### 2020-04-06 paper/paper.tex
Initial commit

%% bare_conf.tex
%% V1.4b
%% 2015/08/26
%% by Michael Shell
%% See:
%% http://www.michaelshell.org/
%% for current contact information.
%%
%% This is a skeleton file demonstrating the use of IEEEtran.cls
%% (requires IEEEtran.cls version 1.8b or later) with an IEEE
%% conference paper.
%%
%% Support sites:
%% http://www.michaelshell.org/tex/ieeetran/
%% http://www.ctan.org/pkg/ieeetran
%% and
%% http://www.ieee.org/

%%*************************************************************************
%% Legal Notice:
%% This code is offered as-is without any warranty either expressed or
%% implied; without even the implied warranty of MERCHANTABILITY or
%% FITNESS FOR A PARTICULAR PURPOSE!
%% User assumes all risk.
%% In no event shall the IEEE or any contributor to this code be liable for
%% any damages or losses, including, but not limited to, incidental,
%% consequential, or any other damages, resulting from the use or misuse
%% of any information contained here.
%%
%% All comments are the opinions of their respective authors and are not
%% necessarily endorsed by the IEEE.
%%
%% This work is distributed under the LaTeX Project Public License (LPPL)
%% ( http://www.latex-project.org/ ) version 1.3, and may be freely used,
%% distributed and modified. A copy of the LPPL, version 1.3, is included
%% in the base LaTeX documentation of all distributions of LaTeX released
%% 2003/12/01 or later.
%% Retain all contribution notices and credits.
%% ** Modified files should be clearly indicated as such, including  **
%% ** renaming them and changing author support contact information. **
%%*************************************************************************

% *** Authors should verify (and, if needed, correct) their LaTeX system  ***
% *** with the testflow diagnostic prior to trusting their LaTeX platform ***
% *** with production work. The IEEE's font choices and paper sizes can   ***
% *** trigger bugs that do not appear when using other class files.       ***                          ***
% The testflow support page is at:
% http://www.michaelshell.org/tex/testflow/

\documentclass[a4paper,conference]{IEEEtran}
% Some Computer Society conferences also require the compsoc mode option,
% but others use the standard conference format.
%
% If IEEEtran.cls has not been installed into the LaTeX system files,
% manually specify the path to it like:
% \documentclass[conference]{../sty/IEEEtran}

% Some very useful LaTeX packages include:
% (uncomment the ones you want to load)

% *** MISC UTILITY PACKAGES ***
%
%\usepackage{ifpdf}
% Heiko Oberdiek's ifpdf.sty is very useful if you need conditional
% compilation based on whether the output is pdf or dvi.
% usage:
% \ifpdf
%   % pdf code
% \else
%   % dvi code
% \fi
% The latest version of ifpdf.sty can be obtained from:
% http://www.ctan.org/pkg/ifpdf
% Also, note that IEEEtran.cls V1.7 and later provides a builtin
% \ifCLASSINFOpdf conditional that works the same way.
% When switching from latex to pdflatex and vice-versa, the compiler may
% have to be run twice to clear warning/error messages.

% *** CITATION PACKAGES ***
%
%\usepackage{cite}
% cite.sty was written by Donald Arseneau
% V1.6 and later of IEEEtran pre-defines the format of the cite.sty package
% \cite{} output to follow that of the IEEE. Loading the cite package will
% result in citation numbers being automatically sorted and properly
% "compressed/ranged". e.g., [1], [9], [2], [7], [5], [6] without using
% cite.sty will become [1], [2], [5]--[7], [9] using cite.sty. cite.sty's
% \cite will automatically add leading space, if needed. Use cite.sty's
% noadjust option (cite.sty V3.8 and later) if you want to turn this off
% such as if a citation ever needs to be enclosed in parenthesis.
% cite.sty is already installed on most LaTeX systems. Be sure and use
% version 5.0 (2009-03-20) and later if using hyperref.sty.
% The latest version can be obtained at:
% http://www.ctan.org/pkg/cite
% The documentation is contained in the cite.sty file itself.
\usepackage{cleveref}

% *** GRAPHICS RELATED PACKAGES ***
%
\usepackage[pdftex]{graphicx}
\DeclareGraphicsExtensions{.pdf,.jpeg,.png}
\usepackage{tabularx}
% graphicx was written by David Carlisle and Sebastian Rahtz. It is
% required if you want graphics, photos, etc. graphicx.sty is already
% installed on most LaTeX systems. The latest version and documentation
% can be obtained at:
% http://www.ctan.org/pkg/graphicx
% Another good source of documentation is "Using Imported Graphics in
% LaTeX2e" by Keith Reckdahl which can be found at:
% http://www.ctan.org/pkg/epslatex
%
% latex, and pdflatex in dvi mode, support graphics in encapsulated
% postscript (.eps) format. pdflatex in pdf mode supports graphics
% in .pdf, .jpeg, .png and .mps (metapost) formats. Users should ensure
% that all non-photo figures use a vector format (.eps, .pdf, .mps) and
% not a bitmapped formats (.jpeg, .png). The IEEE frowns on bitmapped formats
% which can result in "jaggedy"/blurry rendering of lines and letters as
% well as large increases in file sizes.
%
% You can find documentation about the pdfTeX application at:
% http://www.tug.org/applications/pdftex

% *** MATH PACKAGES ***
%
%\usepackage{amsmath}
% A popular package from the American Mathematical Society that provides
% many useful and powerful commands for dealing with mathematics.
%
% Note that the amsmath package sets \interdisplaylinepenalty to 10000
% thus preventing page breaks from occurring within multiline equations. Use:
%\interdisplaylinepenalty=2500
% after loading amsmath to restore such page breaks as IEEEtran.cls normally
% does. amsmath.sty is already installed on most LaTeX systems. The latest
% version and documentation can be obtained at:
% http://www.ctan.org/pkg/amsmath

% *** SPECIALIZED LIST PACKAGES ***
%
%\usepackage{algorithmic}
% algorithmic.sty was written by Peter Williams and Rogerio Brito.
% This package provides an algorithmic environment fo describing algorithms.
% You can use the algorithmic environment in-text or within a figure
% environment to provide for a floating algorithm. Do NOT use the algorithm
% floating environment provided by algorithm.sty (by the same authors) or
% algorithm2e.sty (by Christophe Fiorio) as the IEEE does not use dedicated
% algorithm float types and packages that provide these will not provide
% correct IEEE style captions. The latest version and documentation of
% algorithmic.sty can be obtained at:
% http://www.ctan.org/pkg/algorithms
% Also of interest may be the (relatively newer and more customizable)
% algorithmicx.sty package by Szasz Janos:
% http://www.ctan.org/pkg/algorithmicx

% *** ALIGNMENT PACKAGES ***
%
%\usepackage{array}
% Frank Mittelbach's and David Carlisle's array.sty patches and improves
% the standard LaTeX2e array and tabular environments to provide better
% appearance and additional user controls. As the default LaTeX2e table
% generation code is lacking to the point of almost being broken with
% respect to the quality of the end results, all users are strongly
% advised to use an enhanced (at the very least that provided by array.sty)
% set of table tools. array.sty is already installed on most systems. The
% latest version and documentation can be obtained at:
% http://www.ctan.org/pkg/array

% IEEEtran contains the IEEEeqnarray family of commands that can be used to
% generate multiline equations as well as matrices, tables, etc., of high
% quality.

% *** SUBFIGURE PACKAGES ***
%\ifCLASSOPTIONcompsoc
%  \usepackage[caption=false,font=normalsize,labelfont=sf,textfont=sf]{subfig}
%\else
%  \usepackage[caption=false,font=footnotesize]{subfig}
%\fi
% subfig.sty, written by Steven Douglas Cochran, is the modern replacement
% for subfigure.sty, the latter of which is no longer maintained and is
% incompatible with some LaTeX packages including fixltx2e. However,
% subfig.sty requires and automatically loads Axel Sommerfeldt's caption.sty
% which will override IEEEtran.cls' handling of captions and this will result
% in non-IEEE style figure/table captions. To prevent this problem, be sure
% and invoke subfig.sty's "caption=false" package option (available since
% subfig.sty version 1.3, 2005/06/28) as this is will preserve IEEEtran.cls
% handling of captions.
% Note that the Computer Society format requires a larger sans serif font
% than the serif footnote size font used in traditional IEEE formatting
% and thus the need to invoke different subfig.sty package options depending
% on whether compsoc mode has been enabled.
%
% The latest version and documentation of subfig.sty can be obtained at:
% http://www.ctan.org/pkg/subfig

% *** FLOAT PACKAGES ***
%
%\usepackage{fixltx2e}
% fixltx2e, the successor to the earlier fix2col.sty, was written by
% Frank Mittelbach and David Carlisle. This package corrects a few problems
% in the LaTeX2e kernel, the most notable of which is that in current
% LaTeX2e releases, the ordering of single and double column floats is not
% guaranteed to be preserved. Thus, an unpatched LaTeX2e can allow a
% single column figure to be placed prior to an earlier double column
% figure.
% Be aware that LaTeX2e kernels dated 2015 and later have fixltx2e.sty's
% corrections already built into the system in which case a warning will
% be issued if an attempt is made to load fixltx2e.sty as it is no longer
% needed.
% The latest version and documentation can be found at:
% http://www.ctan.org/pkg/fixltx2e

%\usepackage{stfloats}
% stfloats.sty was written by Sigitas Tolusis. This package gives LaTeX2e
% the ability to do double column floats at the bottom of the page as well
% as the top. (e.g., "\begin{figure*}[!b]" is not normally possible in
% LaTeX2e). It also provides a command:
%\fnbelowfloat
% to enable the placement of footnotes below bottom floats (the standard
% LaTeX2e kernel puts them above bottom floats). This is an invasive package
% which rewrites many portions of the LaTeX2e float routines. It may not work
% with other packages that modify the LaTeX2e float routines. The latest
% version and documentation can be obtained at:
% http://www.ctan.org/pkg/stfloats
% Do not use the stfloats baselinefloat ability as the IEEE does not allow
% \baselineskip to stretch. Authors submitting work to the IEEE should note
% that the IEEE rarely uses double column equations and that authors should try
% to avoid such use. Do not be tempted to use the cuted.sty or midfloat.sty
% packages (also by Sigitas Tolusis) as the IEEE does not format its papers in
% such ways.
% Do not attempt to use stfloats with fixltx2e as they are incompatible.
% Instead, use Morten Hogholm'a dblfloatfix which combines the features
% of both fixltx2e and stfloats:
%
% \usepackage{dblfloatfix}
% The latest version can be found at:
% http://www.ctan.org/pkg/dblfloatfix

% *** PDF, URL AND HYPERLINK PACKAGES ***
%
%\usepackage{url}
% url.sty was written by Donald Arseneau. It provides better support for
% handling and breaking URLs. url.sty is already installed on most LaTeX
% systems. The latest version and documentation can be obtained at:
% http://www.ctan.org/pkg/url
% Basically, \url{my_url_here}.

% *** Do not adjust lengths that control margins, column widths, etc. ***
% *** Do not use packages that alter fonts (such as pslatex).         ***
% There should be no need to do such things with IEEEtran.cls V1.6 and later.
% (Unless specifically asked to do so by the journal or conference you plan
% to submit to, of course. )

% correct bad hyphenation here
\hyphenation{op-tical net-works semi-conduc-tor}

\bibliographystyle{IEEEtran}

\begin{document}
%
% paper title
% Titles are generally capitalized except for words such as a, an, and, as,
% at, but, by, for, in, nor, of, on, or, the, to and up, which are usually
% not capitalized unless they are the first or last word of the title.
% Linebreaks \\ can be used within to get better formatting as desired.
% Do not put math or special symbols in the title.
\title{The Origins of Congestion and\\Network Assisted End-to-End Congestion Control}

% author names and affiliations
% use a multiple column layout for up to three different
% affiliations
\author{\IEEEauthorblockN{Til Mohr}
\IEEEauthorblockA{RWTH Aachen University\\
til.mohr@rwth-aachen.de}}

% conference papers do not typically use \thanks and this command
% is locked out in conference mode. If really needed, such as for
% the acknowledgment of grants, issue a \IEEEoverridecommandlockouts
% after \documentclass

% for over three affiliations, or if they all won't fit within the width
% of the page, use this alternative format:
%
%\author{\IEEEauthorblockN{Michael Shell\IEEEauthorrefmark{1},
%Homer Simpson\IEEEauthorrefmark{2},
%James Kirk\IEEEauthorrefmark{3},
%Montgomery Scott\IEEEauthorrefmark{3} and
%Eldon Tyrell\IEEEauthorrefmark{4}}
%\IEEEauthorblockA{\IEEEauthorrefmark{1}School of Electrical and Computer Engineering\\
%Georgia Institute of Technology,
%Atlanta, Georgia 30332--0250\\ Email: see http://www.michaelshell.org/contact.html}
%\IEEEauthorblockA{\IEEEauthorrefmark{2}Twentieth Century Fox, Springfield, USA\\
%Email: homer@thesimpsons.com}
%\IEEEauthorblockA{\IEEEauthorrefmark{3}Starfleet Academy, San Francisco, California 96678-2391\\
%Telephone: (800) 555--1212, Fax: (888) 555--1212}
%\IEEEauthorblockA{\IEEEauthorrefmark{4}Tyrell Inc., 123 Replicant Street, Los Angeles, California 90210--4321}}

% use for special paper notices
%\IEEEspecialpapernotice{(Invited Paper)}

% make the title area
\maketitle

% no keywords

% For peer review papers, you can put extra information on the cover
% page as needed:
% \ifCLASSOPTIONpeerreview
% \begin{center} \bfseries EDICS Category: 3-BBND \end{center}
% \fi
%
% For peerreview papers, this IEEEtran command inserts a page break and
% creates the second title. It will be ignored for other modes.
\IEEEpeerreviewmaketitle

\section{Introduction}
- Nur Idee bisher!\\
- Rolle des Internets / immer mehr Dienste laufen übers Internet  \textrightarrow{}  (VoiceOverIP...)  \textrightarrow{}  Umso wichtiger wird es, Congestion zu minimieren\\
- Grafik finden, wie Congestion den Einzelnen betrifft (Packet Loss etc.)

\section{Origins of Congestion}
- Wo und Wie entsteht Congestion (in Router/Switches Buffern etc.)

\section{Congestion Avoidance}
- Congestion Avoidance anhand TCP/IP erklären\\
 \textrightarrow{} Slow-Start\\
 \textrightarrow{} RoundTripTime\\
- Vorteile (nur Softwareupdate) und Nachteile (ungenaue Messungen und nicht optimale Maßnahmen)

\section{Network Assisted Congestion Control}
- Congestion Control anhand von ECN (TCP/IP-Erweiterung) erklären\\
 \textrightarrow{} Bit im IP Header setzen + wie TCP damit umgeht\\
  \textrightarrow{} Vorteile (genaue Maßnahmen) und Nachteile (erforderliche Hardware)\\
- XCP (Verallgemeinerung von ECN)
\subsection*{DCTCP}
- DCTCP + Warum nur in Data Center verwendet

\section{Conclusion}

\bibliography{references}

% \begin{thebibliography}{1}
%
% \bibitem{IEEEhowto:kopka}
% H.~Kopka and P.~W. Daly, \emph{A Guide to \LaTeX}, 3rd~ed.\hskip 1em plus
%   0.5em minus 0.4em\relax Harlow, England: Addison-Wesley, 1999.
%
% \end{thebibliography}

% that's all folks
\end{document}

### 2020-04-26 paper/paper.tex
Lots and lots and lots of new stuff

@@
-%% bare_conf.tex
-%% V1.4b
-%% 2015/08/26
-%% by Michael Shell
-%% See:
-%% http://www.michaelshell.org/
-%% for current contact information.
-%%
-%% This is a skeleton file demonstrating the use of IEEEtran.cls
-%% (requires IEEEtran.cls version 1.8b or later) with an IEEE
-%% conference paper.
-%%
-%% Support sites:
-%% http://www.michaelshell.org/tex/ieeetran/
-%% http://www.ctan.org/pkg/ieeetran
-%% and
-%% http://www.ieee.org/
+\documentclass[a4paper,conference]{IEEEtran}
 
-%%*************************************************************************
-%% Legal Notice:
-%% This code is offered as-is without any warranty either expressed or
-%% implied; without even the implied warranty of MERCHANTABILITY or
-%% FITNESS FOR A PARTICULAR PURPOSE!
-%% User assumes all risk.
-%% In no event shall the IEEE or any contributor to this code be liable for
-%% any damages or losses, including, but not limited to, incidental,
-%% consequential, or any other damages, resulting from the use or misuse
-%% of any information contained here.
-%%
-%% All comments are the opinions of their respective authors and are not
-%% necessarily endorsed by the IEEE.
-%%
-%% This work is distributed under the LaTeX Project Public License (LPPL)
-%% ( http://www.latex-project.org/ ) version 1.3, and may be freely used,
-%% distributed and modified. A copy of the LPPL, version 1.3, is included
-%% in the base LaTeX documentation of all distributions of LaTeX released
-%% 2003/12/01 or later.
-%% Retain all contribution notices and credits.
-%% ** Modified files should be clearly indicated as such, including  **
-%% ** renaming them and changing author support contact information. **
-%%*************************************************************************
+% Packages
+\usepackage[pdftex]{graphicx}				% Graphics
+\DeclareGraphicsExtensions{.pdf,.jpeg,.jpg,.png}
+\usepackage{tabularx}						% Tables
+\usepackage{amsmath}						% Math
+\usepackage{mathtools}						% Extra Mathematische Symbole
+\usepackage{algorithmic}					% Algorithmics
+\usepackage{cleveref}						% Citations
+\usepackage{bytefield}						% Bytefield
+\usepackage{tikz}							% tikz-figures
+\usetikzlibrary{positioning,arrows,automata}
 
-% *** Authors should verify (and, if needed, correct) their LaTeX system  ***
-% *** with the testflow diagnostic prior to trusting their LaTeX platform ***
-% *** with production work. The IEEE's font choices and paper sizes can   ***
-% *** trigger bugs that do not appear when using other class files.       ***                          ***
-% The testflow support page is at:
-% http://www.michaelshell.org/tex/testflow/
-
-\documentclass[a4paper,conference]{IEEEtran}
-% Some Computer Society conferences also require the compsoc mode option,
-% but others use the standard conference format.
-%
-% If IEEEtran.cls has not been installed into the LaTeX system files,
-% manually specify the path to it like:
-% \documentclass[conference]{../sty/IEEEtran}
-
-% Some very useful LaTeX packages include:
-% (uncomment the ones you want to load)
-
-% *** MISC UTILITY PACKAGES ***
-%
-%\usepackage{ifpdf}
-% Heiko Oberdiek's ifpdf.sty is very useful if you need conditional
-% compilation based on whether the output is pdf or dvi.
-% usage:
-% \ifpdf
-%   % pdf code
-% \else
-%   % dvi code
-% \fi
-% The latest version of ifpdf.sty can be obtained from:
-% http://www.ctan.org/pkg/ifpdf
-% Also, note that IEEEtran.cls V1.7 and later provides a builtin
-% \ifCLASSINFOpdf conditional that works the same way.
-% When switching from latex to pdflatex and vice-versa, the compiler may
-% have to be run twice to clear warning/error messages.
-
-% *** CITATION PACKAGES ***
-%
-%\usepackage{cite}
-% cite.sty was written by Donald Arseneau
-% V1.6 and later of IEEEtran pre-defines the format of the cite.sty package
-% \cite{} output to follow that of the IEEE. Loading the cite package will
-% result in citation numbers being automatically sorted and properly
-% "compressed/ranged". e.g., [1], [9], [2], [7], [5], [6] without using
-% cite.sty will become [1], [2], [5]--[7], [9] using cite.sty. cite.sty's
-% \cite will automatically add leading space, if needed. Use cite.sty's
-% noadjust option (cite.sty V3.8 and later) if you want to turn this off
-% such as if a citation ever needs to be enclosed in parenthesis.
-% cite.sty is already installed on most LaTeX systems. Be sure and use
-% version 5.0 (2009-03-20) and later if using hyperref.sty.
-% The latest version can be obtained at:
-% http://www.ctan.org/pkg/cite
-% The documentation is contained in the cite.sty file itself.
-\usepackage{cleveref}
-
-% *** GRAPHICS RELATED PACKAGES ***
-%
-\usepackage[pdftex]{graphicx}
-\DeclareGraphicsExtensions{.pdf,.jpeg,.png}
-\usepackage{tabularx}
-% graphicx was written by David Carlisle and Sebastian Rahtz. It is
-% required if you want graphics, photos, etc. graphicx.sty is already
-% installed on most LaTeX systems. The latest version and documentation
-% can be obtained at:
-% http://www.ctan.org/pkg/graphicx
-% Another good source of documentation is "Using Imported Graphics in
-% LaTeX2e" by Keith Reckdahl which can be found at:
-% http://www.ctan.org/pkg/epslatex
-%
-% latex, and pdflatex in dvi mode, support graphics in encapsulated
-% postscript (.eps) format. pdflatex in pdf mode supports graphics
-% in .pdf, .jpeg, .png and .mps (metapost) formats. Users should ensure
-% that all non-photo figures use a vector format (.eps, .pdf, .mps) and
-% not a bitmapped formats (.jpeg, .png). The IEEE frowns on bitmapped formats
-% which can result in "jaggedy"/blurry rendering of lines and letters as
-% well as large increases in file sizes.
-%
-% You can find documentation about the pdfTeX application at:
-% http://www.tug.org/applications/pdftex
-
-% *** MATH PACKAGES ***
-%
-%\usepackage{amsmath}
-% A popular package from the American Mathematical Society that provides
-% many useful and powerful commands for dealing with mathematics.
-%
-% Note that the amsmath package sets \interdisplaylinepenalty to 10000
-% thus preventing page breaks from occurring within multiline equations. Use:
-%\interdisplaylinepenalty=2500
-% after loading amsmath to restore such page breaks as IEEEtran.cls normally
-% does. amsmath.sty is already installed on most LaTeX systems. The latest
-% version and documentation can be obtained at:
-% http://www.ctan.org/pkg/amsmath
-
-% *** SPECIALIZED LIST PACKAGES ***
-%
-%\usepackage{algorithmic}
-% algorithmic.sty was written by Peter Williams and Rogerio Brito.
-% This package provides an algorithmic environment fo describing algorithms.
-% You can use the algorithmic environment in-text or within a figure
-% environment to provide for a floating algorithm. Do NOT use the algorithm
-% floating environment provided by algorithm.sty (by the same authors) or
-% algorithm2e.sty (by Christophe Fiorio) as the IEEE does not use dedicated
-% algorithm float types and packages that provide these will not provide
-% correct IEEE style captions. The latest version and documentation of
-% algorithmic.sty can be obtained at:
-% http://www.ctan.org/pkg/algorithms
-% Also of interest may be the (relatively newer and more customizable)
-% algorithmicx.sty package by Szasz Janos:
-% http://www.ctan.org/pkg/algorithmicx
-
-% *** ALIGNMENT PACKAGES ***
-%
-%\usepackage{array}
-% Frank Mittelbach's and David Carlisle's array.sty patches and improves
-% the standard LaTeX2e array and tabular environments to provide better
-% appearance and additional user controls. As the default LaTeX2e table
-% generation code is lacking to the point of almost being broken with
-% respect to the quality of the end results, all users are strongly
-% advised to use an enhanced (at the very least that provided by array.sty)
-% set of table tools. array.sty is already installed on most systems. The
-% latest version and documentation can be obtained at:
-% http://www.ctan.org/pkg/array
-
-% IEEEtran contains the IEEEeqnarray family of commands that can be used to
-% generate multiline equations as well as matrices, tables, etc., of high
-% quality.
-
-% *** SUBFIGURE PACKAGES ***
-%\ifCLASSOPTIONcompsoc
-%  \usepackage[caption=false,font=normalsize,labelfont=sf,textfont=sf]{subfig}
-%\else
-%  \usepackage[caption=false,font=footnotesize]{subfig}
-%\fi
-% subfig.sty, written by Steven Douglas Cochran, is the modern replacement
-% for subfigure.sty, the latter of which is no longer maintained and is
-% incompatible with some LaTeX packages including fixltx2e. However,
-% subfig.sty requires and automatically loads Axel Sommerfeldt's caption.sty
-% which will override IEEEtran.cls' handling of captions and this will result
-% in non-IEEE style figure/table captions. To prevent this problem, be sure
-% and invoke subfig.sty's "caption=false" package option (available since
-% subfig.sty version 1.3, 2005/06/28) as this is will preserve IEEEtran.cls
-% handling of captions.
-% Note that the Computer Society format requires a larger sans serif font
-% than the serif footnote size font used in traditional IEEE formatting
-% and thus the need to invoke different subfig.sty package options depending
-% on whether compsoc mode has been enabled.
-%
-% The latest version and documentation of subfig.sty can be obtained at:
-% http://www.ctan.org/pkg/subfig
-
-% *** FLOAT PACKAGES ***
-%
-%\usepackage{fixltx2e}
-% fixltx2e, the successor to the earlier fix2col.sty, was written by
-% Frank Mittelbach and David Carlisle. This package corrects a few problems
-% in the LaTeX2e kernel, the most notable of which is that in current
-% LaTeX2e releases, the ordering of single and double column floats is not
-% guaranteed to be preserved. Thus, an unpatched LaTeX2e can allow a
-% single column figure to be placed prior to an earlier double column
-% figure.
-% Be aware that LaTeX2e kernels dated 2015 and later have fixltx2e.sty's
-% corrections already built into the system in which case a warning will
-% be issued if an attempt is made to load fixltx2e.sty as it is no longer
-% needed.
-% The latest version and documentation can be found at:
-% http://www.ctan.org/pkg/fixltx2e
-
-%\usepackage{stfloats}
-% stfloats.sty was written by Sigitas Tolusis. This package gives LaTeX2e
-% the ability to do double column floats at the bottom of the page as well
-% as the top. (e.g., "\begin{figure*}[!b]" is not normally possible in
-% LaTeX2e). It also provides a command:
-%\fnbelowfloat
-% to enable the placement of footnotes below bottom floats (the standard
-% LaTeX2e kernel puts them above bottom floats). This is an invasive package
-% which rewrites many portions of the LaTeX2e float routines. It may not work
-% with other packages that modify the LaTeX2e float routines. The latest
-% version and documentation can be obtained at:
-% http://www.ctan.org/pkg/stfloats
-% Do not use the stfloats baselinefloat ability as the IEEE does not allow
-% \baselineskip to stretch. Authors submitting work to the IEEE should note
-% that the IEEE rarely uses double column equations and that authors should try
-% to avoid such use. Do not be tempted to use the cuted.sty or midfloat.sty
-% packages (also by Sigitas Tolusis) as the IEEE does not format its papers in
-% such ways.
-% Do not attempt to use stfloats with fixltx2e as they are incompatible.
-% Instead, use Morten Hogholm'a dblfloatfix which combines the features
-% of both fixltx2e and stfloats:
-%
-% \usepackage{dblfloatfix}
-% The latest version can be found at:
-% http://www.ctan.org/pkg/dblfloatfix
-
-% *** PDF, URL AND HYPERLINK PACKAGES ***
-%
-%\usepackage{url}
-% url.sty was written by Donald Arseneau. It provides better support for
-% handling and breaking URLs. url.sty is already installed on most LaTeX
-% systems. The latest version and documentation can be obtained at:
-% http://www.ctan.org/pkg/url
-% Basically, \url{my_url_here}.
-
-% *** Do not adjust lengths that control margins, column widths, etc. ***
-% *** Do not use packages that alter fonts (such as pslatex).         ***
-% There should be no need to do such things with IEEEtran.cls V1.6 and later.
-% (Unless specifically asked to do so by the journal or conference you plan
-% to submit to, of course. )
+\begin{filecontents}{tcp.data}
+#Transmission round		Congestion Window Size
+1		1
+2		2
+3		4
+4		8
+5		16
+6		32
+7		16
+8		17
+9		18
+10		19
+11		20
+12		10
+13		5
+14		6
+15		7
+\end{filecontents}
 
@@
 \begin{document}
-%
+
 % paper title
-% Titles are generally capitalized except for words such as a, an, and, as,
-% at, but, by, for, in, nor, of, on, or, the, to and up, which are usually
-% not capitalized unless they are the first or last word of the title.
-% Linebreaks \\ can be used within to get better formatting as desired.
-% Do not put math or special symbols in the title.
 \title{The Origins of Congestion and\\Network Assisted End-to-End Congestion Control}
 
-% author names and affiliations
-% use a multiple column layout for up to three different
-% affiliations
+% author
 \author{\IEEEauthorblockN{Til Mohr}
-\IEEEauthorblockA{RWTH Aachen University\\
-til.mohr@rwth-aachen.de}}
-
-% conference papers do not typically use \thanks and this command
-% is locked out in conference mode. If really needed, such as for
-% the acknowledgment of grants, issue a \IEEEoverridecommandlockouts
-% after \documentclass
-
-% for over three affiliations, or if they all won't fit within the width
-% of the page, use this alternative format:
-%
-%\author{\IEEEauthorblockN{Michael Shell\IEEEauthorrefmark{1},
-%Homer Simpson\IEEEauthorrefmark{2},
-%James Kirk\IEEEauthorrefmark{3},
-%Montgomery Scott\IEEEauthorrefmark{3} and
-%Eldon Tyrell\IEEEauthorrefmark{4}}
-%\IEEEauthorblockA{\IEEEauthorrefmark{1}School of Electrical and Computer Engineering\\
-%Georgia Institute of Technology,
-%Atlanta, Georgia 30332--0250\\ Email: see http://www.michaelshell.org/contact.html}
-%\IEEEauthorblockA{\IEEEauthorrefmark{2}Twentieth Century Fox, Springfield, USA\\
-%Email: homer@thesimpsons.com}
-%\IEEEauthorblockA{\IEEEauthorrefmark{3}Starfleet Academy, San Francisco, California 96678-2391\\
-%Telephone: (800) 555--1212, Fax: (888) 555--1212}
-%\IEEEauthorblockA{\IEEEauthorrefmark{4}Tyrell Inc., 123 Replicant Street, Los Angeles, California 90210--4321}}
+\IEEEauthorblockA{RWTH Aachen University\\til.mohr@rwth-aachen.de}}
 
@@
 \maketitle
-
-% no keywords
-
-% For peer review papers, you can put extra information on the cover
-% page as needed:
-% \ifCLASSOPTIONpeerreview
-% \begin{center} \bfseries EDICS Category: 3-BBND \end{center}
-% \fi
-%
-% For peerreview papers, this IEEEtran command inserts a page break and
-% creates the second title. It will be ignored for other modes.
 \IEEEpeerreviewmaketitle
 
+\begin{abstract}
+This article presents the most common origins of network congestion, how different types of algorithms perceive and handle congestion and what obstacles they face.
+\end{abstract}
+
 \section{Introduction}
-- Nur Idee bisher!\\
-- Rolle des Internets / immer mehr Dienste laufen übers Internet  \textrightarrow{}  (VoiceOverIP...)  \textrightarrow{}  Umso wichtiger wird es, Congestion zu minimieren\\
-- Grafik finden, wie Congestion den Einzelnen betrifft (Packet Loss etc.)
+Congestion was, is and always will be a problem on the Internet, appearing in various forms with different symptoms causing minor, but also major challenges.
+Data sent over the internet is split up in smaller packets that all try to reach their destination the fastest, whether it is by going the shortest or the least congested route.
+In an optimal network all computers, servers etc.
+would have a direct link to each other, thus reducing the importance of bandwidth and  consequently removing the need for buffers at all.
+However, the larger the network would get, the amount of connections would grow quadratic.
+\\Networks today consist of network nodes (routers, switches, ...) which all try to route internet packets as quick as possible whilst keeping the required hardware to a minimum.
+Ideally there would be a steady flow of packets being transmitted at maximum rate.
+However, when a internet node receives a packet while the outgoing connection still is in use, the packet has to be stored in a buffer until the outgoing connection freed up again.
+Therefore, those storage units are essential to every node on a network, but without proper management even well-sized buffers could lead to network problems like packet-loss, jitter and connection delay \cite{appenzeller2004sizing}	.
+\\This paper will present origins of network congestion and analyze how different types of algorithms try to avoid and minimize congestion.
 
 \section{Origins of Congestion}
-- Wo und Wie entsteht Congestion (in Router/Switches Buffern etc.)
+With the ever-growing amount of internet services not only data-centers, but the whole internet in general face new problems with each packet sent.
+Therefore, the global reduction of those complications are an important task to improve the quality of all internet services.
+However, as it is for solving any problem, the origin of those problems must be identified first.
+\\Network delay can generally speaking be categorized into three different types of delay \cite{gettys2012bufferbloat}:
+
+\begin{itemize}
+\item 
+\textit{Transmission delay} describes the delay of transmitting data packets between two internet nodes.
+Typically, packets sent from a origin to a destination node are sent over various different internet nodes, all of which could have a different bandwidth.
+Therefore, a packet send from \textbf{node 1} to \textbf{node n} over \textbf{nodes 2 to n-1} cannot be sent at a faster rate than the lowest bandwidth on the packets path (bottleneck) [Figure \ref{fig:transmission delay}].\[Rate_{max} = \min\{bandwidth_{i}\mid i\in\{1,...,n-1\}\}\]
+
+\item
+\textit{Processing delay} is the time needed to process an individual packet by an internet node.
+This usually means processing the packets IP header, finding its destination and possibly queuing it in buffers.
+
+\item
+\textit{Queuing delay} is the time a package spent waiting for it to be processed or transmitted.
+\end{itemize}
+\begin{figure}
+	\begin{center}
+	\begin{tikzpicture}[->, >=latex, node distance = 1cm and 2.5cm, semithick]
+		% Nodes
+		\node[state] (A)  			   	{1};
+		\node[state] (B) [right=of A] 	{2};
+		\node[state] (C) [below=of B] 	{n-1};
+		\node[state] (D) [right=of C] 	{n};
+		%Arrows
+		\path	
+		(A) edge [above] node {$bandwidth_{1}$} (B)
+		(B) edge [dotted] (C)
+		(C) edge [above] node {$bandwidth_{n-1}$} (D)
+		;
+	\end{tikzpicture}
+	\end{center}
+	\caption{Path a packet takes through the network}
+	\label{fig:transmission delay}
+\end{figure}
+
+Whereas processing delay is almost not noticeable and transmission delay could be improved by upgrading cables or transmission protocols, queuing delay has various causes and therefore is the  most difficult one to reduce.
+There are many different ways to implement buffers in a network node and the effectiveness of each implementation varies from where it is used.
+Traffic in residential network nodes for example differ from nodes used in large data-centers in every aspect, thus hardware and software in those components should be adjusted accordingly.
+\\Sizing buffers is a complex task.
+On the one hand read and write times should as short as possible to minimize additional delay.
+On the other hand buffers should be as big as needed to be robust for future network changes and therefore store lots of data if needed while also keeping in mind the requirements congestion detection algorithms have in order to adjust for congestion accordingly \cite{staff2012bufferbloat}.
+Unifying those goals is very challenging as it leads to increasing costs and physical space occupied by those buffers.
+Additionally, a 2004 paper by Appenzeller found that 99\% of buffers in network nodes, which were sized on a rule-of-thumb rule based on the TCP protocol \cite{villamizar1994high}, could be removed without losing, but rather gaining performance \cite{appenzeller2004sizing}.
+Clearly, balancing speed and size is difficult and thus buffer sizes should be adjusted to their area of use accordingly.
+\begin{figure}
+  \centering
+    \includegraphics[width=0.9\linewidth]{figures/Throughput-Delay}
+  \caption{Relationship between throughput and delay \cite{gettys2012bufferbloat}}
+  \label{fig:throughput-delay-relationship}
+\end{figure}
+\\This isn't the only major cause of queuing delay.
+Without proper buffer management even nodes with well sized buffers will experience performance loss.
+Proper queue management is able to compensate for the unpredictable fluidity of network traffic.
+It is an difficult task to buffer the right amount of packages to keep transmission flow consistent over time.
+As the number of packets being transmitted grows, the \textit{throughput} of the network node increases until packets are being transmitted at bottleneck rate.
+After this, the transmission rate cannot increase anymore, which leads to packets being buffered along their path to prevent them being lost [Figure \ref{fig:throughput-delay-relationship}]\cite{gettys2012bufferbloat}.
+If the incoming amount of traffic keeps to exceed the amount of outgoing traffic, those buffers will fill up (\textit{over-buffering}), which will eventually lead to \textit{bufferbloat} \cite{Allman13commentson,6886125}.
+
+\subsection*{Bufferbloat}
+Bufferbloat refers to frequently full and excessively large buffers in internet nodes.
+This leads to partly unnecessary network delay and damages congestion avoidance algorithms, such as classic TCP \cite{gettys2012bufferbloat,chen2014bufferbloat}.
+Buffering is needed to provide space to queued packets waiting for transmission resulting in a minimization of packet loss.
+In the past however, buffer sizes were smaller than today's thus less packets could be queued and more were dropped, signaling transmission protocols the presence of congestion in this node, so those protocols can adjust.
+With memory getting cheaper over the years, buffers saw an increase in size.
+This has the negative impact, that more packets can be queued before any are dropped, thus those adjustments of congestion avoidance protocols kick in later than usual.
+Thus, buffers are being filled in the whole network which results in an extreme increase in network delay (\textit{latency}) and fluctuation (\textit{jitter}) \cite{gettys2012bufferbloat,staff2012bufferbloat,chen2014bufferbloat}.
+\\So Services, which require low latency at a high data-rate, such as Streaming, will experience a decrease in quality.
 
 \section{Congestion Avoidance}
-- Congestion Avoidance anhand TCP/IP erklären\\
- \textrightarrow{} Slow-Start\\
- \textrightarrow{} RoundTripTime\\
-- Vorteile (nur Softwareupdate) und Nachteile (ungenaue Messungen und nicht optimale Maßnahmen)
+As already mentioned, congestion detection algorithms make up an important part in congestion avoidance protocols.
+\textit{TCP} is such an protocol, based on the transport layer, which adapts its send rate according to available network bandwidth \cite{1209197,jacobson1992tcp}.
+To fully understand, how TCP's congestion avoidance algorithm functions, you will have to understand how TCP works.
+TCP lives on the transport layer.
+It communicates with the application layer via ports, splits up data into packets (and vice versa), and transmits those packages over the network with the use of the internet protocol, thus TCP is often also referred to as \textit{TCP/IP}.
+It is the interface between network components.
+Each of those communication packages have different purposes, but underlay the same structure - the protocol header.
+The TCP header is split up into multiple segments [Figure \ref{fig:TCP_Header}]\cite{jacobson1992tcp,huston2000tcp}:
+\begin{figure}
+\centering
+\begin{bytefield}[bitheight=2.2\baselineskip]{32}
+	\bitbox{16}{Source Port} & \bitbox{16}{Destination Port} \\
+	\bitbox{32}{Sequence Number} \\
+	\bitbox{32}{Acknowledgment Number} \\
+	\bitbox{4}{Data Offset} & \bitbox{6}{Reserved} &
+\bitbox{1}{\tiny U\\R\\G} & \bitbox{1}{\tiny A\\C\\K} &
+\bitbox{1}{\tiny P\\S\\H} & \bitbox{1}{\tiny R\\S\\T} &
+\bitbox{1}{\tiny S\\Y\\N} & \bitbox{1}{\tiny F\\I\\N} &
+\bitbox{16}{Window Size} \\
+\bitbox{16}{Checksum} & \bitbox{16}{Urgent Pointer} \\
+\bitbox{32}{Options} \\
+\bitbox{32}{Data}
+\end{bytefield}
+\caption{TCP Header}
+\label{fig:TCP_Header}
+\end{figure}
+\begin{itemize}
+\item \textit{Source Port} and \textit{Destination Port} specify the ports on which TCP should communicate with the application layer.
+
+\item The \textit{Sequence Numbers} and \textit{Acknowledgment Numbers} specify how much data has been sent so far.
+All bytes in a TCP connection are numbered in increasing order, with both sides of the connection keeping track of those numbers.
+This way packet loss can be detected.
+
+\item TCP can set six different bit-flags in its header.
+The most important one is the \textit{ACK} flag (Acknowledgement bit).
+It is used whenever a packet receives the destination and an acknowledgement packet has to be sent back.
+The \textit{SYN} and \textit{FIN} flags initialize and terminate a connection.
+
+\item The \textit{Window Size} is the maximum amount of data (in bytes) a receiver is willing to receive at any time.
+It is being negotiated at the initialization of a new connection and can only be negotiated again with the \textit{SYN} flag set.
+It is important to set it appropriately to prevent unnecessary packet loss or delay, as the receiver couldn't process the data in time \cite{jacobson1992tcp}.
+\end{itemize}
+TCP uses a handshake system.
+Whenever a client sends a data packet to the server, the server waits until it has received enough data (specified by the window size), and then sends back an acknowledgment package, signaling the successful transmission of data.
+There are also other handshake operations used, for example for establishing or terminating a connection, however those have no further importance to the congestion avoidance protocol \cite{1209197,jacobson1992tcp,huston2000tcp,jacobson1995congestion}.
+
+\subsection{Perceiving Congestion}
+Both partners of a connection are either a sender or a receiver.
+The sender keeps track of the time it takes from when a package sent until the acknowledgment is received.
+This is also known as \textit{round-trip time} (RTT), where a large RTT can indicate some congestion along the path.
+Some variances of TCP make use of this feature in their congestion avoidance algorithms, yet classic TCP does not \cite{huston2000tcp,jacobson1995congestion}.
+RTT also plays an very important role in the detection of packet loss, which further imply congestion.
+The sender side keeps track of the RTT while it is being measured and compares it to the \textit{retransmission timeout} (RTO) \cite{jacobson1992tcp}.
+RTO time is being calculated by estimating the mean and variance of the RTT.
+Whenever 	packets are not acknowledged within this RTO intervall, the package is being detected as lost and will be retransmitted.
+Therefore, it is essential to TCP performance, that the RTO is being determined accurately, as a too low RTO could prompt unnecessary retransmissions and too high RTO could slow down the application TCP is serving \cite{huston2000tcp,jacobson1995congestion}.\\
+Note however, a high RTT can also have other causes than packet loss or congestion.
+A short term drastic decrease in the receivers processing performance can lead to packets being buffered at the receivers end for a prolonged period of time, thus resulting in a higher RTT \cite{huston2000tcp}.
+
+\subsection{Adjusting for Congestion}
+To prevent congestion altogether, packet flow inside a network would have to be conservative, i.e.
+a new packet isn't being sent into the network before an old packet leaves the network.
+A connection in this state is also being described as being \textit{in equilibrium} \cite{jacobson1992tcp}.
+Whenever a connection is in equilibrium, congestion in the network cannot be caused by this TCP connection, thus also controlling congestion.
+TCPs congestion avoidance algorithm can be categorized into three sub-algorithms, that all have the goal to keep the connection in equilibrium, no matter how congestion was perceived.
+Here, a variable called \textit{congestion window} (cwnd) has an important role.
+It describes the maximum amount of packages send during a interval.
+An additional mechanism of keeping a connection in equilibrium is, that each acknowledgement package received by the sender opens the space for a new package to be sent into the network \cite{jacobson1992tcp}.
+
+\subsubsection*{Slow-Start}
+The \textit{slow-start} algorithm is used when a connection has just been established or when multiple packet losses were detected.
+The cwnd variable is set to one package per estimated RTT, because no information about network congestion has been collected yet.
+Whenever an acknowledgment package is being received, cwnd will increase by one package.
+This exponential growth in packages helps to get the packet send rate to the maximum the network can handle quite quickly, resulting in almost immediately maximum network performance \cite{jacobson1992tcp}.
+Slow-start will only increase if cwnd
+
+\subsubsection*{Adaption}
+When the network is congested, buffers in network nodes will fill up exponentially \cite{jacobson1992tcp}.
+Therefore, the network can only stabilize if the amount of packets sent decreases at least as quickly as the queues are growing.
+The formula \cite{jacobson1992tcp} \[cwnd_{i+1}=d*cwnd_{i}, 0<d<1\] has an exponential decrease over time, thus resulting in the network recovering.\\
+If the network isn't congested, the sender should try to raise cwnd slowly to reach the networks maximum throughput again.
+A multiplicative, exponential increase over time, like the slow-start algorithm isn't the best choice here.
+Increasing the amounts of packets sent drastically while cwnd still is high can result in the network being challenge the networks capacity, resulting in heavy congestion.
+Therefore, the best option to raise cwnd to its limits is an additive increase \cite{jacobson1992tcp}:\[cwnd_{i+1}=cwnd_{i} + u\] Van Jacobson proposes in his 1992 paper about TCP congestion avoidance that $d$ will be set to $\frac{1}{2}$ and $u$ will be set to $\frac{1}{cwnd_{i}}$.
+Thus, cwnd can only increase by a maximum of one packet per RTT leading to a smooth approach to the networks limit \cite{jacobson1992tcp}.
+
+\begin{figure}
+  \centering
+    \begin{tikzpicture}[y=.15cm, x=.45cm,font=\sffamily]
+    	% Ranges and dotted lines
+    	\fill[yellow!10]	(0,0) rectangle (6,35);
+    	\fill[red!10] 		(6,0) rectangle (7,35);
+    	\fill[green!10] 	(7,0) rectangle (11,35);
+    	\fill[red!10]		(11,0) rectangle (13,35);
+    	\fill[green!10]		(13,0) rectangle (15.3,35);
+    	\draw[dotted] 		(6,0) -- (6,35);
+    	\draw[dotted] 		(7,0) -- (7,35);
+    	\draw[dotted] 		(11,0) -- (11,35);
+    	\draw[dotted] 		(13,0) -- (13,35);
+    	% Axis
+    	\draw[->] 			(0,0) -- coordinate (x axis mid) (17,0);
+    	\draw[->] 			(0,0) -- coordinate (y axis mid) (0,35);
+    	% Labels
+    	\node[below=0.1cm] at (x axis mid) {\scriptsize Transmission round};
+    	\node[rotate=90, above=0.1cm] at (y axis mid) {\scriptsize Congestion window size};
+    	% Plot
+    	\draw plot[mark=*]
+    		file {tcp.data};
+    \end{tikzpicture}
+  \caption{TCP congestion avoidance protocols. Yellow represents the slow-start sequence, red and green symbolize detected congestion}
+  \label{fig:tcp_congestion_avoidance_protocols}
+\end{figure}
 
 \section{Network Assisted Congestion Control}
-- Congestion Control anhand von ECN (TCP/IP-Erweiterung) erklären\\
- \textrightarrow{} Bit im IP Header setzen + wie TCP damit umgeht\\
-  \textrightarrow{} Vorteile (genaue Maßnahmen) und Nachteile (erforderliche Hardware)\\
-- XCP (Verallgemeinerung von ECN)
+Congestion Control anhand von ECN (TCPIP-Erweiterung) erklaeren
+Bit im IP Header setzen + wie TCP damit umgeht
+Vorteile (genaue Maßnahmen) und Nachteile (erforderliche Hardware)
+XCP (Verallgemeinerung von ECN)
 \subsection*{DCTCP}
-- DCTCP + Warum nur in Data Center verwendet
+DCTCP + Warum nur in Data Center verwendet
 
@@
 
-% \begin{thebibliography}{1}
-%
-% \bibitem{IEEEhowto:kopka}
-% H.~Kopka and P.~W. Daly, \emph{A Guide to \LaTeX}, 3rd~ed.\hskip 1em plus
-%   0.5em minus 0.4em\relax Harlow, England: Addison-Wesley, 1999.
-%
-% \end{thebibliography}
-
-% that's all folks
 \end{document}

### 2020-05-02 paper/paper.tex
Almost done?

@@
 \usepackage{algorithmic}					% Algorithmics
+\usepackage{hyperref}						% More references
 \usepackage{cleveref}						% Citations
 \usepackage{bytefield}						% Bytefield
+\usepackage{moeptikz}						% tikz-packages for networking symbols
 \usepackage{tikz}							% tikz-figures
 \usetikzlibrary{positioning,arrows,automata}
+\tikzstyle{line} = [draw,thick,-latex]
+\tikzstyle{transition} = [font=\small]
 
+% data for figure
 \begin{filecontents}{tcp.data}
@@
 Typically, packets sent from a origin to a destination node are sent over various different internet nodes, all of which could have a different bandwidth.
-Therefore, a packet send from \textbf{node 1} to \textbf{node n} over \textbf{nodes 2 to n-1} cannot be sent at a faster rate than the lowest bandwidth on the packets path (bottleneck) [Figure \ref{fig:transmission delay}].\[Rate_{max} = \min\{bandwidth_{i}\mid i\in\{1,...,n-1\}\}\]
+Therefore, a packet send from \textbf{node 1} to \textbf{node n} over \textbf{nodes 2 to n-1} cannot be sent at a faster rate than the lowest bandwidth on the packets path (bottleneck) [Figure \ref{fig:transmission delay}].\[Rate_{max} \leftarrow \min\{bandwidth_{i}\mid i\in\{1,...,n-1\}\}\]
 
@@
 Clearly, balancing speed and size is difficult and thus buffer sizes should be adjusted to their area of use accordingly.
-\begin{figure}
-  \centering
-    \includegraphics[width=0.9\linewidth]{figures/Throughput-Delay}
-  \caption{Relationship between throughput and delay \cite{gettys2012bufferbloat}}
-  \label{fig:throughput-delay-relationship}
-\end{figure}
 \\This isn't the only major cause of queuing delay.
@@
 As the number of packets being transmitted grows, the \textit{throughput} of the network node increases until packets are being transmitted at bottleneck rate.
-After this, the transmission rate cannot increase anymore, which leads to packets being buffered along their path to prevent them being lost [Figure \ref{fig:throughput-delay-relationship}]\cite{gettys2012bufferbloat}.
+After this, the transmission rate cannot increase anymore, which leads to packets being buffered along their path to prevent them being lost \cite{gettys2012bufferbloat}.
 If the incoming amount of traffic keeps to exceed the amount of outgoing traffic, those buffers will fill up (\textit{over-buffering}), which will eventually lead to \textit{bufferbloat} \cite{Allman13commentson,6886125}.
@@
 \subsection{Perceiving Congestion}
+\label{P_C_marker}
 Both partners of a connection are either a sender or a receiver.
@@
 RTO time is being calculated by estimating the mean and variance of the RTT.
-Whenever 	packets are not acknowledged within this RTO intervall, the package is being detected as lost and will be retransmitted.
-Therefore, it is essential to TCP performance, that the RTO is being determined accurately, as a too low RTO could prompt unnecessary retransmissions and too high RTO could slow down the application TCP is serving \cite{huston2000tcp,jacobson1995congestion}.\\
+Whenever packets are not acknowledged within this RTO intervall, the package is being perceived as lost and will be retransmitted.
+Therefore, it is essential to TCP performance, that the RTO is being determined accurately, as a too low RTO could prompt unnecessary retransmissions and too high RTO could slow down the application TCP is serving \cite{huston2000tcp,jacobson1995congestion}.
+To implicitly notify TCP that congestion occurs, congested network nodes will occasionally drop packets after a certain threshold of queued packets has been exceeded.
+Usually, this is quite smaller the maximum queue length, therefore avoiding actions can be taken in time \cite{ramakrishnan1999proposal}.\\
 Note however, a high RTT can also have other causes than packet loss or congestion.
@@
 \subsection{Adjusting for Congestion}
+\label{AfC}
 To prevent congestion altogether, packet flow inside a network would have to be conservative, i.e.
@@
 An additional mechanism of keeping a connection in equilibrium is, that each acknowledgement package received by the sender opens the space for a new package to be sent into the network \cite{jacobson1992tcp}.
+The principles TCP uses to adjust for congestion are based on \textit{Additive-Increase Multiplicative-Decrease} (AIMD) principle.
 
@@
 This exponential growth in packages helps to get the packet send rate to the maximum the network can handle quite quickly, resulting in almost immediately maximum network performance \cite{jacobson1992tcp}.
-Slow-start will only increase if cwnd
 
@@
 Therefore, the network can only stabilize if the amount of packets sent decreases at least as quickly as the queues are growing.
-The formula \cite{jacobson1992tcp} \[cwnd_{i+1}=d*cwnd_{i}, 0<d<1\] has an exponential decrease over time, thus resulting in the network recovering.\\
+The formula \cite{jacobson1992tcp} \[cwnd_{i+1} \leftarrow d*cwnd_{i}, 0<d<1\] has an exponential decrease over time, thus resulting in the network recovering.\\
 If the network isn't congested, the sender should try to raise cwnd slowly to reach the networks maximum throughput again.
@@
 Increasing the amounts of packets sent drastically while cwnd still is high can result in the network being challenge the networks capacity, resulting in heavy congestion.
-Therefore, the best option to raise cwnd to its limits is an additive increase \cite{jacobson1992tcp}:\[cwnd_{i+1}=cwnd_{i} + u\] Van Jacobson proposes in his 1992 paper about TCP congestion avoidance that $d$ will be set to $\frac{1}{2}$ and $u$ will be set to $\frac{1}{cwnd_{i}}$.
+Therefore, the best option to raise cwnd to its limits is an additive increase \cite{jacobson1992tcp}:\[cwnd_{i+1} \leftarrow cwnd_{i} + u\] Van Jacobson proposes in his 1992 paper about TCP congestion avoidance that $d$ will be set to $\frac{1}{2}$ and $u$ will be set to $\frac{1}{cwnd_{i}}$.
 Thus, cwnd can only increase by a maximum of one packet per RTT leading to a smooth approach to the networks limit \cite{jacobson1992tcp}.
@@
 
+\begin{figure*}[tb]
+	\centering
+		\begin{tikzpicture}[->, >=latex, node distance = 2.5cm, semithick]
+			%network
+			\node[client] (H1) {};
+			\node[router] (R1) [right=of H1] {};
+			\node[router] (R2) [right=of R1] {};
+			\node[router] (R3) [right=of R2] {};
+			\node[client] (H2) [right=of R3] {};
+			\path [line] (H1) -- (R1);
+			\path [line] (R1) -- (R2);
+			\path [line] (R2) -- (R3);
+			\path [line] (R3) -- (H2);
+			\path [line, dotted] (H2) -- ++(0,-35pt) -| (H1);
+			\node[below=0.8cm of R2] (text) [transition] {Congested Signal};
+
+			%packets
+			\node[messageclosed,scale=0.8, fill=black!10] (P1) [above right=0.2cm and 1cm of H1] {};
+			\node[messageclosed,scale=0.8, fill=black!50] (P2) [above left=0.7cm and 0.3cm of R2] {};
+			\node[messageclosed,scale=0.8, fill=black!10] (P3) [above left=0.4cm and 0.2cm of R2] {};
+			\node[messageclosed,scale=0.8, fill=black!10] (P4) [above left=0.1cm and 0.1cm of R2] {};
+			\node[messageclosed,scale=0.8, fill=black!50] (P5) [above right=0.2cm and 1cm of R3] {};
+			\node[above=1.5cm of R1] (P) [transition] {Packet};
+			\node[above right = 1.5cm and 0.7cm of R2] (CN) [transition] {Congested Node};
+			\node[above=1.5 of H2] (MP) [transition] {Marked Packet};
+			\path [line] (P) -- (P1);
+			\path [line] (CN) -- (R2);
+			\path [line] (MP) -- (P5);
+		\end{tikzpicture}
+	\caption{TCP/IP with Explicit Congestion Notification}
+	\label{fig:ecn}
+\end{figure*}
+
 \section{Network Assisted Congestion Control}
-Congestion Control anhand von ECN (TCPIP-Erweiterung) erklaeren
-Bit im IP Header setzen + wie TCP damit umgeht
-Vorteile (genaue Maßnahmen) und Nachteile (erforderliche Hardware)
-XCP (Verallgemeinerung von ECN)
-\subsection*{DCTCP}
-DCTCP + Warum nur in Data Center verwendet
+The problem that comes with classic TCP is that the congestion detection algorithms are very implicit.
+As already mentioned, packet loss for example may have different causes other than network congestion, however, the congestion avoidance algorithm will still interpret it as congestion, thus possibly resulting in unnecessary network delay.
+Therefore, exactly determining whether the network is congested or not is crucial to keep a network stable.
+Here, the network layer comes into play.
+Because most communication should be hidden along the packets path, all network nodes will only communicate on the internet layer (where IP lives), but not above (where TCP lives) \cite{ramakrishnan1999proposal}.\\
+In 2001 a new method was standardized to explicitly notify TCP on congestion.
+\textit{Explicit Congestion Notification} (ECN) is an extension to TCP and IP, allowing end-to-end congestion notification, while also minimizing the amount of packets dropped.
+This is done by modifying the IP header to be able to carry one more bit - the \textit{Congestion Experienced} flag (CE) \cite{ramakrishnan1999proposal}.
+Instead of dropping packets occasionally after the amount of queued packets exceed a threshold [\nameref{P_C_marker}], all packets supporting ECN will now be marked (setting the CE bit in the IP header).
+Thus, congestion can be described explicitly resulting in better short- and long-term network performance.
+Also, if a data packet has been market, the according acknowledgment package will also be marked \cite{ramakrishnan1999proposal,ramakrishnan2001addition,10.1145/205511.205512}.\\
+To sum up so far, without ECN, classic TCP would just drive the network into congestion, and then recover from it.
+The network generally speaking will at least experience some jitter (fluctuating network delay).
+With ECN however, packets in congested nodes will be marked, explicitly informing the TCP sender that the network is congested.
+As a result, avoiding actions can be taken immediately.
+Therefore, we also speak of \textit{Congestion Control} \cite{ramakrishnan1999proposal,ramakrishnan2001addition}.
+
+\subsection{XCP}
+Although TCP is a very dominant protocol in the internet, there are also other transport protocols which could benefit from end-to-end congestion notification.
+The paper "Congestion Control for High Bandwidth-Delay Product Networks"  generalizes the ECN proposal of 2001, creating a new protocol called \textit{eXplicit Congestion Protocol} (XCP), while also improving upon it further by introducing new packet control concepts \cite{katabi2002congestion,1498331}.
+
+\subsubsection*{XCP Header}
+Like TCP, XCP is a window-based congestion control protocol indented for best traffic.
+Underneath the hood however, they are very different.
+A major difference between both protocols is that XCP does not only inform the sender on congestion, but does also provide information about the degree of congestion.
+XCP can achieve this by including a \textit{congestion header} to each packet.
+While the sender maintains a \textit{cwnd} and \textit{RTT} variable, as with TCP, those are also being communicated to network nodes via the congestion header as \textit{H\_cwnd} and \textit{H\_rtt}.
+In addition, the congestion header includes one more field.
+This \textit{H\_feedback} field is also initialized by the sender, but is the only field which can be modified by nodes along the path \cite{katabi2002congestion,1498331}.
+
+\subsubsection*{XCP Sender}
+As with TCP, the XCP sender maintains a congestion window, \textit{cwnd}, and an estimate of the round trip time, \textit{RTT}.
+On packet departure, the \textit{H\_cwnd} and \textit{H\_rtt} fields will be initialized with \textit{cwnd} and \textit{RTT} accordingly.
+The initialization value of the \textit{H\_feedback} field represents a desired send rate $r$:
+\[H\_feedback \leftarrow \frac{r \cdot RTT - cwnd}{cwnd}\]
+This is an improvement to TCPs adjustment algorithms as the desired rate can be reached after one RTT.\\
+The other task an XCP sender has is to adjust the congestion window according to the networks feedback.
+When s is the packet size, the congestion window variable would be adjusted by following formula:
+\[cwnd \leftarrow \max(cwnd + H_feedback, s)\]
+In addition to responding to direct feedback, the XCP sender still needs to handle package loss, though this is very similar to what TCP does \cite{katabi2002congestion}.
+
+\subsubsection*{XCP Receiver}
+The XCP receiver functions exactly the same way like a TCP receiver would except for copying the (modified) congestion header from the data packet to its acknowledgment \cite{katabi2002congestion}.
+
+\subsubsection*{XCP Router}
+The main task of network nodes in XCP, or XCP routers, is to inform the XCP sender of the current state of congestion.
+This is done by computing feedback at every node.
+The objective of XCP is to prevent the queues building up to a point at which packets have to be dropped \cite{katabi2002congestion}.\\
+Feedback is computed by two controllers.
+The \textit{Efficiency Controller} (EC) tries to maximize link utilization while minimizing queue lengths and packet drop rates.
+When the XCP routers link is underused, a positive feedback should be sent.
+If the link is congested however, a negative feedback should be sent in order to keep queue lengths small resulting in no packet drops \cite{katabi2002congestion,1498331}.\\
+To achieve efficiency and equilibrium, this feedback will be allocated to single packets as \textit{H\_feedback}.
+However, the EC does not decide which packets should carry the feedback.
+Hence, the \textit{Fairness Controller} (FC) comes into play.
+As TCP, FC relies on the AIMD principle [\nameref{AfC}].
+If the EC calculated a positive feedback, it will be split up between all packages equally.
+If the feedback is negative however, it will be divided between packets in proportion to their current send-rates, which are determined by \textit{H\_cwnd} and \textit{H\_rtt} \cite{katabi2002congestion,1498331}.\\
+Decoupling the feedback computation allows XCP to quickly acquire the maximum send rate and achieve equilibrium, all while the network allocates bandwidth to all XCP senders equally \cite{katabi2002congestion}.
+
+\subsection{Evaluation of Network Assisted Congestion Control}
+As you can see, including the network layer to detect can have great performance improvements.
+'Understanding XCP: Equilibrium and Fairness' has shown through simulations that many network topologies could benefit from XCP as it clears queues in equilibrium, thus resulting in improved network performance under load \cite{1498331}.
+Another paper has shown that XCP outperforms ordinary TCP in nearly any environment related to average queue lengths, efficiency and throughput \cite{katabi2002congestion}.
+\\However, all those performance comes with a cost.
+Since network nodes have to actively detect congestion, ECN, XCP and other network assisted congestion control protocols rely on active queue management in those nodes.
+Although most hardware being produced today meet those requirements, a large part of internet hardware isn't capable of doing so.
+Because any of those nodes can be congested, it would be counterintuitive to run those protocols, as congestion might or might not be detected, thus resulting in performance loss.
+\\Nevertheless, whenever implementation of XCP, ECN or similar protocols is not a challenge, you can implement network assisted congestion control with ease and see great performance improvements.
+Especially smaller networks with high traffic could definitely benefit from those technologies.
+
+\subsection{DCTCP}
+A lot of internet services run in data centers with large amounts computers communicating with each other, resulting in a lots of traffic inside the data center.
+Optimizing those data centers networking capabilities can result in an increase of quality of services, which they are hosting.
+PAPER OF \textit{DCTCP} proposes a variation of TCP using ECN to create a data transportation protocol, which specially takes hardware purposed for use those environments into account.
 

### 2020-05-02 paper/paper.tex
Almost there

@@
 \usepackage{tikz}							% tikz-figures
-\usetikzlibrary{positioning,arrows,automata}
+\usetikzlibrary{decorations.pathreplacing,positioning,arrows,automata}
 \tikzstyle{line} = [draw,thick,-latex]
@@
 	\centering
-		\begin{tikzpicture}[->, >=latex, node distance = 2.5cm, semithick]
+		\begin{tikzpicture}[-, >=latex, node distance = 2.5cm, semithick]
 			%network
@@
 Although most hardware being produced today meet those requirements, a large part of internet hardware isn't capable of doing so.
-Because any of those nodes can be congested, it would be counterintuitive to run those protocols, as congestion might or might not be detected, thus resulting in performance loss.
+Because any of those nodes can be congested, it would be counterintuitive to run those protocols, as congestion might or might not be detected, thus resulting in performance loss \cite{katabi2002congestion}.
 \\Nevertheless, whenever implementation of XCP, ECN or similar protocols is not a challenge, you can implement network assisted congestion control with ease and see great performance improvements.
@@
 A lot of internet services run in data centers with large amounts computers communicating with each other, resulting in a lots of traffic inside the data center.
-Optimizing those data centers networking capabilities can result in an increase of quality of services, which they are hosting.
-PAPER OF \textit{DCTCP} proposes a variation of TCP using ECN to create a data transportation protocol, which specially takes hardware purposed for use those environments into account.
+Optimizing those data centers networking capabilities can result in an increase of quality of hosted services.
+In 2010 a paper proposed \textit{DCTCP}, a variation of TCP using ECN to create a data transportation protocol, which specially takes hardware purposed for use those environments into account \cite{10.1145/1851275.1851192}.\\
+The general problem that appears in high-traffic environments is the so called \textit{TCP Incast} problem.
+
+\begin{figure}
+	\centering
+		\begin{tikzpicture}[-, >=latex, node distance = 2.5cm, semithick, auto]
+			\node[router, scale=0.8] (Router) {};
+			\node (Internet) [right = 1.5cm of Router, text width=2cm] {\scriptsize Outgoing Connection};
+			\node[switch, scale=0.8] (Sw1) [below left = 0.8cm and 1cm of Router] {};
+			\node[switch, scale=0.8] (Sw2) [below right = 0.8cm and 1cm of Router] {};
+			\node[server, scale=1] (S1) [below left = 0.8cm and 0.25cm of Sw1] {\scriptsize Server};
+			\node[server, scale=1] (S2) [below = 2.1cm of Router] {\scriptsize Server};
+			\node[server, scale=1] (S3) [below right = 0.8cm and 0.25cm of Sw2] {\scriptsize Server};
+
+			\draw[-] (Router) -- (Sw1);
+			\draw[-] (Router) -- (Sw2);
+			\draw[-, dotted] (Router) -- (Internet);
+			\draw[-] (Sw1) -- (S1);
+			\draw[-] (Sw1) -- (S2);
+			\draw[-] (Sw1) -- (S3);
+			\draw[-] (Sw2) -- (S1);
+			\draw[-] (Sw2) -- (S2);
+			\draw[-] (Sw2) -- (S3);
+
+			\draw[decorate,decoration={brace,amplitude=10pt},xshift=-4pt,yshift=0pt] (S1.north -| S1.west) -- (Router.center -| S1.west) node [black,midway,xshift=-0.6cm,align=right] {\scriptsize High\\ \scriptsize'Fan-In'};
+		\end{tikzpicture}
+	\hspace{0.5cm}
+	\caption{Data Center High 'Fan-In'}
+	\label{fig:dc}
+\end{figure}
 

### 2020-05-16 paper/paper.tex
Feedback

@@
 \begin{abstract}
-This article presents the most common origins of network congestion, how different types of algorithms perceive and handle congestion and what obstacles they face.
+In this article we present the most common origins of network congestion.
+To tackle congestion, different types of algorithms will be compared, by first describing how each perceives congestion and later on how they adjust to those circumstances.
+This paper will also take a closer look at network assisted congestion control and a real-use implementation in data centers.
 \end{abstract}
@@
 \section{Introduction}
-Congestion was, is and always will be a problem on the Internet, appearing in various forms with different symptoms causing minor, but also major challenges.
-Data sent over the internet is split up in smaller packets that all try to reach their destination the fastest, whether it is by going the shortest or the least congested route.
-In an optimal network all computers, servers etc.
+Congestion was, is, and always will be a problem on the Internet, appearing in various forms with different symptoms causing minor, but also major challenges.
+Most data sent over the internet is split up in smaller packets that all try to reach their destination the fastest, whether it is by going the shortest or the least congested route.
+In an optimal network all computers, servers, etc.
 would have a direct link to each other, thus reducing the importance of bandwidth and  consequently removing the need for buffers at all.
-However, the larger the network would get, the amount of connections would grow quadratic.
-\\Networks today consist of network nodes (routers, switches, ...) which all try to route internet packets as quick as possible whilst keeping the required hardware to a minimum.
+However, the larger the network grows, the number of links would increase quadratically.
+\\Networks today consist of network nodes (routers, switches, ...) which all try to route internet packets as quickly as possible whilst keeping the required hardware to a minimum.
 Ideally there would be a steady flow of packets being transmitted at maximum rate.
 However, when a internet node receives a packet while the outgoing connection still is in use, the packet has to be stored in a buffer until the outgoing connection freed up again.
-Therefore, those storage units are essential to every node on a network, but without proper management even well-sized buffers could lead to network problems like packet-loss, jitter and connection delay \cite{appenzeller2004sizing}	.
+Therefore, those storage units are essential to every node on a network, but without proper management even well-sized buffers could lead to network problems like packet-loss, jitter, and connection delay \cite{appenzeller2004sizing}.
 \\This paper will present origins of network congestion and analyze how different types of algorithms try to avoid and minimize congestion.
@@
 \section{Origins of Congestion}
-With the ever-growing amount of internet services not only data-centers, but the whole internet in general face new problems with each packet sent.
-Therefore, the global reduction of those complications are an important task to improve the quality of all internet services.
-However, as it is for solving any problem, the origin of those problems must be identified first.
+With the ever-growing amount of internet services not only data-centers, but the whole internet in general faces new challenges with each packet sent.
+Therefore, a network-wide reduction of those complications is an important task to improve the quality of all internet services.
 \\Network delay can generally speaking be categorized into three different types of delay \cite{gettys2012bufferbloat}:
@@
 \textit{Transmission delay} describes the delay of transmitting data packets between two internet nodes.
-Typically, packets sent from a origin to a destination node are sent over various different internet nodes, all of which could have a different bandwidth.
-Therefore, a packet send from \textbf{node 1} to \textbf{node n} over \textbf{nodes 2 to n-1} cannot be sent at a faster rate than the lowest bandwidth on the packets path (bottleneck) [Figure \ref{fig:transmission delay}].\[Rate_{max} \leftarrow \min\{bandwidth_{i}\mid i\in\{1,...,n-1\}\}\]
+Typically, packets sent from an origin to a destination node are sent over various different internet nodes, all of which could have a different bandwidth.
+Therefore, a packet send from \textbf{node 1} to \textbf{node n} over \textbf{nodes 2 to n-1} cannot be sent at a faster rate than the lowest bandwidth on the path of the packet (bottleneck) [Figure \ref{fig:transmission delay}].\[Rate_{max} \leftarrow \min\{bandwidth_{i}\mid i\in\{1,...,n-1\}\}\]
 
@@
 \item
-\textit{Queuing delay} is the time a package spent waiting for it to be processed or transmitted.
+\textit{Queuing delay} is the time a package spends waiting to be processed or transmitted.
 \end{itemize}
@@
 	\end{center}
-	\caption{Path a packet takes through the network}
+	\caption{Path a packet takes through the network. Each circle represents a network node with $1$ being the sender and $n$ being the receiver.}
 	\label{fig:transmission delay}
@@
 There are many different ways to implement buffers in a network node and the effectiveness of each implementation varies from where it is used.
-Traffic in residential network nodes for example differ from nodes used in large data-centers in every aspect, thus hardware and software in those components should be adjusted accordingly.
+Traffic in residential network nodes for example differs from nodes used in large data-centers in every aspect, thus hardware and software in those components should be adjusted accordingly.
 \\Sizing buffers is a complex task.
-On the one hand read and write times should as short as possible to minimize additional delay.
-On the other hand buffers should be as big as needed to be robust for future network changes and therefore store lots of data if needed while also keeping in mind the requirements congestion detection algorithms have in order to adjust for congestion accordingly \cite{staff2012bufferbloat}.
+On the one hand read and write times should be as short as possible to minimize additional delay.
+On the other hand buffers should not be too small, as networks need to be robust for future changes and therefore nodes should store lots of data if needed, while also keeping in mind the requirements of congestion detection algorithms as described in section \nameref{sec:ca} in order to adjust for congestion appropriately \cite{staff2012bufferbloat}.
 Unifying those goals is very challenging as it leads to increasing costs and physical space occupied by those buffers.
-Additionally, a 2004 paper by Appenzeller found that 99\% of buffers in network nodes, which were sized on a rule-of-thumb rule based on the TCP protocol \cite{villamizar1994high}, could be removed without losing, but rather gaining performance \cite{appenzeller2004sizing}.
+Additionally, a 2004 paper by Appenzeller et al.
+\cite{appenzeller2004sizing} found that 99\% of buffers in network nodes, which were sized on a rule-of-thumb rule based on the TCP protocol \cite{villamizar1994high}, could be removed without losing, but rather gaining performance.
 Clearly, balancing speed and size is difficult and thus buffer sizes should be adjusted to their area of use accordingly.
 \\This isn't the only major cause of queuing delay.
-Without proper buffer management even nodes with well sized buffers will experience performance loss.
+Without proper buffer management even nodes with well-sized buffers will experience performance loss.
 Proper queue management is able to compensate for the unpredictable fluidity of network traffic.
-It is an difficult task to buffer the right amount of packages to keep transmission flow consistent over time.
+It is a difficult task to buffer the right amount of packages to keep transmission flow consistent over time.
 As the number of packets being transmitted grows, the \textit{throughput} of the network node increases until packets are being transmitted at bottleneck rate.
-After this, the transmission rate cannot increase anymore, which leads to packets being buffered along their path to prevent them being lost \cite{gettys2012bufferbloat}.
+After this, the transmission rate cannot increase anymore, which leads to packets being buffered along their path to prevent them from being lost \cite{gettys2012bufferbloat}.
 If the incoming amount of traffic keeps to exceed the amount of outgoing traffic, those buffers will fill up (\textit{over-buffering}), which will eventually lead to \textit{bufferbloat} \cite{Allman13commentson,6886125}.
@@
 Bufferbloat refers to frequently full and excessively large buffers in internet nodes.
-This leads to partly unnecessary network delay and damages congestion avoidance algorithms, such as classic TCP \cite{gettys2012bufferbloat,chen2014bufferbloat}.
+This leads to partly unnecessary network delay and damages congestion avoidance algorithms, such as those found in classic TCP \cite{gettys2012bufferbloat,chen2014bufferbloat}.
 Buffering is needed to provide space to queued packets waiting for transmission resulting in a minimization of packet loss.
-In the past however, buffer sizes were smaller than today's thus less packets could be queued and more were dropped, signaling transmission protocols the presence of congestion in this node, so those protocols can adjust.
+In the past however, buffer sizes were smaller than today's thus less packets could be queued and more were dropped, signaling transmission protocols the presence of congestion in this node early in.
 With memory getting cheaper over the years, buffers saw an increase in size.
 This has the negative impact, that more packets can be queued before any are dropped, thus those adjustments of congestion avoidance protocols kick in later than usual.
-Thus, buffers are being filled in the whole network which results in an extreme increase in network delay (\textit{latency}) and fluctuation (\textit{jitter}) \cite{gettys2012bufferbloat,staff2012bufferbloat,chen2014bufferbloat}.
-\\So Services, which require low latency at a high data-rate, such as Streaming, will experience a decrease in quality.
+Therefore, buffers are being filled in the whole network which results in an extreme increase in network delay (\textit{latency}) and fluctuation (\textit{jitter}) \cite{gettys2012bufferbloat,staff2012bufferbloat,chen2014bufferbloat}.
+\\So Services, which require low latency at a high data-rate, such as streaming, will experience a decrease in quality.
 
 \section{Congestion Avoidance}
+\label{sec:ca}
 As already mentioned, congestion detection algorithms make up an important part in congestion avoidance protocols.
-\textit{TCP} is such an protocol, based on the transport layer, which adapts its send rate according to available network bandwidth \cite{1209197,jacobson1992tcp}.
-To fully understand, how TCP's congestion avoidance algorithm functions, you will have to understand how TCP works.
+\textit{TCP} is such a protocol, based on the transport layer, which adapts its send rate according to available network bandwidth \cite{1209197,jacobson1992tcp}.
+To fully understand, how TCP's congestion avoidance algorithm operates, one needs to understand how TCP works.
 TCP lives on the transport layer.
-It communicates with the application layer via ports, splits up data into packets (and vice versa), and transmits those packages over the network with the use of the internet protocol, thus TCP is often also referred to as \textit{TCP/IP}.
-It is the interface between network components.
-Each of those communication packages have different purposes, but underlay the same structure - the protocol header.
-The TCP header is split up into multiple segments [Figure \ref{fig:TCP_Header}]\cite{jacobson1992tcp,huston2000tcp}:
+It communicates with the application layer via sockets, splits up data into packets (and vice versa), and transmits those packages over the network with the help of the internet protocol, thus TCP is often also referred to as \textit{TCP/IP}.
+It is the interface between network components and computer programs.
+Each of those packets have different purposes, but underlay the same structure - the protocol header.
+The TCP header is split up into multiple segments [Figure \ref{fig:TCP_Header}]:
 \begin{figure}
 \centering
-\begin{bytefield}[bitheight=2.2\baselineskip]{32}
+\begin{bytefield}[bitheight=2.2\baselineskip ,bitwidth=0.635\baselineskip]{32}
 	\bitbox{16}{Source Port} & \bitbox{16}{Destination Port} \\
@@
 	\bitbox{32}{Acknowledgment Number} \\
-	\bitbox{4}{Data Offset} & \bitbox{6}{Reserved} &
+	\bitbox{4}{Data Offset} & \bitbox{4}{Rsrvd} &
+\bitbox{1}{\tiny C\\W\\R} & \bitbox{1}{\tiny E\\C\\E} &
 \bitbox{1}{\tiny U\\R\\G} & \bitbox{1}{\tiny A\\C\\K} &
@@
 \end{bytefield}
-\caption{TCP Header}
+\caption{Basic structure of the TCP header.
+It is split up into multiple 32-bit blocks.
+The data block contains all the application data to be transmitted \cite{ietf-tcpm-rfc793bis-16,jacobson1992tcp,huston2000tcp}.}
 \label{fig:TCP_Header}
@@
 This way packet loss can be detected.
+As the sequence numbers are limited to 32 bit, they cannot increase endlessly.
+Thus, if all possible sequence numbers are used up, it wraps around to 0 to keep the connection up and running \cite{postel1981transmission}.
 
-\item TCP can set six different bit-flags in its header.
+\item TCP can set eight different bit-flags in its header.
 The most important one is the \textit{ACK} flag (Acknowledgement bit).
-It is used whenever a packet receives the destination and an acknowledgement packet has to be sent back.
+It is used whenever a packet arrives at its destination and an acknowledgement packet has to be sent back.
 The \textit{SYN} and \textit{FIN} flags initialize and terminate a connection.
@@
 \item The \textit{Window Size} is the maximum amount of data (in bytes) a receiver is willing to receive at any time.
-It is being negotiated at the initialization of a new connection and can only be negotiated again with the \textit{SYN} flag set.
+It is being negotiated at the initialization of a new connection and can only be negotiated again wile the \textit{SYN} flag is set.
 It is important to set it appropriately to prevent unnecessary packet loss or delay, as the receiver couldn't process the data in time \cite{jacobson1992tcp}.
@@
 Whenever a client sends a data packet to the server, the server waits until it has received enough data (specified by the window size), and then sends back an acknowledgment package, signaling the successful transmission of data.
-There are also other handshake operations used, for example for establishing or terminating a connection, however those have no further importance to the congestion avoidance protocol \cite{1209197,jacobson1992tcp,huston2000tcp,jacobson1995congestion}.
+There are also other handshake operations used, for example for establishing or terminating a connection, however, those have no further importance to the congestion avoidance protocol \cite{1209197,jacobson1992tcp,huston2000tcp,jacobson1995congestion}.
 
@@
 \label{P_C_marker}
-Both partners of a connection are either a sender or a receiver.
+Both partners of a connection are both sender and receiver, as both transmit packets.
+In the following however, we will simplify this and specify one partner as the sender, the other as the receiver.
 The sender keeps track of the time it takes from when a package sent until the acknowledgment is received.
-This is also known as \textit{round-trip time} (RTT), where a large RTT can indicate some congestion along the path.
+This is also known as \textit{round-trip time} (RTT).
+Comparing the current RTT to previously measured RTTs can inform the sender about the current state of congestion.
 Some variances of TCP make use of this feature in their congestion avoidance algorithms, yet classic TCP does not \cite{huston2000tcp,jacobson1995congestion}.
-RTT also plays an very important role in the detection of packet loss, which further imply congestion.
+RTT also plays a very important role in the detection of packet loss, which further implies congestion.
 The sender side keeps track of the RTT while it is being measured and compares it to the \textit{retransmission timeout} (RTO) \cite{jacobson1992tcp}.
 RTO time is being calculated by estimating the mean and variance of the RTT.
-Whenever packets are not acknowledged within this RTO intervall, the package is being perceived as lost and will be retransmitted.
+Whenever packets are not acknowledged within this RTO interval, the TCP sender regards this packet as lost and will retransmit its data.
 Therefore, it is essential to TCP performance, that the RTO is being determined accurately, as a too low RTO could prompt unnecessary retransmissions and too high RTO could slow down the application TCP is serving \cite{huston2000tcp,jacobson1995congestion}.
 To implicitly notify TCP that congestion occurs, congested network nodes will occasionally drop packets after a certain threshold of queued packets has been exceeded.
-Usually, this is quite smaller the maximum queue length, therefore avoiding actions can be taken in time \cite{ramakrishnan1999proposal}.\\
+Usually, this is quite smaller than the maximum queue length, therefore avoiding actions can be taken in time \cite{ramakrishnan1999proposal}.\\
 Note however, a high RTT can also have other causes than packet loss or congestion.
@@
 Here, a variable called \textit{congestion window} (cwnd) has an important role.
-It describes the maximum amount of packages send during a interval.
-An additional mechanism of keeping a connection in equilibrium is, that each acknowledgement package received by the sender opens the space for a new package to be sent into the network \cite{jacobson1992tcp}.
+It describes the maximum amount of packets sent during an average RTT.
+An additional mechanism of keeping a connection in equilibrium is, that each acknowledgment packet received by the sender opens the space for a new packet to be sent into the network \cite{jacobson1992tcp}.
 The principles TCP uses to adjust for congestion are based on \textit{Additive-Increase Multiplicative-Decrease} (AIMD) principle.
@@
 The \textit{slow-start} algorithm is used when a connection has just been established or when multiple packet losses were detected.
-The cwnd variable is set to one package per estimated RTT, because no information about network congestion has been collected yet.
-Whenever an acknowledgment package is being received, cwnd will increase by one package.
-This exponential growth in packages helps to get the packet send rate to the maximum the network can handle quite quickly, resulting in almost immediately maximum network performance \cite{jacobson1992tcp}.
+The cwnd variable is set to one packet per estimated RTT, because no information about network congestion has been collected yet.
+Whenever an acknowledgment packet is being received, cwnd will increase by one packet.
+This doubles cwnd each RTT, resulting in an exponential growth over time, which helps to get the packet send rate to the maximum the network can handle quite quickly, contributing to almost immediately maximum network performance \cite{jacobson1992tcp}.
 
@@
 Therefore, the network can only stabilize if the amount of packets sent decreases at least as quickly as the queues are growing.
-The formula \cite{jacobson1992tcp} \[cwnd_{i+1} \leftarrow d*cwnd_{i}, 0<d<1\] has an exponential decrease over time, thus resulting in the network recovering.\\
+The formula \cite{jacobson1992tcp} \[cwnd_{i+1} \leftarrow d \cdot cwnd_{i}, 0<d<1\] has an exponential decrease over time, thus resulting in the network recovering.\\
 If the network isn't congested, the sender should try to raise cwnd slowly to reach the networks maximum throughput again.
 A multiplicative, exponential increase over time, like the slow-start algorithm isn't the best choice here.
-Increasing the amounts of packets sent drastically while cwnd still is high can result in the network being challenge the networks capacity, resulting in heavy congestion.
+Increasing the amounts of packets sent drastically while cwnd still is high can challenge the networks capacity, resulting in heavy congestion.
 Therefore, the best option to raise cwnd to its limits is an additive increase \cite{jacobson1992tcp}:\[cwnd_{i+1} \leftarrow cwnd_{i} + u\] Van Jacobson proposes in his 1992 paper about TCP congestion avoidance that $d$ will be set to $\frac{1}{2}$ and $u$ will be set to $\frac{1}{cwnd_{i}}$.
@@
 		\end{tikzpicture}
-	\caption{TCP/IP with Explicit Congestion Notification}
+	\caption{Congestion Control Algorithm. A congested node marks buffered packets to describe congestion. This signal will be sent back to the original sender.}
 	\label{fig:ecn}
@@
 Here, the network layer comes into play.
-Because most communication should be hidden along the packets path, all network nodes will only communicate on the internet layer (where IP lives), but not above (where TCP lives) \cite{ramakrishnan1999proposal}.\\
+Because most communication should be hidden along the path of the packet, all network nodes will only communicate on the network layer (where IP lives), but not above (where TCP lives) \cite{ramakrishnan1999proposal}.\\
 In 2001 a new method was standardized to explicitly notify TCP on congestion.
@@
 This is done by modifying the IP header to be able to carry one more bit - the \textit{Congestion Experienced} flag (CE) \cite{ramakrishnan1999proposal}.
-Instead of dropping packets occasionally after the amount of queued packets exceed a threshold [\nameref{P_C_marker}], all packets supporting ECN will now be marked (setting the CE bit in the IP header).
+Instead of dropping packets occasionally after the amount of queued packets exceed a threshold [\nameref{P_C_marker}], all packets supporting ECN will now be marked (setting the CE bit in the IP header) [Figure \ref{fig:ecn}].
 Thus, congestion can be described explicitly resulting in better short- and long-term network performance.
-Also, if a data packet has been market, the according acknowledgment package will also be marked \cite{ramakrishnan1999proposal,ramakrishnan2001addition,10.1145/205511.205512}.\\
+Finally, if a data packet has been marked, the according acknowledgment packet will also be marked to inform the sender of congestion \cite{ramakrishnan1999proposal,ramakrishnan2001addition,10.1145/205511.205512}.\\
 To sum up so far, without ECN, classic TCP would just drive the network into congestion, and then recover from it.
@@
 As a result, avoiding actions can be taken immediately.
-Therefore, we also speak of \textit{Congestion Control} \cite{ramakrishnan1999proposal,ramakrishnan2001addition}.
+In addition, the need to retransmit packets would decrease drastically, as no packets are being dropped anymore to indicate congestion.
+Therefore, this central algorithm is also being called \textit{Congestion Control} \cite{ramakrishnan1999proposal,ramakrishnan2001addition}.
 
 \subsection{XCP}
-Although TCP is a very dominant protocol in the internet, there are also other transport protocols which could benefit from end-to-end congestion notification.
+Although TCP is a very dominant protocol on the internet, there are also other transport protocols that could benefit from end-to-end congestion notification.
 The paper "Congestion Control for High Bandwidth-Delay Product Networks"  generalizes the ECN proposal of 2001, creating a new protocol called \textit{eXplicit Congestion Protocol} (XCP), while also improving upon it further by introducing new packet control concepts \cite{katabi2002congestion,1498331}.
@@
 \subsubsection*{XCP Header}
-Like TCP, XCP is a window-based congestion control protocol indented for best traffic.
+Like TCP, XCP is a window-based congestion control protocol intended for the best effort.
 Underneath the hood however, they are very different.
@@
 The other task an XCP sender has is to adjust the congestion window according to the networks feedback.
-When s is the packet size, the congestion window variable would be adjusted by following formula:
+When $s$ is the packet size, the congestion window variable would be adjusted by following formula:
 \[cwnd \leftarrow \max(cwnd + H_feedback, s)\]
-In addition to responding to direct feedback, the XCP sender still needs to handle package loss, though this is very similar to what TCP does \cite{katabi2002congestion}.
+In addition to responding to direct feedback, the XCP sender still needs to handle packet loss, though this is very similar to what TCP does \cite{katabi2002congestion}.
 
@@
 The \textit{Efficiency Controller} (EC) tries to maximize link utilization while minimizing queue lengths and packet drop rates.
-When the XCP routers link is underused, a positive feedback should be sent.
-If the link is congested however, a negative feedback should be sent in order to keep queue lengths small resulting in no packet drops \cite{katabi2002congestion,1498331}.\\
+When the XCP routers link is underused, positive feedback should be sent.
+If the link is congested however, negative feedback should be sent in order to keep queue lengths small resulting in no packet drops \cite{katabi2002congestion,1498331}.\\
 To achieve efficiency and equilibrium, this feedback will be allocated to single packets as \textit{H\_feedback}.
@@
 As TCP, FC relies on the AIMD principle [\nameref{AfC}].
-If the EC calculated a positive feedback, it will be split up between all packages equally.
+If the EC calculated positive feedback, it will be split up between all packets equally.
 If the feedback is negative however, it will be divided between packets in proportion to their current send-rates, which are determined by \textit{H\_cwnd} and \textit{H\_rtt} \cite{katabi2002congestion,1498331}.\\
@@
 \subsection{Evaluation of Network Assisted Congestion Control}
-As you can see, including the network layer to detect can have great performance improvements.
+As you can see, including the network layer to detect congestion can have great performance improvements.
 'Understanding XCP: Equilibrium and Fairness' has shown through simulations that many network topologies could benefit from XCP as it clears queues in equilibrium, thus resulting in improved network performance under load \cite{1498331}.
-Another paper has shown that XCP outperforms ordinary TCP in nearly any environment related to average queue lengths, efficiency and throughput \cite{katabi2002congestion}.
-\\However, all those performance comes with a cost.
-Since network nodes have to actively detect congestion, ECN, XCP and other network assisted congestion control protocols rely on active queue management in those nodes.
-Although most hardware being produced today meet those requirements, a large part of internet hardware isn't capable of doing so.
+Another paper has shown that XCP outperforms ordinary TCP in nearly any environment related to average queue lengths, efficiency, and throughput \cite{katabi2002congestion}.
+\\However, all this performance comes at a cost.
+Since network nodes have to actively detect congestion, ECN, XCP, and other network assisted congestion control protocols rely on active queue management in those nodes.
+Although there is hardware being produced today that meet those requirements, a large part of internet hardware isn't capable of doing so.
 Because any of those nodes can be congested, it would be counterintuitive to run those protocols, as congestion might or might not be detected, thus resulting in performance loss \cite{katabi2002congestion}.
-\\Nevertheless, whenever implementation of XCP, ECN or similar protocols is not a challenge, you can implement network assisted congestion control with ease and see great performance improvements.
+This is the reason you do not see those protocols in wide area networks (WAN) too often, let alone the internet.
+\\Nevertheless, whenever the implementation of XCP, ECN, or similar protocols is not a challenge, you can implement network assisted congestion control with ease and see great performance improvements.
 Especially smaller networks with high traffic could definitely benefit from those technologies.
@@
 \subsection{DCTCP}
-A lot of internet services run in data centers with large amounts computers communicating with each other, resulting in a lots of traffic inside the data center.
-Optimizing those data centers networking capabilities can result in an increase of quality of hosted services.
-In 2010 a paper proposed \textit{DCTCP}, a variation of TCP using ECN to create a data transportation protocol, which specially takes hardware purposed for use those environments into account \cite{10.1145/1851275.1851192}.\\
-The general problem that appears in high-traffic environments is the so called \textit{TCP Incast} problem.
+A lot of internet services run in data centers with large amounts of computers communicating with each other, resulting in a lot of traffic inside the data center.
+Optimizing those data centers networking capabilities can result in an increase in application productivity.
+In 2010 a paper proposed \textit{DCTCP}, a variation of TCP using ECN to create a data transportation protocol, which especially takes hardware purposed for use in those environments into account \cite{10.1145/1851275.1851192}.\\
+The general problem that appears in high-traffic environments is the so-called \textit{TCP Incast} problem.
+It describes a drastic reduction in application throughput.
+To understand where TCP incast originates, you fist need to understand how most data centers are built.
+A typical data center contains a set of racks, each holding a large number of servers, the switches, that connect those racks, and connecting links, that connect those switches to other parts of the data center or the internet [Figure \ref{fig:dc}].
+This design pattern is also known as \textit{Partition/Aggregate} and is most commonly used in data centers.
+It however results in a high 'fan-in', as multiple servers (leaves) are connected to just one link (root) \cite{10.1145/1851275.1851192,10.1145/1592681.1592693}.
+In addition, most applications require high bandwidth, low latency connections, also enabling them to send many parallel requests to other servers.
+Combining those requirements with the design pattern, it becomes clear, that those switches are being challenged.
+Besides, many data center switches have relatively small, shared buffers.
+Therefore, it is common that queue lengths stay high for a prolonged period of time, resulting in a collapse of throughput \cite{10.1145/1851275.1851192,10.1145/1592681.1592693}.\\
+The goal of DCTCP is to keep data center traffic constantly flowing, which means meeting the requirements applications take while keeping relatively small buffer sizes inside switches in mind.
+To achieve this, DCTCP implements ECN.
+Unlike TCP however, if a DCTCP sender registers congestion, the window size will be reduced by a fraction of the amount of marked packets it received.
+Therefore, excessive link underutilization can be prevented.
+In addition, the DCTCP receiver will only send one ACK packet for every $m$ consecutive data packet, where all $1,...,m$ packets before had the same CE-flag configuration.
+Here, $m$ is a variable initialized while the DCTCP connection is being established and cannot be changed without renegotiation.
+Therefore, fewer packets are sent through the network, but the TCP sender is still able to identify lost packets and register congestion \cite{10.1145/1851275.1851192}.
+\\Being able to react to congestion immediately enables DCTCP to reduce buffer overflows and timeouts, thus largely resolving incast problems.
+The paper proposing DCTCP has found, that their protocol increases traffic by 110\% while using 90\% less buffer space compared to TCP \cite{10.1145/1851275.1851192}.
 
@@
 \section{Conclusion}
+In this paper we have analyzed origins of congestion.
+Different factors play a role in network congestion, however especially the filling up of buffers at congested nodes will cause traffic delay.
+Further, we shine a light on the functionality of congestion avoidance algorithms, especially classic TCP.
+However, because those algorithms quite often do cause network stress rather than resolving it, we took a closer look at network assisted end-to-end congestion control.
+ECN, an addition to TCP, and XCP are two standardized protocols applying those principles, but haven't been very successful in wide area networks (WAN).
+In smaller networks, the implementation of them is easier.
+DCTCP, a variation of ECN, as an example, can successfully resolve many issues occurring in data centers.
 

### 2020-05-16 paper/Übungsblatt 2/Übungsblatt.tex
Feedback

\documentclass[11pt]{article}

%Packages
\usepackage{amsfonts}	      %Mathematische Zeichen und Fonts
\usepackage{mathtools}        %Extra Mathematische Symbole
\usepackage{extarrows}	      %Extra Pfeile
\usepackage{listings}         %Codeansicht
\usepackage{scrlayer-scrpage} %Seitenkopf
\usepackage{tikz}             %tikz
\usepackage{enumitem}		  %Enumerate
\usepackage{listings}		  %Code snippets
\usepackage{amsmath}

\usetikzlibrary{arrows, automata, positioning}
\pagestyle{scrheadings}

\begin{document}

%Header
\ihead{\textbf{Betriebssysteme und Systemsoftware \\ Übungsblatt X} \\Tutorium 1}
\ohead{Andrés Montoya, 405409 \\ Til Mohr, 405959\\ Marc Ludevid Wulf, 405401}

%Seiteninhalt
\paragraph{Aufgabe 2.2}
\begin{enumerate}[label = \alph*)]
\item \texttt{echo "BuS 1020: Abgabe der 2. Uebung am 11.5."| sed 's/1/2/'}
%test
\item Siehe $script.sh$.

\item Dieses Skript sucht alle C-Programme in diesem Directory und allen Subdirectories und addiert die Anzahl der Zeilen zusammen.

\item Siehe $verzeichnis\_struktur.sh$.

\end{enumerate}

\paragraph{Aufgabe 2.3}
\begin{enumerate}[label = \alph*)]
\item Systemcalls sind die Schnittstelle zwischen dem User-Mode und dem Kernel-Mode. Durch Systemcalls können die Anwendungsprogramme und Standardbibliiotheken auf Kernel-Funktionalität zugreifen.

\item Accept entfernt das älteste Element von der Warteschlange an anstehenden Verbindungen die an einem Socket anliegen um diese Verbindung zu bearbeiten. Open lädt eine Datei um mit dieser arbieten zu können. Dazu wird ein File descriptor zurückgegeben. Write schreibt dann von einem übergebenen buffer in die Datei die durch einen file descriptor beschrieben wird. Mmap erzeugt eine neue Mapping in dem virtuellen Addressspeicher des Prozesses. Brk steht für program break und dieser markiert das Ende von dem Speicher der einem Prozess zusteht. Durch das verschieben kann man den zu Verfügung gestellten Speicher also erweitern oder verkleinern.

\item Strace kann dazu genutzt werden um die Systemcalls die ein Anwendungsprogramm ausführt zu loggen.

\item Da \verb|ls| nur die Namen der Dateien in dem Ordner ausgeben muss, sind nur wenige Systemcalls notwending (20 bei uns). \verb|ls -la| gibt aber mehr Dateien aus (auch die die mit \verb|.| anfangen) und zu dem Namen zusätzlich noch andere Daten aus (Berechtigungen, User, Date, etc.). Deswegen sind hier deutlich mehr Systemcalls notwendig (294 bei uns). Zudem werden bei \verb|ls| hauptsächlich \verb|openstat| (9) und \verb|fstat| (10) verwendet, bei \verb|ls -la| allerdings eher \verb|lstat| (233) und \verb|openstat| (38). Daran kann man sehen, dass auch nur eine Paremeteränderung eines Commdands die Funktionsweise des Befehls komplett verändern kann.
\end{enumerate}

\end{document}

### 2020-05-16 slides/slides.tex
Feedback

\documentclass[english,aspectratio=43,t]{beamer}
% useful options:
%   aspectratio=169

\usepackage[utf8]{inputenc}
\usepackage{babel}
\usepackage{fixltx2e}
\usepackage{graphicx}
\usepackage{longtable}
\usepackage{float}
\usepackage{wrapfig}
\usepackage{soul}
\usepackage{textcomp}
\usepackage{marvosym}
\usepackage{wasysym}
\usepackage{latexsym}
\usepackage{amssymb}
\usepackage{hyperref}
\tolerance=1000
\usepackage{listings}
\usepackage[iso]{isodate}

% select theme before using any of its features
\usetheme{Comsys}

\title[Cool]{The Origins of Congestion\\and Network assisted end-to-end\\Congestion Control}
\author{Til Mohr}
\authorextra{Advisor: Jan Rüth}
\subtitle{Proseminar}
\date{\today}
%\location{Aachen}
%\projectlogo[scale=0.5]{logos/rwth_comsys_cropped}
%\logoextra{logos/rwth_comsys_cropped}

%\titlegraphics{images/presentation}
%\begin{titlepicture}
%	\includegraphics[width=3cm]{images/presentation}
%\end{titlepicture}

\begin{document}

\titleframe

\begin{frame}
	\frametitle{Outline}
	\tableofcontents
\end{frame}

\section{First Topic}
\label{sec:first_topic}
\frame\sectionpage

\begin{frame}{Blocks}
	\begin{cols}{0.5,auto}
		\begin{block}{Block with Title}
			Content of block with title
		\end{block}
		\begin{whiteblock}{White Block}
			Just a blue frame around the text
		\end{whiteblock}
		\begin{exampleblock}{This is an Example Block}
			Example Block
		\end{exampleblock}
		\begin{alertblock}{This is an Alert Block}
			Alert Block
		\end{alertblock}
	\col
		\begin{block}{}
			Content of block without title
		\end{block}
		\begin{whiteblock}{}
			Without title
		\end{whiteblock}
		\begin{grayblock}{}
			Grayed out block
		\end{grayblock}
		\begin{exampleblock}{}
			Example Block without Title
		\end{exampleblock}
		\begin{redblock}{}
			Red Block without Title
		\end{redblock}
	\end{cols}
	\pushdown
	\begin{alertblock}{}
		\begin{center}Don't overload your slides, beware of the shadow!\end{center}
	\end{alertblock}
\end{frame}

\begin{frame}{Itemize}
	\begin{cols}{1,2*1}
		\begin{itemize}
		\item item
		\item item
		\item []
		\item item
			\begin{itemize}
			\item subitem
				\begin{itemize}
				\item subsubitem
				\end{itemize}
			\end{itemize}
		\end{itemize}
	\col
		\begin{enumerate}
		\item Enumerated item
		\item Enumerated item
		\item[]
		\item Enumerated item
		\item Enumerated item
		\end{enumerate}
	\end{cols}

	\vspace{1em}

	\begin{cols}{auto}
		\begin{block}{Block with Items}
			\begin{itemize}
			\item item
				\begin{itemize}
				\item subitem
					\begin{itemize}
					\item subsubitem
					\end{itemize}
				\end{itemize}
			\end{itemize}
		\end{block}
		\begin{block}{}
			\begin{enumerate}
			\item item
			\item item
			\end{enumerate}
		\end{block}
	\col
		\begin{exampleblock}{}
			\begin{itemize}
			\item item
				\begin{itemize}
				\item subitem
					\begin{itemize}
					\item subsubitem
					\end{itemize}
				\end{itemize}
			\end{itemize}
		\end{exampleblock}
		\begin{alertblock}{}
			\begin{itemize}
			\item item
				\begin{itemize}
				\item subitem
					\begin{itemize}
					\item subsubitem
					\end{itemize}
				\end{itemize}
			\end{itemize}
		\end{alertblock}
	\col
		\begin{whiteblock}{}
			\begin{itemize}
			\item item
				\begin{itemize}
				\item subitem
					\begin{itemize}
					\item subsubitem
					\end{itemize}
				\end{itemize}
			\end{itemize}
		\end{whiteblock}
		\begin{grayblock}{}
			\begin{itemize}
			\item item
				\begin{itemize}
				\item subitem
					\begin{itemize}
					\item subsubitem
					\end{itemize}
				\end{itemize}
			\end{itemize}
		\end{grayblock}
	\end{cols}
\end{frame}

\begin{frame}{Illustrative Pictures right of Main Text}
	\begin{itemize}
	\item A picture can be inserted which is right aligned
		\begin{itemize}
		\item It is vertically aligned with the position\ilgraphics[height=2em]{logos/rwth}
		\\it was inserted in the source
		\end{itemize}
	\item There are two variants
		\begin{itemize}
		\item A variant based on \textit{{\textbackslash}includegraphics} which takes the same arguments as \textit{{\textbackslash}includegraphics}
		\item An environment for any content
		\begin{illustration}
			\begin{tikzpicture}
				\draw[draw=red] (0,0) circle (1em);
				\draw[draw=blue] (0,0) rectangle(2em,2em);
			\end{tikzpicture}
		\end{illustration}
		\end{itemize}
	\end{itemize}
\end{frame}

\singlemessageframe{This Frame contains\\only a single Message!}

\begin{frame}{Sizes}%
	\showsizes[cm]%
	\showfonts
\end{frame}

\end{document}

### 2020-05-16 paper/Übungsblatt 2/code/script.sh

@@ -1,0 +1,11 @@
+#!/bin/bash
+while true;
+do
+if ps -p "$1" > /dev/null
+then
+echo Process $1 is running.
+else
+echo Process $1 is not running.
+fi
+sleep $2;
+done

### 2020-05-16 paper/Übungsblatt 2/code/script_2.4_a.sh

@@ -1,0 +1,5 @@
+#!/bin/bash
+
+for e in $(ls -a); do
+    echo $(stat -c '%s %n' "$e")
+done

### 2020-05-16 paper/Übungsblatt 2/code/script_2.4_b.sh

@@ -1,0 +1,5 @@
+#!/bin/bash
+
+for e in $(ls -a); do
+    echo $(stat -c '%n %s' "$e")
+done

### 2020-05-16 paper/Übungsblatt 2/code/script_2.4_c.sh

@@ -1,0 +1,17 @@
+#!/bin/bash
+
+# Parameter is team size
+teams () {
+    FILE=$(sort teamnamen.txt | uniq -c | sed 's/^ *//' | grep -c "^$1")
+    echo ${FILE}
+}
+
+if [ $# -eq 0 ]; then
+    teams 3
+else
+    if [ $1 -eq "0" ]; then
+        echo $(grep -cvP '\S' teamnamen.txt)
+    else 
+        teams "$1"
+    fi
+fi

### 2020-05-16 paper/Übungsblatt 2/code/script_2.4_d.sh

@@ -1,0 +1,5 @@
+#!/bin/bash
+
+for i in $(seq 0 3); do
+    echo $i $(source ./script_2.4_c.sh $i)
+done

### 2020-05-16 paper/Übungsblatt 2/code/stack_exercise.c

@@ -1,0 +1,79 @@
+#include <limits.h>
+#include <stdio.h>
+#include <stdlib.h>
+#include <stdbool.h>
+
+/**
+ * Struktur, die Eintraege des Stacks kapselt.
+ */
+typedef struct StackNode
+{
+    // Zeiger auf den Command-String
+    char *command;
+    // Zeiger auf den naechsten Knoten des Stacks
+    struct StackNode *next_node;
+} StackNode;
+
+/**
+ * Erstellt einen neuen Knoten des Stacks und haengt ihn am Anfang an
+ * Gibt den neuen Startknoten zurück.
+ */
+StackNode *newNode(char *command)
+{
+    // allokiere neuen Speicherplatz für den naechsten Knoten auf dem Heap
+    StackNode *node = (StackNode *)malloc(sizeof(StackNode));
+    node->command = command;
+    node->next_node = NULL;
+    return node;
+}
+
+/**
+ * Teste, ob Stack leer ist
+ */
+bool isEmpty(StackNode *root)
+{
+    // Falls root auf NULL zeigt, ist der Stack leer
+    return !root;
+}
+
+/**
+ * Fuege einen neuen Knoten dem Stack hinzu.
+ */
+void push(StackNode **pointer2root, char *command)
+{
+    StackNode *node = newNode(command);
+    node->next_node = *pointer2root;
+    *pointer2root = node;
+}
+
+/**
+ * Gib den obersten Knoten zurueck und loesche diesen vom Stack (inklusive Speicherplatzfreigabe)
+ */
+bool pop(StackNode **pointer2root, char **pointer2command)
+{
+    if(isEmpty(*pointer2root)) return false;
+    StackNode *root = *pointer2root;
+    *pointer2command = ((*root).command);
+    *pointer2root = ((*root).next_node);
+    free(root);
+    return true;
+}
+
+int main()
+{
+    StackNode *root = NULL;
+
+    push(&root, "create: grades.txt");
+    push(&root, "add: Max Mustermann 1.0");
+    push(&root, "delete: grades.txt");
+
+    char *command;
+    while (pop(&root, &command))
+    {
+        printf("'%s' popped from stack \n", command);
+    }
+
+    printf("Stack is empty. Exit \n");
+
+    return 0;
+}

### 2020-05-16 paper/Übungsblatt 2/code/verzeichnis_struktur.sh

@@ -1,0 +1,34 @@
+#!/bin/bash
+
+# Basename
+# Anzahl tabs
+print_directory () {
+    for e in "$1"*; do
+        if [ -f "$e" ]; then
+            for i in $(seq 1 $2); do
+                printf "\t";
+            done
+            OUTPUT=$(sed 's/^\.//g' <<<$e)
+            OUTPUT=$(sed 's:^/::g' <<<$OUTPUT)
+            printf "File: $OUTPUT \n";
+        fi
+    done
+
+    for e in "$1"*; do
+        if [ -d "$e" ]; then
+            for i in $(seq 1 $2); do
+                printf "\t";
+            done
+            OUTPUT=$(sed 's/^\.//g' <<<$e)
+            OUTPUT=$(sed 's:^/::g' <<<$OUTPUT)
+            printf "Directory: $OUTPUT \n"
+            print_directory "$e/" $(($2+1))
+        fi
+    done
+}
+
+if [ $# -eq 0 ]; then
+    print_directory "./" 0
+else
+    print_directory "$1" 0
+fi

### 2020-05-21 paper/paper.tex
Citation Titles didn't have upper-case error fix

@@
 It is split up into multiple 32-bit blocks.
-The data block contains all the application data to be transmitted \cite{ietf-tcpm-rfc793bis-16,jacobson1992tcp,huston2000tcp}.}
+The data block contains all the application data to be transmitted \cite{jacobson1992tcp,ietf-tcpm-rfc793bis-16,huston2000tcp}.}
 \label{fig:TCP_Header}
@@
 The most important one is the \textit{ACK} flag (Acknowledgement bit).
-It is used whenever a packet arrives at its destination and an acknowledgement packet has to be sent back.
+Whenever a data packet arrives at its destination, a packet acknowledging the arrival of this packet will be send back.
+This acknowledgement packet always has the \textit{ACK} flag set.
 The \textit{SYN} and \textit{FIN} flags initialize and terminate a connection.
@@
 \item The \textit{Window Size} is the maximum amount of data (in bytes) a receiver is willing to receive at any time.
-It is being negotiated at the initialization of a new connection and can only be negotiated again wile the \textit{SYN} flag is set.
-It is important to set it appropriately to prevent unnecessary packet loss or delay, as the receiver couldn't process the data in time \cite{jacobson1992tcp}.
+It is important to set it appropriately to prevent unnecessary packet loss or delay, as the receiver couldn't process the data in time.
+Thus, it is will also be constantly adjusted to improve the connections reliability \cite{jacobson1992tcp}.
 \end{itemize}
@@
 To implicitly notify TCP that congestion occurs, congested network nodes will occasionally drop packets after a certain threshold of queued packets has been exceeded.
-Usually, this is quite smaller than the maximum queue length, therefore avoiding actions can be taken in time \cite{ramakrishnan1999proposal}.\\
-Note however, a high RTT can also have other causes than packet loss or congestion.
-A short term drastic decrease in the receivers processing performance can lead to packets being buffered at the receivers end for a prolonged period of time, thus resulting in a higher RTT \cite{huston2000tcp}.
+Usually, this is quite smaller than the maximum queue length, therefore avoiding actions can be taken in time \cite{ramakrishnan1999proposal}.
 
@@
 The problem that comes with classic TCP is that the congestion detection algorithms are very implicit.
-As already mentioned, packet loss for example may have different causes other than network congestion, however, the congestion avoidance algorithm will still interpret it as congestion, thus possibly resulting in unnecessary network delay.
+Packet loss for example may have different causes other than network congestion, such as simple cable interference, however, the congestion avoidance algorithm will still interpret it as congestion, thus possibly resulting in unnecessary network delay.
 Therefore, exactly determining whether the network is congested or not is crucial to keep a network stable.
@@
 Since network nodes have to actively detect congestion, ECN, XCP, and other network assisted congestion control protocols rely on active queue management in those nodes.
-Although there is hardware being produced today that meet those requirements, a large part of internet hardware isn't capable of doing so.
+Although there is some hardware being produced today that meet ECNs requirements, a large part of internet hardware isn't capable of doing so.
 Because any of those nodes can be congested, it would be counterintuitive to run those protocols, as congestion might or might not be detected, thus resulting in performance loss \cite{katabi2002congestion}.

### 2020-05-21 paper/Übungsblatt 2/code/script.sh

@@ -1,11 +1,0 @@
-#!/bin/bash
-while true;
-do
-if ps -p "$1" > /dev/null
-then
-echo Process $1 is running.
-else
-echo Process $1 is not running.
-fi
-sleep $2;
-done

### 2020-05-21 paper/Übungsblatt 2/code/script_2.4_a.sh

@@ -1,5 +1,0 @@
-#!/bin/bash
-
-for e in $(ls -a); do
-    echo $(stat -c '%s %n' "$e")
-done

### 2020-05-21 paper/Übungsblatt 2/code/script_2.4_b.sh

@@ -1,5 +1,0 @@
-#!/bin/bash
-
-for e in $(ls -a); do
-    echo $(stat -c '%n %s' "$e")
-done

### 2020-05-21 paper/Übungsblatt 2/code/script_2.4_c.sh

@@ -1,17 +1,0 @@
-#!/bin/bash
-
-# Parameter is team size
-teams () {
-    FILE=$(sort teamnamen.txt | uniq -c | sed 's/^ *//' | grep -c "^$1")
-    echo ${FILE}
-}
-
-if [ $# -eq 0 ]; then
-    teams 3
-else
-    if [ $1 -eq "0" ]; then
-        echo $(grep -cvP '\S' teamnamen.txt)
-    else 
-        teams "$1"
-    fi
-fi

### 2020-05-21 paper/Übungsblatt 2/code/script_2.4_d.sh

@@ -1,5 +1,0 @@
-#!/bin/bash
-
-for i in $(seq 0 3); do
-    echo $i $(source ./script_2.4_c.sh $i)
-done

### 2020-05-21 paper/Übungsblatt 2/code/stack_exercise.c

@@ -1,79 +1,0 @@
-#include <limits.h>
-#include <stdio.h>
-#include <stdlib.h>
-#include <stdbool.h>
-
-/**
- * Struktur, die Eintraege des Stacks kapselt.
- */
-typedef struct StackNode
-{
-    // Zeiger auf den Command-String
-    char *command;
-    // Zeiger auf den naechsten Knoten des Stacks
-    struct StackNode *next_node;
-} StackNode;
-
-/**
- * Erstellt einen neuen Knoten des Stacks und haengt ihn am Anfang an
- * Gibt den neuen Startknoten zurück.
- */
-StackNode *newNode(char *command)
-{
-    // allokiere neuen Speicherplatz für den naechsten Knoten auf dem Heap
-    StackNode *node = (StackNode *)malloc(sizeof(StackNode));
-    node->command = command;
-    node->next_node = NULL;
-    return node;
-}
-
-/**
- * Teste, ob Stack leer ist
- */
-bool isEmpty(StackNode *root)
-{
-    // Falls root auf NULL zeigt, ist der Stack leer
-    return !root;
-}
-
-/**
- * Fuege einen neuen Knoten dem Stack hinzu.
- */
-void push(StackNode **pointer2root, char *command)
-{
-    StackNode *node = newNode(command);
-    node->next_node = *pointer2root;
-    *pointer2root = node;
-}
-
-/**
- * Gib den obersten Knoten zurueck und loesche diesen vom Stack (inklusive Speicherplatzfreigabe)
- */
-bool pop(StackNode **pointer2root, char **pointer2command)
-{
-    if(isEmpty(*pointer2root)) return false;
-    StackNode *root = *pointer2root;
-    *pointer2command = ((*root).command);
-    *pointer2root = ((*root).next_node);
-    free(root);
-    return true;
-}
-
-int main()
-{
-    StackNode *root = NULL;
-
-    push(&root, "create: grades.txt");
-    push(&root, "add: Max Mustermann 1.0");
-    push(&root, "delete: grades.txt");
-
-    char *command;
-    while (pop(&root, &command))
-    {
-        printf("'%s' popped from stack \n", command);
-    }
-
-    printf("Stack is empty. Exit \n");
-
-    return 0;
-}

### 2020-05-21 paper/Übungsblatt 2/code/verzeichnis_struktur.sh

@@ -1,34 +1,0 @@
-#!/bin/bash
-
-# Basename
-# Anzahl tabs
-print_directory () {
-    for e in "$1"*; do
-        if [ -f "$e" ]; then
-            for i in $(seq 1 $2); do
-                printf "\t";
-            done
-            OUTPUT=$(sed 's/^\.//g' <<<$e)
-            OUTPUT=$(sed 's:^/::g' <<<$OUTPUT)
-            printf "File: $OUTPUT \n";
-        fi
-    done
-
-    for e in "$1"*; do
-        if [ -d "$e" ]; then
-            for i in $(seq 1 $2); do
-                printf "\t";
-            done
-            OUTPUT=$(sed 's/^\.//g' <<<$e)
-            OUTPUT=$(sed 's:^/::g' <<<$OUTPUT)
-            printf "Directory: $OUTPUT \n"
-            print_directory "$e/" $(($2+1))
-        fi
-    done
-}
-
-if [ $# -eq 0 ]; then
-    print_directory "./" 0
-else
-    print_directory "$1" 0
-fi

### 2020-05-25 paper/paper.tex
Feedback

@@
 \item 
-\textit{Transmission delay} describes the delay of transmitting data packets between two internet nodes.
+\textit{Transmission delay} is the time it takes to send a packet from one node to another.
+More specifically, it describes the delay of transmitting data packets between two internet nodes depending on bandwidth and length of the connection.
 Typically, packets sent from an origin to a destination node are sent over various different internet nodes, all of which could have a different bandwidth.
-Therefore, a packet send from \textbf{node 1} to \textbf{node n} over \textbf{nodes 2 to n-1} cannot be sent at a faster rate than the lowest bandwidth on the path of the packet (bottleneck) [Figure \ref{fig:transmission delay}].\[Rate_{max} \leftarrow \min\{bandwidth_{i}\mid i\in\{1,...,n-1\}\}\]
+Therefore, a packet send from \textbf{node 1} to \textbf{node n} over \textbf{nodes 2 to n-1} cannot be received at a faster rate than the lowest bandwidth on the path of the packet (bottleneck) [Figure \ref{fig:transmission delay}].\[Rate_{max} \leftarrow \min\{bandwidth_{i}\mid i\in\{1,...,n-1\}\}\]
+
+\item
+Each link creates an intentional gap between two packets being sent.
+This pause is often necessary to allow the receiver to prepare for a new package, however, depending on the transmission protocol, other purposes may apply.
+This kind of delay is known as \textit{Inter-Packet delay}.
 
@@
 
-Whereas processing delay is almost not noticeable and transmission delay could be improved by upgrading cables or transmission protocols, queuing delay has various causes and therefore is the  most difficult one to reduce.
+It is important to understand that delay has a direct impact on congestion.
+When delay increases, the networks throughput can decrease as packets will stay inside the network for a longer time.
+If the network however continues to receive new packets at constant rate, more and more packets will have to be stored, resulting in congestion.
+Generally speaking, to tackle congestion, firstly delay can be reduced, and secondly the networks storages, the buffers, can be improved regarding efficient queue management and buffer size.
+Finally, transmission protocols can also react to congestion and adapt their send rate to recover from congestion [\nameref{sec:ca}] and even control congestion [\nameref{sec:cc}].
+\\Whereas processing delay is almost not noticeable and transmission and inter-packet delay could be improved by upgrading cables or transmission protocols, queuing delay has various causes and therefore is the  most difficult one to reduce.
 There are many different ways to implement buffers in a network node and the effectiveness of each implementation varies from where it is used.
@@
 Bufferbloat refers to frequently full and excessively large buffers in internet nodes.
-This leads to partly unnecessary network delay and damages congestion avoidance algorithms, such as those found in classic TCP \cite{gettys2012bufferbloat,chen2014bufferbloat}.
+This leads to partly unnecessary network delay and damages congestion avoidance algorithms, such as those found in classic TCP, which will be explained in the section \nameref{sec:ca} \cite{gettys2012bufferbloat,chen2014bufferbloat}.
 Buffering is needed to provide space to queued packets waiting for transmission resulting in a minimization of packet loss.
@@
 To fully understand, how TCP's congestion avoidance algorithm operates, one needs to understand how TCP works.
-TCP lives on the transport layer.
+In the following TCP's congestion avoidance algorithm will be explained using the TCP Specification \cite{ietf-tcpm-rfc793bis-16,postel1981transmission}, also being refered to as \textit{classic TCP}.
+Note, there are also other TCP implemtations improving upon it.
+The underlying concept however is the same.
+\\TCP lives on the transport layer.
 It communicates with the application layer via sockets, splits up data into packets (and vice versa), and transmits those packages over the network with the help of the internet protocol, thus TCP is often also referred to as \textit{TCP/IP}.
@@
 \subsection{Perceiving Congestion}
-\label{P_C_marker}
+\label{sec:P_C_marker}
 Both partners of a connection are both sender and receiver, as both transmit packets.
@@
 Usually, this is quite smaller than the maximum queue length, therefore avoiding actions can be taken in time \cite{ramakrishnan1999proposal}.
-
-\subsection{Adjusting for Congestion}
-\label{AfC}
-To prevent congestion altogether, packet flow inside a network would have to be conservative, i.e.
-a new packet isn't being sent into the network before an old packet leaves the network.
-A connection in this state is also being described as being \textit{in equilibrium} \cite{jacobson1992tcp}.
-Whenever a connection is in equilibrium, congestion in the network cannot be caused by this TCP connection, thus also controlling congestion.
-TCPs congestion avoidance algorithm can be categorized into three sub-algorithms, that all have the goal to keep the connection in equilibrium, no matter how congestion was perceived.
-Here, a variable called \textit{congestion window} (cwnd) has an important role.
-It describes the maximum amount of packets sent during an average RTT.
-An additional mechanism of keeping a connection in equilibrium is, that each acknowledgment packet received by the sender opens the space for a new packet to be sent into the network \cite{jacobson1992tcp}.
-The principles TCP uses to adjust for congestion are based on \textit{Additive-Increase Multiplicative-Decrease} (AIMD) principle.
-
-\subsubsection*{Slow-Start}
-The \textit{slow-start} algorithm is used when a connection has just been established or when multiple packet losses were detected.
-The cwnd variable is set to one packet per estimated RTT, because no information about network congestion has been collected yet.
-Whenever an acknowledgment packet is being received, cwnd will increase by one packet.
-This doubles cwnd each RTT, resulting in an exponential growth over time, which helps to get the packet send rate to the maximum the network can handle quite quickly, contributing to almost immediately maximum network performance \cite{jacobson1992tcp}.
-
-\subsubsection*{Adaption}
-When the network is congested, buffers in network nodes will fill up exponentially \cite{jacobson1992tcp}.
-Therefore, the network can only stabilize if the amount of packets sent decreases at least as quickly as the queues are growing.
-The formula \cite{jacobson1992tcp} \[cwnd_{i+1} \leftarrow d \cdot cwnd_{i}, 0<d<1\] has an exponential decrease over time, thus resulting in the network recovering.\\
-If the network isn't congested, the sender should try to raise cwnd slowly to reach the networks maximum throughput again.
-A multiplicative, exponential increase over time, like the slow-start algorithm isn't the best choice here.
-Increasing the amounts of packets sent drastically while cwnd still is high can challenge the networks capacity, resulting in heavy congestion.
-Therefore, the best option to raise cwnd to its limits is an additive increase \cite{jacobson1992tcp}:\[cwnd_{i+1} \leftarrow cwnd_{i} + u\] Van Jacobson proposes in his 1992 paper about TCP congestion avoidance that $d$ will be set to $\frac{1}{2}$ and $u$ will be set to $\frac{1}{cwnd_{i}}$.
-Thus, cwnd can only increase by a maximum of one packet per RTT leading to a smooth approach to the networks limit \cite{jacobson1992tcp}.
 
@@
 
+\subsection{Adjusting for Congestion}
+\label{sec:AfC}
+To prevent congestion altogether, packet flow inside a network would have to be conservative.
+A connection in this state is also being described as being \textit{in equilibrium} \cite{jacobson1992tcp}.
+This simply means, the networks works at full load whilst no additional delay or congestion is being created.
+However, it is quite impossible to keep a full network in this state forever.
+The goal of TCP's congestion avoidance algorithms, which are based on an \textit{Additive-Increase Multiplicative-Decrease} (AIMD) principle, is to react to the current state of congestion and achieve equilibrium.
+Those algorithms can be categorized into three sub-algorithms, that all have the goal to keep the connection in equilibrium, no matter how congestion was perceived.
+Here, a variable called \textit{congestion window} (cwnd) has an important role.
+It describes the maximum amount of data sent during an average RTT.
+An additional mechanism of keeping a connection in equilibrium is, that each acknowledgment packet received by the sender opens the space for a new packet to be sent into the network \cite{jacobson1992tcp}.
+The principles TCP uses to adjust for congestion are based on \textit{Additive-Increase Multiplicative-Decrease} (AIMD) principle.
+
+\subsubsection*{Slow-Start}
+The \textit{slow-start} algorithm is used when a connection has just been established or when multiple packet losses were detected.
+The cwnd variable is set to one packet per estimated RTT, because no information about network congestion has been collected yet.
+Whenever an acknowledgment packet is being received, cwnd will increase by one packet.
+This doubles cwnd each RTT, resulting in an exponential growth over time, which helps to get the packet send rate to the maximum the network can handle quite quickly, contributing to almost immediately maximum network performance \cite{jacobson1992tcp}.
+
+\subsubsection*{Adaption}
+When the network is congested, buffers in network nodes will fill up exponentially \cite{jacobson1992tcp}.
+Therefore, the network can only stabilize if the amount of packets sent decreases at least as quickly as the queues are growing.
+The formula \cite{jacobson1992tcp} \[cwnd_{i+1} \leftarrow d \cdot cwnd_{i}, 0<d<1\] has an exponential decrease over time, thus resulting in the network recovering.\\
+If the network isn't congested, the sender should try to raise cwnd slowly to reach the networks maximum throughput again.
+A multiplicative, exponential increase over time, like the slow-start algorithm isn't the best choice here.
+Increasing the amounts of packets sent drastically while cwnd still is high can challenge the networks capacity, resulting in heavy congestion.
+Therefore, the best option to raise cwnd to its limits is an additive increase \cite{jacobson1992tcp}:\[cwnd_{i+1} \leftarrow cwnd_{i} + u\] Van Jacobson proposes in his 1992 paper about TCP congestion avoidance that $d$ will be set to $\frac{1}{2}$ and $u$ will be set to $\frac{1}{cwnd_{i}}$.
+Thus, cwnd can only increase by a maximum of one packet per RTT leading to a smooth approach to the networks limit \cite{jacobson1992tcp}.
+
 \section{Network Assisted Congestion Control}
+\label{sec:cc}
 The problem that comes with classic TCP is that the congestion detection algorithms are very implicit.
@@
 This is done by modifying the IP header to be able to carry one more bit - the \textit{Congestion Experienced} flag (CE) \cite{ramakrishnan1999proposal}.
-Instead of dropping packets occasionally after the amount of queued packets exceed a threshold [\nameref{P_C_marker}], all packets supporting ECN will now be marked (setting the CE bit in the IP header) [Figure \ref{fig:ecn}].
+Instead of dropping packets occasionally after the amount of queued packets exceed a threshold [\nameref{sec:P_C_marker}], all packets supporting ECN will now be marked (setting the CE bit in the IP header) [Figure \ref{fig:ecn}].
 Thus, congestion can be described explicitly resulting in better short- and long-term network performance.
@@
 \[H\_feedback \leftarrow \frac{r \cdot RTT - cwnd}{cwnd}\]
-This is an improvement to TCPs adjustment algorithms as the desired rate can be reached after one RTT.\\
+This is an improvement to TCP's adjustment algorithms as the desired rate can be reached after one RTT.\\
 The other task an XCP sender has is to adjust the congestion window according to the networks feedback.
@@
 Hence, the \textit{Fairness Controller} (FC) comes into play.
-As TCP, FC relies on the AIMD principle [\nameref{AfC}].
+As TCP, FC relies on the AIMD principle [\nameref{sec:AfC}].
 If the EC calculated positive feedback, it will be split up between all packets equally.
@@
 Although there is some hardware being produced today that meet ECNs requirements, a large part of internet hardware isn't capable of doing so.
-Because any of those nodes can be congested, it would be counterintuitive to run those protocols, as congestion might or might not be detected, thus resulting in performance loss \cite{katabi2002congestion}.
+As many routers do not run on software but hardware only to improve performance, they can only handle specific packets.
+Whenever they cannot recognize a packet, it will simply be dropped \cite{katabi2002congestion}.
+So, although ECN packets differ from ordinary TCP/IP packets by only one bit, not all routers can handle this change and will drop those packets.
 This is the reason you do not see those protocols in wide area networks (WAN) too often, let alone the internet.
 \\Nevertheless, whenever the implementation of XCP, ECN, or similar protocols is not a challenge, you can implement network assisted congestion control with ease and see great performance improvements.
-Especially smaller networks with high traffic could definitely benefit from those technologies.
-
-\subsection{DCTCP}
-A lot of internet services run in data centers with large amounts of computers communicating with each other, resulting in a lot of traffic inside the data center.
-Optimizing those data centers networking capabilities can result in an increase in application productivity.
-In 2010 a paper proposed \textit{DCTCP}, a variation of TCP using ECN to create a data transportation protocol, which especially takes hardware purposed for use in those environments into account \cite{10.1145/1851275.1851192}.\\
-The general problem that appears in high-traffic environments is the so-called \textit{TCP Incast} problem.
-It describes a drastic reduction in application throughput.
-To understand where TCP incast originates, you fist need to understand how most data centers are built.
-A typical data center contains a set of racks, each holding a large number of servers, the switches, that connect those racks, and connecting links, that connect those switches to other parts of the data center or the internet [Figure \ref{fig:dc}].
-This design pattern is also known as \textit{Partition/Aggregate} and is most commonly used in data centers.
-It however results in a high 'fan-in', as multiple servers (leaves) are connected to just one link (root) \cite{10.1145/1851275.1851192,10.1145/1592681.1592693}.
-In addition, most applications require high bandwidth, low latency connections, also enabling them to send many parallel requests to other servers.
-Combining those requirements with the design pattern, it becomes clear, that those switches are being challenged.
-Besides, many data center switches have relatively small, shared buffers.
-Therefore, it is common that queue lengths stay high for a prolonged period of time, resulting in a collapse of throughput \cite{10.1145/1851275.1851192,10.1145/1592681.1592693}.\\
-The goal of DCTCP is to keep data center traffic constantly flowing, which means meeting the requirements applications take while keeping relatively small buffer sizes inside switches in mind.
-To achieve this, DCTCP implements ECN.
-Unlike TCP however, if a DCTCP sender registers congestion, the window size will be reduced by a fraction of the amount of marked packets it received.
-Therefore, excessive link underutilization can be prevented.
-In addition, the DCTCP receiver will only send one ACK packet for every $m$ consecutive data packet, where all $1,...,m$ packets before had the same CE-flag configuration.
-Here, $m$ is a variable initialized while the DCTCP connection is being established and cannot be changed without renegotiation.
-Therefore, fewer packets are sent through the network, but the TCP sender is still able to identify lost packets and register congestion \cite{10.1145/1851275.1851192}.
-\\Being able to react to congestion immediately enables DCTCP to reduce buffer overflows and timeouts, thus largely resolving incast problems.
-The paper proposing DCTCP has found, that their protocol increases traffic by 110\% while using 90\% less buffer space compared to TCP \cite{10.1145/1851275.1851192}.
+Especially smaller networks with high traffic could definitely benefit from those technologies [\nameref{sec:DCTCP}].
 
@@
 
+\subsection{DCTCP}
+\label{sec:DCTCP}
+A lot of internet services run in data centers with large amounts of computers communicating with each other, resulting in a lot of traffic inside the data center.
+Optimizing those data centers networking capabilities can result in an increase in application productivity.
+In 2010 a paper proposed \textit{DCTCP}, a variation of TCP using ECN to create a data transportation protocol, which especially takes hardware purposed for use in those environments into account \cite{10.1145/1851275.1851192}.\\
+The general problem that appears in high-traffic environments is the so-called \textit{TCP Incast} problem.
+It describes a drastic reduction in application throughput.
+To understand where TCP incast originates, you fist need to understand how most data centers are built.
+A typical data center contains a set of racks, each holding a large number of servers, the switches, that connect those racks, and connecting links, that connect those switches to other parts of the data center or the internet [Figure \ref{fig:dc}].
+This design pattern is also known as \textit{Partition/Aggregate} and is most commonly used in data centers.
+It however results in a high 'fan-in', as multiple servers (leaves) are connected to just one link (root) \cite{10.1145/1851275.1851192,10.1145/1592681.1592693}.
+In addition, most applications require high bandwidth, low latency connections, also enabling them to send many parallel requests to other servers.
+Combining those requirements with the design pattern, it becomes clear, that those switches are being challenged.
+Besides, many data center switches have relatively small, shared buffers.
+Therefore, it is common that queue lengths stay high for a prolonged period of time, resulting in a collapse of throughput \cite{10.1145/1851275.1851192,10.1145/1592681.1592693}.\\
+The goal of DCTCP is to keep data center traffic constantly flowing, which means meeting the requirements applications take while keeping relatively small buffer sizes inside switches in mind.
+To achieve this, DCTCP implements ECN.
+Unlike TCP however, if a DCTCP sender registers congestion, the window size will be reduced by a fraction of the amount of marked packets it received.
+Therefore, excessive link underutilization can be prevented.
+In addition, the DCTCP receiver will only send one ACK packet for every $m$ consecutive data packet, where all $1,...,m$ packets before had the same CE-flag configuration.
+Here, $m$ is a variable initialized while the DCTCP connection is being established and cannot be changed without renegotiation.
+Therefore, fewer packets are sent through the network, but the TCP sender is still able to identify lost packets and register congestion \cite{10.1145/1851275.1851192}.
+\\Being bandwidth greedy transmission protocol, data centers using DCTCP can see a performance drop in nodes that use other transmission protocols such as TCP or ECN.
+As DCTCP tries to utilize the links as much as possible, non-greedy protocols will be neglected and will see a drop in performance.
+Thus, data centers should avoid using DCTCP in addition to other protocols to fully make use of DCTCP's capabilities.
+\\To sum up, being able to react to congestion immediately enables DCTCP to reduce buffer overflows and timeouts, thus largely resolving incast problems.
+The paper proposing DCTCP has found, that their protocol increases traffic by 110\% while using 90\% less buffer space compared to TCP \cite{10.1145/1851275.1851192}.
+
 \section{Conclusion}

### 2020-05-27 paper/paper.tex
Some more improvements

@@
 % author
-\author{\IEEEauthorblockN{Til Mohr}
-\IEEEauthorblockA{RWTH Aachen University\\til.mohr@rwth-aachen.de}}
+%\author{\IEEEauthorblockN{Til Mohr}
+%\IEEEauthorblockA{RWTH Aachen University\\til.mohr@rwth-aachen.de}}
 
@@
 Congestion was, is, and always will be a problem on the Internet, appearing in various forms with different symptoms causing minor, but also major challenges.
-Most data sent over the internet is split up in smaller packets that all try to reach their destination the fastest, whether it is by going the shortest or the least congested route.
+Most data sent over the internet is split up in smaller packets that all try to reach their destination the fastest, whether it is by taking the shortest or the least congested route.
 In an optimal network all computers, servers, etc.
@@
 Therefore, a network-wide reduction of those complications is an important task to improve the quality of all internet services.
-\\Network delay can generally speaking be categorized into three different types of delay \cite{gettys2012bufferbloat}:
+\\Network delay can generally speaking be categorized into four different types of delay \cite{gettys2012bufferbloat}:
 
@@
 \item
-Each link creates an intentional gap between two packets being sent.
+Each node creates an intentional gap between two packets being sent.
 This pause is often necessary to allow the receiver to prepare for a new package, however, depending on the transmission protocol, other purposes may apply.
@@
 Generally speaking, to tackle congestion, firstly delay can be reduced, and secondly the networks storages, the buffers, can be improved regarding efficient queue management and buffer size.
-Finally, transmission protocols can also react to congestion and adapt their send rate to recover from congestion [\nameref{sec:ca}] and even control congestion [\nameref{sec:cc}].
-\\Whereas processing delay is almost not noticeable and transmission and inter-packet delay could be improved by upgrading cables or transmission protocols, queuing delay has various causes and therefore is the  most difficult one to reduce.
+Finally, transmission protocols can also react to congestion and adapt their send and data rate to recover from congestion [\nameref{sec:ca}] and even control congestion [\nameref{sec:cc}].
+\\Whereas processing delay is almost not noticeable and transmission and inter-packet delay could be improved by upgrading cables, hardware, or transmission protocols, queuing delay has various causes and therefore is the  most difficult one to reduce.
 There are many different ways to implement buffers in a network node and the effectiveness of each implementation varies from where it is used.
@@
 To fully understand, how TCP's congestion avoidance algorithm operates, one needs to understand how TCP works.
-In the following TCP's congestion avoidance algorithm will be explained using the TCP Specification \cite{ietf-tcpm-rfc793bis-16,postel1981transmission}, also being refered to as \textit{classic TCP}.
+In the following, TCP's congestion avoidance algorithm will be explained using the TCP Specification \cite{ietf-tcpm-rfc793bis-16,postel1981transmission}, also being refered to as \textit{classic TCP}.
 Note, there are also other TCP implemtations improving upon it.
@@
 A connection in this state is also being described as being \textit{in equilibrium} \cite{jacobson1992tcp}.
-This simply means, the networks works at full load whilst no additional delay or congestion is being created.
-However, it is quite impossible to keep a full network in this state forever.
+Simply speaking this means, the network works at full load whilst no additional delay or congestion is being created.
+However, it is quite challenging to keep a full network in this state forever.
 The goal of TCP's congestion avoidance algorithms, which are based on an \textit{Additive-Increase Multiplicative-Decrease} (AIMD) principle, is to react to the current state of congestion and achieve equilibrium.
@@
 'Understanding XCP: Equilibrium and Fairness' has shown through simulations that many network topologies could benefit from XCP as it clears queues in equilibrium, thus resulting in improved network performance under load \cite{1498331}.
-Another paper has shown that XCP outperforms ordinary TCP in nearly any environment related to average queue lengths, efficiency, and throughput \cite{katabi2002congestion}.
+Another paper has shown that XCP outperforms ordinary TCP in nearly any environment in aspects of average queue lengths, efficiency, and throughput \cite{katabi2002congestion}.
 \\However, all this performance comes at a cost.
@@
 So, although ECN packets differ from ordinary TCP/IP packets by only one bit, not all routers can handle this change and will drop those packets.
-This is the reason you do not see those protocols in wide area networks (WAN) too often, let alone the internet.
+This is the reason you do not see these kind of protocols in wide area networks (WAN) too often, let alone the internet.
 \\Nevertheless, whenever the implementation of XCP, ECN, or similar protocols is not a challenge, you can implement network assisted congestion control with ease and see great performance improvements.
@@
 Therefore, fewer packets are sent through the network, but the TCP sender is still able to identify lost packets and register congestion \cite{10.1145/1851275.1851192}.
-\\Being bandwidth greedy transmission protocol, data centers using DCTCP can see a performance drop in nodes that use other transmission protocols such as TCP or ECN.
+\\However, being bandwidth greedy transmission protocol, data centers using DCTCP can see a performance drop when they additionally use other transmission protocols such as TCP or ECN.
 As DCTCP tries to utilize the links as much as possible, non-greedy protocols will be neglected and will see a drop in performance.

### 2020-05-27 paper/paper.tex
Einreichen Feedback

@@
 \\Networks today consist of network nodes (routers, switches, ...) which all try to route internet packets as quickly as possible whilst keeping the required hardware to a minimum.
-Ideally there would be a steady flow of packets being transmitted at maximum rate.
-However, when a internet node receives a packet while the outgoing connection still is in use, the packet has to be stored in a buffer until the outgoing connection freed up again.
-Therefore, those storage units are essential to every node on a network, but without proper management even well-sized buffers could lead to network problems like packet-loss, jitter, and connection delay \cite{appenzeller2004sizing}.
+Ideally there would be a steady flow of packets being transmitted at the maximum rate.
+However, when an internet node receives a packet while the outgoing connection still is in use, the packet has to be stored in a buffer until the outgoing connection freed up again.
+Therefore, those storage units are essential to every node on a network, but without proper management even well-sized buffers could lead to network problems like packet loss, jitter, and connection delay \cite{appenzeller2004sizing}.
 \\This paper will present origins of network congestion and analyze how different types of algorithms try to avoid and minimize congestion.
@@
 \section{Origins of Congestion}
-With the ever-growing amount of internet services not only data-centers, but the whole internet in general faces new challenges with each packet sent.
+With the ever-growing amount of internet services not only data centers, but the whole internet in general faces new challenges with each packet sent.
 Therefore, a network-wide reduction of those complications is an important task to improve the quality of all internet services.
@@
 More specifically, it describes the delay of transmitting data packets between two internet nodes depending on bandwidth and length of the connection.
-Typically, packets sent from an origin to a destination node are sent over various different internet nodes, all of which could have a different bandwidth.
+Typically, packets sent from an origin to a destination node are sent over various internet nodes, all of which could have a different bandwidth.
 Therefore, a packet send from \textbf{node 1} to \textbf{node n} over \textbf{nodes 2 to n-1} cannot be received at a faster rate than the lowest bandwidth on the path of the packet (bottleneck) [Figure \ref{fig:transmission delay}].\[Rate_{max} \leftarrow \min\{bandwidth_{i}\mid i\in\{1,...,n-1\}\}\]
@@
 \textit{Processing delay} is the time needed to process an individual packet by an internet node.
-This usually means processing the packets IP header, finding its destination and possibly queuing it in buffers.
+This usually means processing the packets IP header, finding its destination, and possibly queuing it in buffers.
 
@@
 It is important to understand that delay has a direct impact on congestion.
-When delay increases, the networks throughput can decrease as packets will stay inside the network for a longer time.
-If the network however continues to receive new packets at constant rate, more and more packets will have to be stored, resulting in congestion.
-Generally speaking, to tackle congestion, firstly delay can be reduced, and secondly the networks storages, the buffers, can be improved regarding efficient queue management and buffer size.
+When delay increases, the throughput of the network can decrease as packets will stay inside the network for a longer time.
+If the network however continues to receive new packets at a constant rate, more and more packets will have to be stored, resulting in congestion.
+Generally speaking, to tackle congestion, firstly delay can be reduced, and secondly the storages inside the network, the buffers, can be improved regarding efficient queue management and buffer size.
 Finally, transmission protocols can also react to congestion and adapt their send and data rate to recover from congestion [\nameref{sec:ca}] and even control congestion [\nameref{sec:cc}].
@@
 There are many different ways to implement buffers in a network node and the effectiveness of each implementation varies from where it is used.
-Traffic in residential network nodes for example differs from nodes used in large data-centers in every aspect, thus hardware and software in those components should be adjusted accordingly.
+Traffic in residential network nodes for example differs from nodes used in large data centers in every aspect, thus hardware and software in those components should be adjusted accordingly.
 \\Sizing buffers is a complex task.
 On the one hand read and write times should be as short as possible to minimize additional delay.
-On the other hand buffers should not be too small, as networks need to be robust for future changes and therefore nodes should store lots of data if needed, while also keeping in mind the requirements of congestion detection algorithms as described in section \nameref{sec:ca} in order to adjust for congestion appropriately \cite{staff2012bufferbloat}.
+On the other hand, buffers should not be too small, as networks need to be robust for future changes and therefore nodes should store lots of data if needed, while also keeping in mind the requirements of congestion detection algorithms as described in section \nameref{sec:ca} in order to adjust for congestion appropriately \cite{staff2012bufferbloat}.
 Unifying those goals is very challenging as it leads to increasing costs and physical space occupied by those buffers.
@@
 \label{sec:ca}
-As already mentioned, congestion detection algorithms make up an important part in congestion avoidance protocols.
+As already mentioned, congestion detection algorithms make up an important part of congestion avoidance protocols.
 \textit{TCP} is such a protocol, based on the transport layer, which adapts its send rate according to available network bandwidth \cite{1209197,jacobson1992tcp}.
 To fully understand, how TCP's congestion avoidance algorithm operates, one needs to understand how TCP works.
-In the following, TCP's congestion avoidance algorithm will be explained using the TCP Specification \cite{ietf-tcpm-rfc793bis-16,postel1981transmission}, also being refered to as \textit{classic TCP}.
-Note, there are also other TCP implemtations improving upon it.
+In the following, TCP's congestion avoidance algorithm will be explained using the TCP Specification \cite{ietf-tcpm-rfc793bis-16,postel1981transmission}, also being referred to as \textit{classic TCP}.
+Note, there are also other TCP implementations improving upon it.
 The underlying concept however is the same.
@@
 The most important one is the \textit{ACK} flag (Acknowledgement bit).
-Whenever a data packet arrives at its destination, a packet acknowledging the arrival of this packet will be send back.
-This acknowledgement packet always has the \textit{ACK} flag set.
+Whenever a data packet arrives at its destination, a packet acknowledging the arrival of this packet will be sent back.
+This acknowledgment packet always has the \textit{ACK} flag set.
 The \textit{SYN} and \textit{FIN} flags initialize and terminate a connection.
@@
 It is important to set it appropriately to prevent unnecessary packet loss or delay, as the receiver couldn't process the data in time.
-Thus, it is will also be constantly adjusted to improve the connections reliability \cite{jacobson1992tcp}.
+Thus, it is will also be constantly adjusted to improve the reliability of the connection \cite{jacobson1992tcp}.
 \end{itemize}
@@
 When the network is congested, buffers in network nodes will fill up exponentially \cite{jacobson1992tcp}.
-Therefore, the network can only stabilize if the amount of packets sent decreases at least as quickly as the queues are growing.
+Therefore, the network can only stabilize if the number of packets sent decreases at least as quickly as the queues are growing.
 The formula \cite{jacobson1992tcp} \[cwnd_{i+1} \leftarrow d \cdot cwnd_{i}, 0<d<1\] has an exponential decrease over time, thus resulting in the network recovering.\\
-If the network isn't congested, the sender should try to raise cwnd slowly to reach the networks maximum throughput again.
+If the network isn't congested, the sender should try to raise cwnd slowly to reach the network's maximum throughput again.
 A multiplicative, exponential increase over time, like the slow-start algorithm isn't the best choice here.
-Increasing the amounts of packets sent drastically while cwnd still is high can challenge the networks capacity, resulting in heavy congestion.
+Increasing the number of packets sent drastically while cwnd still is high can challenge the networks capacity, resulting in heavy congestion.
 Therefore, the best option to raise cwnd to its limits is an additive increase \cite{jacobson1992tcp}:\[cwnd_{i+1} \leftarrow cwnd_{i} + u\] Van Jacobson proposes in his 1992 paper about TCP congestion avoidance that $d$ will be set to $\frac{1}{2}$ and $u$ will be set to $\frac{1}{cwnd_{i}}$.
-Thus, cwnd can only increase by a maximum of one packet per RTT leading to a smooth approach to the networks limit \cite{jacobson1992tcp}.
+Thus, cwnd can only increase by a maximum of one packet per RTT leading to a smooth approach to the network's limit \cite{jacobson1992tcp}.
 
@@
 In 2001 a new method was standardized to explicitly notify TCP on congestion.
-\textit{Explicit Congestion Notification} (ECN) is an extension to TCP and IP, allowing end-to-end congestion notification, while also minimizing the amount of packets dropped.
+\textit{Explicit Congestion Notification} (ECN) is an extension to TCP and IP, allowing end-to-end congestion notification, while also minimizing the number of packets dropped.
 This is done by modifying the IP header to be able to carry one more bit - the \textit{Congestion Experienced} flag (CE) \cite{ramakrishnan1999proposal}.
-Instead of dropping packets occasionally after the amount of queued packets exceed a threshold [\nameref{sec:P_C_marker}], all packets supporting ECN will now be marked (setting the CE bit in the IP header) [Figure \ref{fig:ecn}].
+Instead of dropping packets occasionally after the number of queued packets exceed a threshold [\nameref{sec:P_C_marker}], all packets supporting ECN will now be marked (setting the CE bit in the IP header) [Figure \ref{fig:ecn}].
 Thus, congestion can be described explicitly resulting in better short- and long-term network performance.
@@
 As a result, avoiding actions can be taken immediately.
-In addition, the need to retransmit packets would decrease drastically, as no packets are being dropped anymore to indicate congestion.
+Also, the need to retransmit packets would decrease drastically, as no packets are being dropped anymore to indicate congestion.
 Therefore, this central algorithm is also being called \textit{Congestion Control} \cite{ramakrishnan1999proposal,ramakrishnan2001addition}.
@@
 \subsection{XCP}
-Although TCP is a very dominant protocol on the internet, there are also other transport protocols that could benefit from end-to-end congestion notification.
+Although TCP is a very dominant protocol on the internet, also other transport protocols could benefit from end-to-end congestion notification.
 The paper "Congestion Control for High Bandwidth-Delay Product Networks"  generalizes the ECN proposal of 2001, creating a new protocol called \textit{eXplicit Congestion Protocol} (XCP), while also improving upon it further by introducing new packet control concepts \cite{katabi2002congestion,1498331}.
@@
 This is an improvement to TCP's adjustment algorithms as the desired rate can be reached after one RTT.\\
-The other task an XCP sender has is to adjust the congestion window according to the networks feedback.
-When $s$ is the packet size, the congestion window variable would be adjusted by following formula:
+The other task an XCP sender has is to adjust the congestion window according to the network's feedback.
+When $s$ is the packet size, the congestion window variable would be adjusted by the following formula:
 \[cwnd \leftarrow \max(cwnd + H_feedback, s)\]
@@
 \subsubsection*{XCP Receiver}
-The XCP receiver functions exactly the same way like a TCP receiver would except for copying the (modified) congestion header from the data packet to its acknowledgment \cite{katabi2002congestion}.
+The XCP receiver functions the same way as a TCP receiver would except for copying the (modified) congestion header from the data packet to its acknowledgment \cite{katabi2002congestion}.
 
@@
 If the EC calculated positive feedback, it will be split up between all packets equally.
-If the feedback is negative however, it will be divided between packets in proportion to their current send-rates, which are determined by \textit{H\_cwnd} and \textit{H\_rtt} \cite{katabi2002congestion,1498331}.\\
+If the feedback is negative however, it will be divided between packets in proportion to their current send rates, which are determined by \textit{H\_cwnd} and \textit{H\_rtt} \cite{katabi2002congestion,1498331}.\\
 Decoupling the feedback computation allows XCP to quickly acquire the maximum send rate and achieve equilibrium, all while the network allocates bandwidth to all XCP senders equally \cite{katabi2002congestion}.
@@
 So, although ECN packets differ from ordinary TCP/IP packets by only one bit, not all routers can handle this change and will drop those packets.
-This is the reason you do not see these kind of protocols in wide area networks (WAN) too often, let alone the internet.
+This is the reason you do not see these kinds of protocols in wide area networks (WAN) too often, let alone the internet.
 \\Nevertheless, whenever the implementation of XCP, ECN, or similar protocols is not a challenge, you can implement network assisted congestion control with ease and see great performance improvements.
-Especially smaller networks with high traffic could definitely benefit from those technologies [\nameref{sec:DCTCP}].
+Especially smaller networks with high traffic could benefit from those technologies [\nameref{sec:DCTCP}].
 
@@
 A lot of internet services run in data centers with large amounts of computers communicating with each other, resulting in a lot of traffic inside the data center.
-Optimizing those data centers networking capabilities can result in an increase in application productivity.
+Optimizing those data centers networking capabilities can increase application productivity.
 In 2010 a paper proposed \textit{DCTCP}, a variation of TCP using ECN to create a data transportation protocol, which especially takes hardware purposed for use in those environments into account \cite{10.1145/1851275.1851192}.\\
@@
 Besides, many data center switches have relatively small, shared buffers.
-Therefore, it is common that queue lengths stay high for a prolonged period of time, resulting in a collapse of throughput \cite{10.1145/1851275.1851192,10.1145/1592681.1592693}.\\
+Therefore, it is common that queue lengths stay high for a prolonged time, resulting in a collapse of throughput \cite{10.1145/1851275.1851192,10.1145/1592681.1592693}.\\
 The goal of DCTCP is to keep data center traffic constantly flowing, which means meeting the requirements applications take while keeping relatively small buffer sizes inside switches in mind.
 To achieve this, DCTCP implements ECN.
-Unlike TCP however, if a DCTCP sender registers congestion, the window size will be reduced by a fraction of the amount of marked packets it received.
+Unlike TCP however, if a DCTCP sender registers congestion, the window size will be reduced by a fraction of the number of marked packets it received.
 Therefore, excessive link underutilization can be prevented.
@@
 Therefore, fewer packets are sent through the network, but the TCP sender is still able to identify lost packets and register congestion \cite{10.1145/1851275.1851192}.
-\\However, being bandwidth greedy transmission protocol, data centers using DCTCP can see a performance drop when they additionally use other transmission protocols such as TCP or ECN.
+\\However, being a bandwidth greedy transmission protocol, data centers using DCTCP can see a performance drop when they additionally use other transmission protocols such as TCP or ECN.
 As DCTCP tries to utilize the links as much as possible, non-greedy protocols will be neglected and will see a drop in performance.

### 2020-06-15 paper/paper.tex
Multiple improvements

@@
 % paper title
-\title{The Origins of Congestion and\\Network Assisted End-to-End Congestion Control}
+\title{The Origins of Congestion and\\Network-Assisted End-to-End Congestion Control}
 
 % author
-%\author{\IEEEauthorblockN{Til Mohr}
-%\IEEEauthorblockA{RWTH Aachen University\\til.mohr@rwth-aachen.de}}
+\author{\IEEEauthorblockN{Til Mohr}
+\IEEEauthorblockA{RWTH Aachen University\\til.mohr@rwth-aachen.de}}
 
@@
 \begin{abstract}
-In this article we present the most common origins of network congestion.
-To tackle congestion, different types of algorithms will be compared, by first describing how each perceives congestion and later on how they adjust to those circumstances.
-This paper will also take a closer look at network assisted congestion control and a real-use implementation in data centers.
+This article gives an introduction to the most common origins of network congestion.
+Congestion lays in different types of delay, buffer sizes and their queue-management.
+While there are multiple approaches to minimize congestion, this paper will present and compare different types of algorithms used in transport protocols.
+TCP will be used to describe congestion avoidance, by first explaining how the protocol perceives congestion and later on how it adjusts to the current state of congestion.
+To compare to this approach, this paper will also take a closer look at network-assisted congestion control and a real-use implementation in data centers.
 \end{abstract}
@@
 \section{Introduction}
-Congestion was, is, and always will be a problem on the Internet, appearing in various forms with different symptoms causing minor, but also major challenges.
+Congestion was, is, and always will be a problem on the Internet, appearing in various forms with different symptoms causing minor but also major challenges.
 Most data sent over the internet is split up in smaller packets that all try to reach their destination the fastest, whether it is by taking the shortest or the least congested route.
@@
 However, when an internet node receives a packet while the outgoing connection still is in use, the packet has to be stored in a buffer until the outgoing connection freed up again.
-Therefore, those storage units are essential to every node on a network, but without proper management even well-sized buffers could lead to network problems like packet loss, jitter, and connection delay \cite{appenzeller2004sizing}.
-\\This paper will present origins of network congestion and analyze how different types of algorithms try to avoid and minimize congestion.
+Therefore, those storage units are essential to every node on a network but without proper management even well-sized buffers could lead to network problems like packet loss, jitter, and connection delay \cite{appenzeller2004sizing}.
+\\This paper will present the origins of network congestion [\autoref{sec:oc}] and analyze how TCP's congestion avoidance [\autoref{sec:ca}] protocols and network-assisted congestion control [\autoref{sec:cc}] try to avoid and minimize congestion.
 
 \section{Origins of Congestion}
-With the ever-growing amount of internet services not only data centers, but the whole internet in general faces new challenges with each packet sent.
+\label{sec:oc}
+With the ever-growing amount of internet services not only data centers but the whole internet in general faces new challenges with each packet sent.
 Therefore, a network-wide reduction of those complications is an important task to improve the quality of all internet services.
@@
 Typically, packets sent from an origin to a destination node are sent over various internet nodes, all of which could have a different bandwidth.
-Therefore, a packet send from \textbf{node 1} to \textbf{node n} over \textbf{nodes 2 to n-1} cannot be received at a faster rate than the lowest bandwidth on the path of the packet (bottleneck) [Figure \ref{fig:transmission delay}].\[Rate_{max} \leftarrow \min\{bandwidth_{i}\mid i\in\{1,...,n-1\}\}\]
+Therefore, a packet send from \textbf{node 1} to \textbf{node n} over \textbf{nodes 2 to n-1} cannot be received at a faster rate than the lowest bandwidth on the path of the packet (bottleneck) [\autoref{fig:transmission delay}].\[Rate_{max} \leftarrow \min\{bandwidth_{i}\mid i\in\{1,...,n-1\}\}\]
 
@@
 	\begin{center}
-	\begin{tikzpicture}[->, >=latex, node distance = 1cm and 2.5cm, semithick]
+	\begin{tikzpicture}[->, node distance = 1cm and 2.5cm]
 		% Nodes
@@
 When delay increases, the throughput of the network can decrease as packets will stay inside the network for a longer time.
-If the network however continues to receive new packets at a constant rate, more and more packets will have to be stored, resulting in congestion.
-Generally speaking, to tackle congestion, firstly delay can be reduced, and secondly the storages inside the network, the buffers, can be improved regarding efficient queue management and buffer size.
-Finally, transmission protocols can also react to congestion and adapt their send and data rate to recover from congestion [\nameref{sec:ca}] and even control congestion [\nameref{sec:cc}].
+However, if the network continues to receive new packets at a constant rate, more and more packets will have to be stored, resulting in congestion.
+Generally speaking, to tackle congestion, firstly delay can be reduced, and secondly, the storages inside the network, the buffers, can be improved regarding efficient queue management and buffer size.
+Finally, transmission protocols can also react to congestion and adapt their send and data rate to recover from congestion [\autoref{sec:ca}] and even control congestion [\autoref{sec:cc}].
 \\Whereas processing delay is almost not noticeable and transmission and inter-packet delay could be improved by upgrading cables, hardware, or transmission protocols, queuing delay has various causes and therefore is the  most difficult one to reduce.
 There are many different ways to implement buffers in a network node and the effectiveness of each implementation varies from where it is used.
-Traffic in residential network nodes for example differs from nodes used in large data centers in every aspect, thus hardware and software in those components should be adjusted accordingly.
+Traffic in residential network nodes, for example, differs from nodes used in large data centers in every aspect, thus hardware and software in those components should be adjusted accordingly.
 \\Sizing buffers is a complex task.
-On the one hand read and write times should be as short as possible to minimize additional delay.
-On the other hand, buffers should not be too small, as networks need to be robust for future changes and therefore nodes should store lots of data if needed, while also keeping in mind the requirements of congestion detection algorithms as described in section \nameref{sec:ca} in order to adjust for congestion appropriately \cite{staff2012bufferbloat}.
+On the one hand, read and write times should be as short as possible to minimize additional delay.
+On the other hand, buffers should not be too small, as networks need to be robust for future changes and therefore nodes should store lots of data if needed while also keeping in mind the requirements of congestion detection algorithms as described in \autoref{sec:ca} in order to adjust for congestion appropriately \cite{staff2012bufferbloat}.
 Unifying those goals is very challenging as it leads to increasing costs and physical space occupied by those buffers.
 Additionally, a 2004 paper by Appenzeller et al.
-\cite{appenzeller2004sizing} found that 99\% of buffers in network nodes, which were sized on a rule-of-thumb rule based on the TCP protocol \cite{villamizar1994high}, could be removed without losing, but rather gaining performance.
+\cite{appenzeller2004sizing} found that 99\% of buffers in network nodes, which were sized on a rule-of-thumb rule based on the TCP protocol \cite{villamizar1994high}, could be removed without losing but rather gaining performance.
 Clearly, balancing speed and size is difficult and thus buffer sizes should be adjusted to their area of use accordingly.
 \\This isn't the only major cause of queuing delay.
-Without proper buffer management even nodes with well-sized buffers will experience performance loss.
+Without proper buffer management, even nodes with well-sized buffers will experience performance loss.
 Proper queue management is able to compensate for the unpredictable fluidity of network traffic.
@@
 After this, the transmission rate cannot increase anymore, which leads to packets being buffered along their path to prevent them from being lost \cite{gettys2012bufferbloat}.
-If the incoming amount of traffic keeps to exceed the amount of outgoing traffic, those buffers will fill up (\textit{over-buffering}), which will eventually lead to \textit{bufferbloat} \cite{Allman13commentson,6886125}.
+If the incoming amount of traffic keeps exceeding the amount of outgoing traffic, those buffers will fill up (\textit{over-buffering}), which will eventually lead to \textit{bufferbloat} \cite{Allman13commentson,6886125}.
 
@@
 Bufferbloat refers to frequently full and excessively large buffers in internet nodes.
-This leads to partly unnecessary network delay and damages congestion avoidance algorithms, such as those found in classic TCP, which will be explained in the section \nameref{sec:ca} \cite{gettys2012bufferbloat,chen2014bufferbloat}.
+This leads to partly unnecessary network delay and damages congestion avoidance algorithms, such as those found in classic TCP, which will be explained in  \autoref{sec:ca} \cite{gettys2012bufferbloat,chen2014bufferbloat}.
 Buffering is needed to provide space to queued packets waiting for transmission resulting in a minimization of packet loss.
@@
 It is the interface between network components and computer programs.
-Each of those packets have different purposes, but underlay the same structure - the protocol header.
-The TCP header is split up into multiple segments [Figure \ref{fig:TCP_Header}]:
+Each of those packets have different purposes but underlay the same structure - the protocol header.
+The TCP header is split up into multiple segments [\autoref{fig:TCP_Header}]:
 \begin{figure}
@@
 Whenever a client sends a data packet to the server, the server waits until it has received enough data (specified by the window size), and then sends back an acknowledgment package, signaling the successful transmission of data.
-There are also other handshake operations used, for example for establishing or terminating a connection, however, those have no further importance to the congestion avoidance protocol \cite{1209197,jacobson1992tcp,huston2000tcp,jacobson1995congestion}.
+There are also other handshake operations used, for example, for establishing or terminating a connection, however, those have no further importance to the congestion avoidance protocol \cite{1209197,jacobson1992tcp,huston2000tcp,jacobson1995congestion}.
 
@@
 Both partners of a connection are both sender and receiver, as both transmit packets.
-In the following however, we will simplify this and specify one partner as the sender, the other as the receiver.
+In the following however, one partner will be specified as the sender, the other as the receiver.
 The sender keeps track of the time it takes from when a package sent until the acknowledgment is received.
@@
   \centering
-    \begin{tikzpicture}[y=.15cm, x=.45cm,font=\sffamily]
+    \begin{tikzpicture}[y=.15cm, x=.45cm]
     	% Ranges and dotted lines
@@
     	% Labels
-    	\node[below=0.1cm] at (x axis mid) {\scriptsize Transmission round};
-    	\node[rotate=90, above=0.1cm] at (y axis mid) {\scriptsize Congestion window size};
+    	\node[below=0.1cm] at (x axis mid) {Transmission round};
+    	\node[rotate=90, above=0.1cm] at (y axis mid) {Congestion window size};
     	% Plot
@@
     \end{tikzpicture}
-  \caption{TCP congestion avoidance protocols. Yellow represents the slow-start sequence, red and green symbolize detected congestion}
+  \caption{TCP congestion avoidance protocols. Yellow represents the slow-start sequence, red and green symbolize detected congestion. The data provided is symbolic.}
   \label{fig:tcp_congestion_avoidance_protocols}
@@
 	\centering
-		\begin{tikzpicture}[-, >=latex, node distance = 2.5cm, semithick]
+		\begin{tikzpicture}[->, node distance = 2.5cm]
 			%network
@@
 The \textit{slow-start} algorithm is used when a connection has just been established or when multiple packet losses were detected.
-The cwnd variable is set to one packet per estimated RTT, because no information about network congestion has been collected yet.
+The cwnd variable is set to one packet per estimated RTT because no information about network congestion has been collected yet.
 Whenever an acknowledgment packet is being received, cwnd will increase by one packet.
@@
 A multiplicative, exponential increase over time, like the slow-start algorithm isn't the best choice here.
-Increasing the number of packets sent drastically while cwnd still is high can challenge the networks capacity, resulting in heavy congestion.
+Increasing the number of packets sent drastically while cwnd still is high can challenge the network's capacity, resulting in heavy congestion.
 Therefore, the best option to raise cwnd to its limits is an additive increase \cite{jacobson1992tcp}:\[cwnd_{i+1} \leftarrow cwnd_{i} + u\] Van Jacobson proposes in his 1992 paper about TCP congestion avoidance that $d$ will be set to $\frac{1}{2}$ and $u$ will be set to $\frac{1}{cwnd_{i}}$.
@@
 
-\section{Network Assisted Congestion Control}
+\section{Network-Assisted Congestion Control}
 \label{sec:cc}
 The problem that comes with classic TCP is that the congestion detection algorithms are very implicit.
-Packet loss for example may have different causes other than network congestion, such as simple cable interference, however, the congestion avoidance algorithm will still interpret it as congestion, thus possibly resulting in unnecessary network delay.
+Packet loss, for example, may have different causes other than network congestion, such as simple cable interference, however, the congestion avoidance algorithm will still interpret it as congestion, thus possibly resulting in unnecessary network delay.
 Therefore, exactly determining whether the network is congested or not is crucial to keep a network stable.
 Here, the network layer comes into play.
-Because most communication should be hidden along the path of the packet, all network nodes will only communicate on the network layer (where IP lives), but not above (where TCP lives) \cite{ramakrishnan1999proposal}.\\
+Because most communication should be hidden along the path of the packet, all network nodes will only communicate on the network layer (where IP lives) but not above (where TCP lives) \cite{ramakrishnan1999proposal}.\\
 In 2001 a new method was standardized to explicitly notify TCP on congestion.
-\textit{Explicit Congestion Notification} (ECN) is an extension to TCP and IP, allowing end-to-end congestion notification, while also minimizing the number of packets dropped.
+\textit{Explicit Congestion Notification} (ECN) is an extension to TCP and IP, allowing end-to-end congestion notification while also minimizing the number of packets dropped.
 This is done by modifying the IP header to be able to carry one more bit - the \textit{Congestion Experienced} flag (CE) \cite{ramakrishnan1999proposal}.
-Instead of dropping packets occasionally after the number of queued packets exceed a threshold [\nameref{sec:P_C_marker}], all packets supporting ECN will now be marked (setting the CE bit in the IP header) [Figure \ref{fig:ecn}].
+Instead of dropping packets occasionally after the number of queued packets exceeds a threshold [\autoref{sec:P_C_marker}], all packets supporting ECN will now be marked (setting the CE bit in the IP header) [\autoref{fig:ecn}].
 Thus, congestion can be described explicitly resulting in better short- and long-term network performance.
@@
 Also, the need to retransmit packets would decrease drastically, as no packets are being dropped anymore to indicate congestion.
-Therefore, this central algorithm is also being called \textit{Congestion Control} \cite{ramakrishnan1999proposal,ramakrishnan2001addition}.
+Therefore, this central principal is also being called \textit{Congestion Control} \cite{ramakrishnan1999proposal,ramakrishnan2001addition}.
 
@@
 Although TCP is a very dominant protocol on the internet, also other transport protocols could benefit from end-to-end congestion notification.
-The paper "Congestion Control for High Bandwidth-Delay Product Networks"  generalizes the ECN proposal of 2001, creating a new protocol called \textit{eXplicit Congestion Protocol} (XCP), while also improving upon it further by introducing new packet control concepts \cite{katabi2002congestion,1498331}.
+The paper "Congestion Control for High Bandwidth-Delay Product Networks"  generalizes the ECN proposal of 2001, creating a new protocol called \textit{eXplicit Congestion Protocol} (XCP) while also improving upon it further by introducing new packet control concepts \cite{katabi2002congestion,1498331}.
 
@@
 Underneath the hood however, they are very different.
-A major difference between both protocols is that XCP does not only inform the sender on congestion, but does also provide information about the degree of congestion.
+A major difference between both protocols is that XCP does not only inform the sender on congestion but does also provide information about the degree of congestion.
 XCP can achieve this by including a \textit{congestion header} to each packet.
@@
 In addition, the congestion header includes one more field.
-This \textit{H\_feedback} field is also initialized by the sender, but is the only field which can be modified by nodes along the path \cite{katabi2002congestion,1498331}.
+This \textit{H\_feedback} field is also initialized by the sender but is the only field which can be modified by nodes along the path \cite{katabi2002congestion,1498331}.
 
@@
 Hence, the \textit{Fairness Controller} (FC) comes into play.
-As TCP, FC relies on the AIMD principle [\nameref{sec:AfC}].
+As TCP, FC relies on the AIMD principle [\autoref{sec:AfC}].
 If the EC calculated positive feedback, it will be split up between all packets equally.
@@
 
-\subsection{Evaluation of Network Assisted Congestion Control}
-As you can see, including the network layer to detect congestion can have great performance improvements.
+\subsection{Evaluation of Network-Assisted Congestion Control}
+As shown, including the network layer to detect congestion can have great performance improvements.
 'Understanding XCP: Equilibrium and Fairness' has shown through simulations that many network topologies could benefit from XCP as it clears queues in equilibrium, thus resulting in improved network performance under load \cite{1498331}.
@@
 \\However, all this performance comes at a cost.
-Since network nodes have to actively detect congestion, ECN, XCP, and other network assisted congestion control protocols rely on active queue management in those nodes.
+Since network nodes have to actively detect congestion, ECN, XCP, and other network-assisted congestion control protocols rely on active queue management in those nodes.
 Although there is some hardware being produced today that meet ECNs requirements, a large part of internet hardware isn't capable of doing so.
@@
 This is the reason you do not see these kinds of protocols in wide area networks (WAN) too often, let alone the internet.
-\\Nevertheless, whenever the implementation of XCP, ECN, or similar protocols is not a challenge, you can implement network assisted congestion control with ease and see great performance improvements.
-Especially smaller networks with high traffic could benefit from those technologies [\nameref{sec:DCTCP}].
+\\Nevertheless, whenever the implementation of XCP, ECN, or similar protocols is not a challenge, you can implement network-assisted congestion control with ease and see great performance improvements.
+Especially smaller networks with high traffic could benefit from those technologies [\autoref{sec:DCTCP}].
 
@@
 	\centering
-		\begin{tikzpicture}[-, >=latex, node distance = 2.5cm, semithick, auto]
+		\begin{tikzpicture}[node distance = 2.5cm]
 			\node[router, scale=0.8] (Router) {};
-			\node (Internet) [right = 1.5cm of Router, text width=2cm] {\scriptsize Outgoing Connection};
+			\node (Internet) [right = 1.5cm of Router, text width=2cm] {Outgoing Connection};
 			\node[switch, scale=0.8] (Sw1) [below left = 0.8cm and 1cm of Router] {};
 			\node[switch, scale=0.8] (Sw2) [below right = 0.8cm and 1cm of Router] {};
-			\node[server, scale=1] (S1) [below left = 0.8cm and 0.25cm of Sw1] {\scriptsize Server};
-			\node[server, scale=1] (S2) [below = 2.1cm of Router] {\scriptsize Server};
-			\node[server, scale=1] (S3) [below right = 0.8cm and 0.25cm of Sw2] {\scriptsize Server};
+			\node[server, scale=1] (S1) [below left = 0.8cm and 0.25cm of Sw1] {Server};
+			\node[server, scale=1] (S2) [below = 2.1cm of Router] {Server};
+			\node[server, scale=1] (S3) [below right = 0.8cm and 0.25cm of Sw2] {Server};
 
@@
 
-			\draw[decorate,decoration={brace,amplitude=10pt},xshift=-4pt,yshift=0pt] (S1.north -| S1.west) -- (Router.center -| S1.west) node [black,midway,xshift=-0.6cm,align=right] {\scriptsize High\\ \scriptsize'Fan-In'};
+			\draw[decorate,decoration={brace,amplitude=10pt},xshift=-4pt,yshift=0pt] (S1.north -| S1.west) -- (Router.center -| S1.west) node [black,midway,xshift=-1cm,align=right] {High\\ 'Fan-In'};
 		\end{tikzpicture}
-	\hspace{0.5cm}
+	\hspace{1cm}
 	\caption{Data Center High 'Fan-In'}
@@
 To understand where TCP incast originates, you fist need to understand how most data centers are built.
-A typical data center contains a set of racks, each holding a large number of servers, the switches, that connect those racks, and connecting links, that connect those switches to other parts of the data center or the internet [Figure \ref{fig:dc}].
+A typical data center contains a set of racks, each holding a large number of servers, the switches, that connect those racks, and connecting links, that connect those switches to other parts of the data center or the internet [\autoref{fig:dc}].
 This design pattern is also known as \textit{Partition/Aggregate} and is most commonly used in data centers.
@@
 Here, $m$ is a variable initialized while the DCTCP connection is being established and cannot be changed without renegotiation.
-Therefore, fewer packets are sent through the network, but the TCP sender is still able to identify lost packets and register congestion \cite{10.1145/1851275.1851192}.
+Therefore, fewer packets are sent through the network but the TCP sender is still able to identify lost packets and register congestion \cite{10.1145/1851275.1851192}.
 \\However, being a bandwidth greedy transmission protocol, data centers using DCTCP can see a performance drop when they additionally use other transmission protocols such as TCP or ECN.
@@
 \section{Conclusion}
-In this paper we have analyzed origins of congestion.
-Different factors play a role in network congestion, however especially the filling up of buffers at congested nodes will cause traffic delay.
-Further, we shine a light on the functionality of congestion avoidance algorithms, especially classic TCP.
-However, because those algorithms quite often do cause network stress rather than resolving it, we took a closer look at network assisted end-to-end congestion control.
-ECN, an addition to TCP, and XCP are two standardized protocols applying those principles, but haven't been very successful in wide area networks (WAN).
+This paper has analyzed the origins of congestion.
+Different factors play a role in network congestion, however especially full buffers at congested nodes will cause traffic delay.
+Further, we assess the functionality of congestion avoidance algorithms, especially classic TCP.
+However, since those algorithms quite often do cause network stress rather than resolving it, we took a closer look at network-assisted end-to-end congestion control.
+ECN, an addition to TCP, and XCP are two standardized protocols applying those principles but haven't been very successful in wide area networks (WAN).
 In smaller networks, the implementation of them is easier.

### 2020-07-05 paper/paper.tex
Update

@@
 This article gives an introduction to the most common origins of network congestion.
-Congestion lays in different types of delay, buffer sizes and their queue-management.
+Congestion lies in different types of delay, buffer sizes, and their queue-management.
 While there are multiple approaches to minimize congestion, this paper will present and compare different types of algorithms used in transport protocols.
@@
 \\Networks today consist of network nodes (routers, switches, ...) which all try to route internet packets as quickly as possible whilst keeping the required hardware to a minimum.
-Ideally there would be a steady flow of packets being transmitted at the maximum rate.
+Ideally, there would be a steady flow of packets being transmitted at the maximum rate.
 However, when an internet node receives a packet while the outgoing connection still is in use, the packet has to be stored in a buffer until the outgoing connection freed up again.
-Therefore, those storage units are essential to every node on a network but without proper management even well-sized buffers could lead to network problems like packet loss, jitter, and connection delay \cite{appenzeller2004sizing}.
+Therefore, those storage units are essential to every node on a network but without proper management, even well-sized buffers could lead to network problems like packet loss, jitter, and connection delay \cite{appenzeller2004sizing}.
 \\This paper will present the origins of network congestion [\autoref{sec:oc}] and analyze how TCP's congestion avoidance [\autoref{sec:ca}] protocols and network-assisted congestion control [\autoref{sec:cc}] try to avoid and minimize congestion.
@@
 With the ever-growing amount of internet services not only data centers but the whole internet in general faces new challenges with each packet sent.
-Therefore, a network-wide reduction of those complications is an important task to improve the quality of all internet services.
-\\Network delay can generally speaking be categorized into four different types of delay \cite{gettys2012bufferbloat}:
+Buffers are crucial storages inside network nodes.
+Their purpose is to store incoming packets while the outgoing connection is still in use.
+If a buffer however starts to fill up, other buffers in the same network will experience the same, thus creating congestion.
+\\Therefore, a network-wide reduction of those complications is an important task to improve the quality of all internet services.
+While delay can cause congestion by filling up buffers, buffers itself are also an origin of congestion.
+
+\subsection{Delay}
+\label{sec:delay}
+Network delay can generally speaking be categorized into four different types of delay \cite{gettys2012bufferbloat}:
+
+\begin{figure}
+	\begin{center}
+	\begin{tikzpicture}[->, node distance = 1cm and 2.5cm]
+		% Nodes
+		\node[state] (A)  			  [transition]	{1};
+		\node[state] (B) [right=of A] [transition]	{2};
+		\node[state] (C) [below=of B] [transition] 	{n-1};
+		\node[state] (D) [right=of C] [transition] 	{n};
+		%Arrows
+		\path	
+		(A) edge [above] node [transition] {$bandwidth_{1}$} (B)
+		(B) edge [dotted] (C)
+		(C) edge [above] node [transition] {$bandwidth_{n-1}$} (D)
+		;
+	\end{tikzpicture}
+	\end{center}
+	\caption{Path a packet takes through the network. Each circle represents a network node with $1$ being the sender and $n$ being the receiver.}
+	\label{fig:transmission delay}
+\end{figure}
 
@@
 \item
-Each node creates an intentional gap between two packets being sent.
-This pause is often necessary to allow the receiver to prepare for a new package, however, depending on the transmission protocol, other purposes may apply.
-This kind of delay is known as \textit{Inter-Packet delay}.
+The delay known as \textit{Inter-Packet delay} describes the gap each node intentionally creates between two packets being sent.
+This pause is often necessary to allow the receiver to prepare for the arrival of a new packet, however, depending on the transmission protocol, other purposes may also apply.
 
@@
 \item
-\textit{Queuing delay} is the time a package spends waiting to be processed or transmitted.
+\textit{Queuing delay} is the time a packet spends waiting to be processed or transmitted.
 \end{itemize}
-\begin{figure}
-	\begin{center}
-	\begin{tikzpicture}[->, node distance = 1cm and 2.5cm]
-		% Nodes
-		\node[state] (A)  			   	{1};
-		\node[state] (B) [right=of A] 	{2};
-		\node[state] (C) [below=of B] 	{n-1};
-		\node[state] (D) [right=of C] 	{n};
-		%Arrows
-		\path	
-		(A) edge [above] node {$bandwidth_{1}$} (B)
-		(B) edge [dotted] (C)
-		(C) edge [above] node {$bandwidth_{n-1}$} (D)
-		;
-	\end{tikzpicture}
-	\end{center}
-	\caption{Path a packet takes through the network. Each circle represents a network node with $1$ being the sender and $n$ being the receiver.}
-	\label{fig:transmission delay}
-\end{figure}
 
@@
 Traffic in residential network nodes, for example, differs from nodes used in large data centers in every aspect, thus hardware and software in those components should be adjusted accordingly.
-\\Sizing buffers is a complex task.
+
+\subsection{Buffers}
+\label{sec:buffers}
+
+Sizing buffers is a complex task.
 On the one hand, read and write times should be as short as possible to minimize additional delay.
-On the other hand, buffers should not be too small, as networks need to be robust for future changes and therefore nodes should store lots of data if needed while also keeping in mind the requirements of congestion detection algorithms as described in \autoref{sec:ca} in order to adjust for congestion appropriately \cite{staff2012bufferbloat}.
+On the other hand, buffers should not be too small, as networks need to be robust for future changes and bursts and therefore nodes should store lots of data if needed while also keeping in mind the requirements of congestion detection algorithms as described in \autoref{sec:ca} to adjust for congestion appropriately \cite{staff2012bufferbloat}.
 Unifying those goals is very challenging as it leads to increasing costs and physical space occupied by those buffers.
@@
 \cite{appenzeller2004sizing} found that 99\% of buffers in network nodes, which were sized on a rule-of-thumb rule based on the TCP protocol \cite{villamizar1994high}, could be removed without losing but rather gaining performance.
-Clearly, balancing speed and size is difficult and thus buffer sizes should be adjusted to their area of use accordingly.
+Balancing speed and size is difficult and thus buffer sizes should be adjusted to their area of use accordingly.
 \\This isn't the only major cause of queuing delay.
@@
 Proper queue management is able to compensate for the unpredictable fluidity of network traffic.
-It is a difficult task to buffer the right amount of packages to keep transmission flow consistent over time.
-As the number of packets being transmitted grows, the \textit{throughput} of the network node increases until packets are being transmitted at bottleneck rate.
+It is a difficult task to buffer the right amount of packets to keep transmission flow consistent over time.
+As the number of packets being transmitted grows, the \textit{throughput} of the network node increases until packets are being transmitted at the bottleneck rate.
 After this, the transmission rate cannot increase anymore, which leads to packets being buffered along their path to prevent them from being lost \cite{gettys2012bufferbloat}.
@@
 
-\subsection*{Bufferbloat}
+\subsubsection*{Bufferbloat}
 Bufferbloat refers to frequently full and excessively large buffers in internet nodes.
@@
 Buffering is needed to provide space to queued packets waiting for transmission resulting in a minimization of packet loss.
-In the past however, buffer sizes were smaller than today's thus less packets could be queued and more were dropped, signaling transmission protocols the presence of congestion in this node early in.
+In the past, however, buffer sizes were smaller than today's thus less packets could be queued and more were dropped, signaling transmission protocols the presence of congestion in this node early in.
 With memory getting cheaper over the years, buffers saw an increase in size.
@@
 \label{sec:ca}
-As already mentioned, congestion detection algorithms make up an important part of congestion avoidance protocols.
-\textit{TCP} is such a protocol, based on the transport layer, which adapts its send rate according to available network bandwidth \cite{1209197,jacobson1992tcp}.
-To fully understand, how TCP's congestion avoidance algorithm operates, one needs to understand how TCP works.
-In the following, TCP's congestion avoidance algorithm will be explained using the TCP Specification \cite{ietf-tcpm-rfc793bis-16,postel1981transmission}, also being referred to as \textit{classic TCP}.
-Note, there are also other TCP implementations improving upon it.
-The underlying concept however is the same.
-\\TCP lives on the transport layer.
-It communicates with the application layer via sockets, splits up data into packets (and vice versa), and transmits those packages over the network with the help of the internet protocol, thus TCP is often also referred to as \textit{TCP/IP}.
-It is the interface between network components and computer programs.
-Each of those packets have different purposes but underlay the same structure - the protocol header.
-The TCP header is split up into multiple segments [\autoref{fig:TCP_Header}]:
 \begin{figure}
@@
 \end{figure}
+As already mentioned, congestion detection algorithms make up an important part of congestion avoidance protocols.
+\textit{TCP} is such a protocol, based on the transport layer, which adapts its send rate according to available network bandwidth \cite{1209197,jacobson1992tcp}.
+To fully understand, how TCP's congestion avoidance algorithm operates, one needs to understand how TCP works.
+In the following, TCP's congestion avoidance algorithm will be explained using the TCP Specification \cite{ietf-tcpm-rfc793bis-16,postel1981transmission}, also being referred to as \textit{classic TCP}.
+Note, there are also other TCP implementations improving upon it.
+The underlying concept however is the same.
+\\TCP lives on the transport layer.
+It communicates with the application layer via sockets, splits up data into packets (and vice versa), and transmits those packets over the network with the help of the internet protocol, thus TCP is often also referred to as \textit{TCP/IP}.
+It is the interface between network components and computer programs.
+Each of those packets have different purposes but underlay the same structure - the protocol header.
+The TCP header is split up into multiple segments [\autoref{fig:TCP_Header}]:
+
 \begin{itemize}
@@
 TCP uses a handshake system.
-Whenever a client sends a data packet to the server, the server waits until it has received enough data (specified by the window size), and then sends back an acknowledgment package, signaling the successful transmission of data.
+Whenever a client sends a data packet to the server, the server waits until it has received enough data (specified by the window size), and then sends back an acknowledgment packet, signaling the successful transmission of data.
 There are also other handshake operations used, for example, for establishing or terminating a connection, however, those have no further importance to the congestion avoidance protocol \cite{1209197,jacobson1992tcp,huston2000tcp,jacobson1995congestion}.
@@
 Both partners of a connection are both sender and receiver, as both transmit packets.
-In the following however, one partner will be specified as the sender, the other as the receiver.
-The sender keeps track of the time it takes from when a package sent until the acknowledgment is received.
+In the following, however, one partner will be specified as the sender, the other as the receiver.
+The sender keeps track of the time it takes from when a packet sent until the acknowledgment is received.
 This is also known as \textit{round-trip time} (RTT).
@@
     	% Labels
-    	\node[below=0.1cm] at (x axis mid) {Transmission round};
-    	\node[rotate=90, above=0.1cm] at (y axis mid) {Congestion window size};
+    	\node[below=0.1cm] at (x axis mid) [transition] {Transmission round};
+    	\node[rotate=90, above=0.1cm] at (y axis mid) [transition] {Congestion window size};
     	% Plot
@@
 An additional mechanism of keeping a connection in equilibrium is, that each acknowledgment packet received by the sender opens the space for a new packet to be sent into the network \cite{jacobson1992tcp}.
-The principles TCP uses to adjust for congestion are based on \textit{Additive-Increase Multiplicative-Decrease} (AIMD) principle.
 
@@
 Also, the need to retransmit packets would decrease drastically, as no packets are being dropped anymore to indicate congestion.
-Therefore, this central principal is also being called \textit{Congestion Control} \cite{ramakrishnan1999proposal,ramakrishnan2001addition}.
+Therefore, this central principle is also being called \textit{Congestion Control} \cite{ramakrishnan1999proposal,ramakrishnan2001addition}.
 
@@
 The paper "Congestion Control for High Bandwidth-Delay Product Networks"  generalizes the ECN proposal of 2001, creating a new protocol called \textit{eXplicit Congestion Protocol} (XCP) while also improving upon it further by introducing new packet control concepts \cite{katabi2002congestion,1498331}.
+A major addition to this congestion control protocol is a feedback variable in the protocols header, which represents the amount of congestion in the network and is being computed along the data packets path.
 
@@
 While the sender maintains a \textit{cwnd} and \textit{RTT} variable, as with TCP, those are also being communicated to network nodes via the congestion header as \textit{H\_cwnd} and \textit{H\_rtt}.
-In addition, the congestion header includes one more field.
-This \textit{H\_feedback} field is also initialized by the sender but is the only field which can be modified by nodes along the path \cite{katabi2002congestion,1498331}.
+Additionally, the congestion header includes one more field.
+This \textit{H\_feedback} field is also initialized by the sender but is the only field that can be modified by nodes along the path \cite{katabi2002congestion,1498331}.
 
@@
 When $s$ is the packet size, the congestion window variable would be adjusted by the following formula:
-\[cwnd \leftarrow \max(cwnd + H_feedback, s)\]
+\[cwnd \leftarrow \max(cwnd + H\_feedback, s)\]
 In addition to responding to direct feedback, the XCP sender still needs to handle packet loss, though this is very similar to what TCP does \cite{katabi2002congestion}.
@@
 When the XCP routers link is underused, positive feedback should be sent.
-If the link is congested however, negative feedback should be sent in order to keep queue lengths small resulting in no packet drops \cite{katabi2002congestion,1498331}.\\
+If the link is congested however, negative feedback should be sent to keep queue lengths minimal resulting in no packet drops \cite{katabi2002congestion,1498331}.
+The exact computation of feedback is beyond the scope of this paper, is however explained into detail in a paper by Wydrowsky et al.
+\cite{1498331}.\\
 To achieve efficiency and equilibrium, this feedback will be allocated to single packets as \textit{H\_feedback}.
@@
 This is the reason you do not see these kinds of protocols in wide area networks (WAN) too often, let alone the internet.
-\\Nevertheless, whenever the implementation of XCP, ECN, or similar protocols is not a challenge, you can implement network-assisted congestion control with ease and see great performance improvements.
+\\Nevertheless, whenever the implementation of XCP, ECN, or similar protocols is not a challenge, can implement network-assisted congestion control with ease and see great performance improvements.
 Especially smaller networks with high traffic could benefit from those technologies [\autoref{sec:DCTCP}].
-
-\begin{figure}
-	\centering
-		\begin{tikzpicture}[node distance = 2.5cm]
-			\node[router, scale=0.8] (Router) {};
-			\node (Internet) [right = 1.5cm of Router, text width=2cm] {Outgoing Connection};
-			\node[switch, scale=0.8] (Sw1) [below left = 0.8cm and 1cm of Router] {};
-			\node[switch, scale=0.8] (Sw2) [below right = 0.8cm and 1cm of Router] {};
-			\node[server, scale=1] (S1) [below left = 0.8cm and 0.25cm of Sw1] {Server};
-			\node[server, scale=1] (S2) [below = 2.1cm of Router] {Server};
-			\node[server, scale=1] (S3) [below right = 0.8cm and 0.25cm of Sw2] {Server};
-
-			\draw[-] (Router) -- (Sw1);
-			\draw[-] (Router) -- (Sw2);
-			\draw[-, dotted] (Router) -- (Internet);
-			\draw[-] (Sw1) -- (S1);
-			\draw[-] (Sw1) -- (S2);
-			\draw[-] (Sw1) -- (S3);
-			\draw[-] (Sw2) -- (S1);
-			\draw[-] (Sw2) -- (S2);
-			\draw[-] (Sw2) -- (S3);
-
-			\draw[decorate,decoration={brace,amplitude=10pt},xshift=-4pt,yshift=0pt] (S1.north -| S1.west) -- (Router.center -| S1.west) node [black,midway,xshift=-1cm,align=right] {High\\ 'Fan-In'};
-		\end{tikzpicture}
-	\hspace{1cm}
-	\caption{Data Center High 'Fan-In'}
-	\label{fig:dc}
-\end{figure}
 
@@
 It describes a drastic reduction in application throughput.
-To understand where TCP incast originates, you fist need to understand how most data centers are built.
+To understand where TCP incast originates, one fist needs to understand how most data centers are built.
 A typical data center contains a set of racks, each holding a large number of servers, the switches, that connect those racks, and connecting links, that connect those switches to other parts of the data center or the internet [\autoref{fig:dc}].
@@
 It however results in a high 'fan-in', as multiple servers (leaves) are connected to just one link (root) \cite{10.1145/1851275.1851192,10.1145/1592681.1592693}.
-In addition, most applications require high bandwidth, low latency connections, also enabling them to send many parallel requests to other servers.
+Additionally, most applications require high bandwidth, low latency connections, also enabling them to send many parallel requests to other servers.
 Combining those requirements with the design pattern, it becomes clear, that those switches are being challenged.
 Besides, many data center switches have relatively small, shared buffers.
-Therefore, it is common that queue lengths stay high for a prolonged time, resulting in a collapse of throughput \cite{10.1145/1851275.1851192,10.1145/1592681.1592693}.\\
+Therefore, it is common that many packets are being transmitted at the same time (\textit{bursty retransmissions}), resulting in a collapse of throughput \cite{10.1145/1851275.1851192,10.1145/1592681.1592693}.\\
 The goal of DCTCP is to keep data center traffic constantly flowing, which means meeting the requirements applications take while keeping relatively small buffer sizes inside switches in mind.
@@
 The paper proposing DCTCP has found, that their protocol increases traffic by 110\% while using 90\% less buffer space compared to TCP \cite{10.1145/1851275.1851192}.
+Therefore it is unsurprising that DCTCP is very popular in data centers.
+
+\begin{figure}
+	\centering
+		\begin{tikzpicture}[node distance = 2.5cm]
+			\node[router, scale=0.8] (Router) {};
+			\node (Internet) [right = 1.5cm of Router, text width=2cm] {Outgoing Connection};
+			\node[switch, scale=0.8] (Sw1) [below left = 0.8cm and 1cm of Router] {};
+			\node[switch, scale=0.8] (Sw2) [below right = 0.8cm and 1cm of Router] {};
+			\node[server, scale=1] (S1) [below left = 0.8cm and 0.25cm of Sw1] [transition] {Server};
+			\node[server, scale=1] (S2) [below = 2.1cm of Router] [transition] {Server};
+			\node[server, scale=1] (S3) [below right = 0.8cm and 0.25cm of Sw2] [transition] {Server};
+
+			\draw[-] (Router) -- (Sw1);
+			\draw[-] (Router) -- (Sw2);
+			\draw[-, dotted] (Router) -- (Internet);
+			\draw[-] (Sw1) -- (S1);
+			\draw[-] (Sw1) -- (S2);
+			\draw[-] (Sw1) -- (S3);
+			\draw[-] (Sw2) -- (S1);
+			\draw[-] (Sw2) -- (S2);
+			\draw[-] (Sw2) -- (S3);
+
+			\draw[decorate,decoration={brace,amplitude=10pt},xshift=-4pt,yshift=0pt] (S1.north -| S1.west) -- (Router.center -| S1.west) node [black,midway,xshift=-1cm,align=right] [transition] {High\\ 'Fan-In'};
+		\end{tikzpicture}
+	\hspace{1cm}
+	\caption{Data Center High 'Fan-In'}
+	\label{fig:dc}
+\end{figure}
 
@@
 This paper has analyzed the origins of congestion.
-Different factors play a role in network congestion, however especially full buffers at congested nodes will cause traffic delay.
+Different factors play a role in network congestion, however, especially full buffers at congested nodes will cause traffic delay.
 Further, we assess the functionality of congestion avoidance algorithms, especially classic TCP.

### 2020-07-05 slides/slides.tex
Multiple improvements
Update

@@
 \usepackage{marvosym}
+\usepackage{latexsym}
+\usepackage{amsfonts,amsmath,amssymb,amsthm} 
 \usepackage{wasysym}
-\usepackage{latexsym}
-\usepackage{amssymb}
 \usepackage{hyperref}
+\usepackage{mathtools}
 \tolerance=1000
@@
 \usepackage[iso]{isodate}
+\usepackage[backend=biber]{biblatex}
+
+\usepackage{bytefield}						% Bytefield
+\usepackage{moeptikz}						% tikz-packages for networking symbols
+\usepackage{tikz}							% tikz-figures
+\usetikzlibrary{decorations.pathreplacing,positioning,arrows,automata}
+\tikzstyle{line} = [draw,thick,-latex]
+\tikzstyle{transition} = [font=\small]
 
@@
 \usetheme{Comsys}
+\addbibresource{references.bib}
 
-\title[Cool]{The Origins of Congestion\\and Network assisted end-to-end\\Congestion Control}
+% data for figure
+\begin{filecontents}{tcp.data}
+#Transmission round		Congestion Window Size
+1		1
+2		2
+3		4
+4		8
+5		16
+6		32
+7		16
+8		17
+9		18
+10		19
+11		20
+12		10
+13		5
+14		6
+15		7
+\end{filecontents}
+
+\title[Cool]{The Origins of Congestion\\and Network-assisted end-to-end\\Congestion Control}
 \author{Til Mohr}
 \authorextra{Advisor: Jan Rüth}
-\subtitle{Proseminar}
+%\subtitle{Proseminar}
 \date{\today}
@@
 
-\section{First Topic}
-\label{sec:first_topic}
-\frame\sectionpage
+\section{Introduction}
+\label{sec:intro}
+\AtBeginSection[]
+{
+  \begin{frame}
+	\frametitle{Outline}
+    \tableofcontents[currentsection]
+  \end{frame}
+}
 
-\begin{frame}{Blocks}
-	\begin{cols}{0.5,auto}
-		\begin{block}{Block with Title}
-			Content of block with title
-		\end{block}
-		\begin{whiteblock}{White Block}
-			Just a blue frame around the text
-		\end{whiteblock}
-		\begin{exampleblock}{This is an Example Block}
-			Example Block
-		\end{exampleblock}
-		\begin{alertblock}{This is an Alert Block}
-			Alert Block
+\begin{frame}{Introduction}
+  \nocite{*}
+\begin{columns}
+	\begin{column}{.5\textwidth}
+		\only<1>{
+		\begin{figure}
+		\centering
+		\begin{tikzpicture}
+			\node [client, scale=0.7] (Client1) {};
+			\node [router, scale=0.7] (Router1) [below right = 0.35cm and 2cm of Client1] {};
+			\node [server, scale=0.7] (Server1) [above right = 0.35cm and 2cm of Router1] {};
+			\node [router, scale=0.7] (Router2) [below right = 2cm and 0.75cm of Client1] {};
+			\node [router, scale=0.7] (Router3) [right = 2.25cm of Router2] {};
+			\node [server, scale=0.7] (Server2) [below left = 1.75cm and -0.1cm of Router2] {};
+			\node [client, scale=0.7] (Client2) [below right = 1.75cm and 0.1cm of Router3] {};
+
+			\draw[-] (Client1) -- (Router2);
+			\draw[-] (Router1) -- (Router2);
+			\draw[-] (Server1) -- (Router1);
+			\draw[-] (Router1) -- (Router3);
+			\draw[-] (Router2) -- (Server2);
+			\draw[-] (Router3) -- (Server2);
+			\draw[-] (Router3) -- (Client2);
+		\end{tikzpicture}
+		\end{figure}
+		}
+
+		\only<2->{
+		\begin{figure}
+		\centering
+		\begin{tikzpicture}
+			\node[state, scale=0.7] (Client1) {};
+			\node[state, scale=0.7] (Router1) [below right = 0.35cm and 2cm of Client1] {};
+			\node[state, scale=0.7] (Server1) [above right = 0.35cm and 2cm of Router1] {};
+			\node[state, scale=0.7] (Router2) [below right = 2cm and 0.75cm of Client1] {};
+			\node[state, scale=0.7] (Router3) [right = 2.25cm of Router2] {};
+			\node[state, scale=0.7] (Server2) [below left = 1.75cm and -0.1cm of Router2] {};
+			\node[state, scale=0.7] (Client2) [below right = 1.75cm and 0.1cm of Router3] {};
+
+			\only<2>{
+			\draw[-] (Client1) -- (Router2);
+			\draw[-] (Router1) -- (Router2);
+			\draw[-] (Server1) -- (Router1);
+			\draw[-] (Router1) -- (Router3);
+			\draw[-] (Router2) -- (Server2);
+			\draw[-] (Router3) -- (Server2);
+			\draw[-] (Router3) -- (Client2);
+			}
+			\only<3->{
+			\path
+				(Client1) edge [left] node [transition] {10} (Router2)
+				(Router1) edge [above left] node [transition] {3} (Router2)
+				(Server1) edge [above] node [transition] {25} (Router1)
+				(Router1) edge [below left] node [transition] {30} (Router3)
+				(Router2) edge [left] node [transition] {9} (Server2)
+				(Router3) edge [below right] node [transition] {28} (Server2)
+				(Router3) edge [left] node [transition] {13} (Client2)
+			;
+			}
+		\end{tikzpicture}
+		\end{figure}
+		}
+	\end{column}
+
+	\only<1-2>{\begin{column}{.45\textwidth}\end{column}}
+	\only<3->{
+	\begin{column}{.45\textwidth}
+		\begin{itemize}
+		\item Links have limitations.
+		\item<4-> Data is split up into smaller packets.
+		\item<5-> Network nodes contain buffers storing packets.
+		\end{itemize}
+		\only<5->{
+		\begin{alertblock}{}
+		Origins of Congestion are versatile.
 		\end{alertblock}
-	\col
-		\begin{block}{}
-			Content of block without title
-		\end{block}
-		\begin{whiteblock}{}
-			Without title
-		\end{whiteblock}
-		\begin{grayblock}{}
-			Grayed out block
-		\end{grayblock}
-		\begin{exampleblock}{}
-			Example Block without Title
-		\end{exampleblock}
-		\begin{redblock}{}
-			Red Block without Title
-		\end{redblock}
-	\end{cols}
-	\pushdown
-	\begin{alertblock}{}
-		\begin{center}Don't overload your slides, beware of the shadow!\end{center}
-	\end{alertblock}
+		}
+	\end{column}
+	}
+\end{columns}
 \end{frame}
 
-\begin{frame}{Itemize}
-	\begin{cols}{1,2*1}
+\section{Origins of Congestion}
+\label{sec:oc}
+\AtBeginSection[]
+{
+  \begin{frame}
+	\frametitle{Outline}
+    \tableofcontents[currentsection]
+  \end{frame}
+}
+
+\subsection{Buffers}
+\label{sec:buffers}
+\begin{frame}{Buffers}
+~\\
+\begin{block}{Purpose}
+Buffers store packets in network nodes waiting for transmission.
+\end{block}
+~\\
+\only<2->{
+Sizing Buffers:
 \begin{itemize}
-		\item item
-		\item item
-		\item []
-		\item item
+\item Buffer size should be large enough to handle packet bursts.
+\item Too large buffers result in an increase of queuing delay and \textit{over-buffering}.
+\end{itemize}
+}
+~\\
+\only<3->{
+\begin{enumerate}
+\item[$\Rightarrow$] Buffers should be designed according to their application environment.
+\end{enumerate}
+}
+\end{frame}
+
+\begin{frame}{Bufferbloat}
 \begin{itemize}
-			\item subitem
+\item Refers to frequently full, excessively large buffers.
+\item<2-> Queued packets will not be dropped, resulting in:
+\only<2->{
 	\begin{itemize}
-				\item subsubitem
+	\item Unnecessary network delay (\textit{latency}) and packet delay fluctuation (\textit{jitter}).
+	\item Increased congestion, as TCP is not being notified of congestion.
+	\item Decreases network throughput.
 	\end{itemize}
+}
+\item<3-> With memory getting cheaper over the years, buffers saw an increase in size.
 \end{itemize}
-		\end{itemize}
-	\col
-		\begin{enumerate}
-		\item Enumerated item
-		\item Enumerated item
-		\item[]
-		\item Enumerated item
-		\item Enumerated item
-		\end{enumerate}
-	\end{cols}
+\end{frame}
 
-	\vspace{1em}
+\subsection{Delay}
+\label{sec:delay}
+\begin{frame}{Delay}
+~\\
+\centering
+\begin{figure}
+\centering
+	\begin{tikzpicture}
+		\node (net1) {};
+		\node[router, scale=0.8] (Router1) [right = 1.5cm of net1] {};
+		\node[router, scale=0.8] (Router2) [right = 3.2cm of Router1] {};
+		\node[router, scale=0.8] (Router3) [right = 3.2cm of Router2] {};
+		\node (net2) [right = 1.5cm of Router3] {};
 
-	\begin{cols}{auto}
-		\begin{block}{Block with Items}
+		\draw[-, dotted] (net1) -- (Router1);
+		\draw[-] (Router1) -- (Router2);
+		\draw[-] (Router2) -- (Router3);
+		\draw[-, dotted] (Router3) -- (net2);
+	\end{tikzpicture}
+\end{figure}
+~\\
+\begin{minipage}{\textwidth}
+\begin{columns}
+%Content
+\only<2->{
+	\begin{column}{0.5\textwidth}
+		\begin{block}{Transmission Delay}
+			Depends on bandwidth and length of the connection.
+		\end{block}
+	\end{column}
+}
+\only<3->{
+	\begin{column}{0.5\textwidth}
+		\begin{block}{Processing Delay}
+			Time it takes to process a single packet.
+		\end{block}
+	\end{column}
+}
+%Placeholders
+\only<1>{\begin{column}{0.5\textwidth}\end{column}}
+\only<1-2>{\begin{column}{0.5\textwidth}\end{column}}
+\end{columns}
+\end{minipage}
+
+\begin{minipage}{\textwidth}
+\begin{columns}
+\only<4->{
+	\begin{column}{0.5\textwidth}
+		\begin{block}{Inter-Packet Delay}
+			Allows the receiver to prepare for transmission.
+		\end{block}
+	\end{column}
+}
+\only<5->{
+	\begin{column}{0.5\textwidth}
+		\begin{block}{Queuing Delay}
+			Time a packet waits in the buffer waiting for transmission.
+		\end{block}
+	\end{column}
+}
+\only<1-3>{\begin{column}{0.5\textwidth}\end{column}}
+\only<1-4>{\begin{column}{0.5\textwidth}\end{column}}
+\end{columns}
+\end{minipage}
+\end{frame}
+
+\subsection{Overview over Congestion}
+\label{sec:ooc}
+
+\begin{frame}{Overview over Congestion}
+\begin{columns}
+\begin{column}{0.5\textwidth}
+	\begin{block}{Buffers}
 		\begin{itemize}
-			\item item
-				\begin{itemize}
-				\item subitem
-					\begin{itemize}
-					\item subsubitem
-					\end{itemize}
-				\end{itemize}
+		\only<2->{\pgfsetfillopacity{0.2}}
+		\item Buffer size
+		\only<2->{\pgfsetfillopacity{1}}
+		\item Buffer management
 		\end{itemize}
 	\end{block}
-		\begin{block}{}
-			\begin{enumerate}
-			\item item
-			\item item
-			\end{enumerate}
+\end{column}
+
+\begin{column}{0.5\textwidth}
+	\begin{block}{Delay}
+		\begin{itemize}
+		\only<2->{\pgfsetfillopacity{0.2}}
+		\item Transmission Delay
+		\item Processing Delay
+		\item Inter-Packet Delay
+		\only<2->{\pgfsetfillopacity{1}}
+		\item Queuing Delay
+		\end{itemize}
 	\end{block}
-	\col
-		\begin{exampleblock}{}
+\end{column}
+\end{columns}
+~\\~\\
+\only<2->{
 \begin{itemize}
-			\item item
-				\begin{itemize}
-				\item subitem
-					\begin{itemize}
-					\item subsubitem
+\item Congestion control algorithms can improve queuing delay, number of queued packets and buffer management.
 \end{itemize}
-				\end{itemize}
-			\end{itemize}
-		\end{exampleblock}
-		\begin{alertblock}{}
-			\begin{itemize}
-			\item item
-				\begin{itemize}
-				\item subitem
-					\begin{itemize}
-					\item subsubitem
-					\end{itemize}
-				\end{itemize}
-			\end{itemize}
-		\end{alertblock}
-	\col
-		\begin{whiteblock}{}
-			\begin{itemize}
-			\item item
-				\begin{itemize}
-				\item subitem
-					\begin{itemize}
-					\item subsubitem
-					\end{itemize}
-				\end{itemize}
-			\end{itemize}
-		\end{whiteblock}
-		\begin{grayblock}{}
-			\begin{itemize}
-			\item item
-				\begin{itemize}
-				\item subitem
-					\begin{itemize}
-					\item subsubitem
-					\end{itemize}
-				\end{itemize}
-			\end{itemize}
-		\end{grayblock}
-	\end{cols}
+}
 \end{frame}
 
-\begin{frame}{Illustrative Pictures right of Main Text}
+\section{TCP Congestion Avoidance}
+\label{sec:ca}
+\AtBeginSection[]
+{
+  \begin{frame}
+	\frametitle{Outline}
+    \tableofcontents[currentsection]
+  \end{frame}
+}
+
+\begin{frame}{TCP}
+~\\
+\begin{block}{Transmission Control Protocol}
 \begin{itemize}
-	\item A picture can be inserted which is right aligned
+\item Lives on transport layer, communicates with application layer via sockets ($\rightarrow$ ports) and uses IP as internet protocol ($\rightarrow$ TCP/IP).
+\item Ensures successful transmission of all data.
+\end{itemize}
+\end{block}
 \begin{itemize}
-		\item It is vertically aligned with the position\ilgraphics[height=2em]{logos/rwth}
-		\\it was inserted in the source
+\item<2-> Transmitted data needs to be acknowledged.
+\item<2-> Adjusts to the current state of congestion.
+\item<3-> Many different implementations exist.
 \end{itemize}
-	\item There are two variants
+\end{frame}
+
+\begin{frame}[fragile]{TCP Header}
+\centering
+\begin{figure}
+\centering
+\begin{bytefield}[bitheight=2.2\baselineskip, bitwidth=0.635\baselineskip]{32}
+\bitbox{16}{Source Port} & \bitbox{16}{Destination Port} \\
+\bitbox{32}{Sequence Number} \\
+\bitbox{32}{Acknowledgment Number} \\
+\bitbox{4}{Data Offset} & \bitbox{4}{Rsrvd} & \bitbox{1}{\tiny C\\W\\R} & \bitbox{1}{\tiny E\\C\\E} & \bitbox{1}{\tiny U\\R\\G} & \bitbox{1}{\tiny A\\C\\K} & \bitbox{1}{\tiny P\\S\\H} & \bitbox{1}{\tiny R\\S\\T} & \bitbox{1}{\tiny S\\Y\\N} & \bitbox{1}{\tiny F\\I\\N} & \bitbox{16}{Window Size} \\
+\bitbox{16}{Checksum} & \bitbox{16}{Urgent Pointer} \\
+\bitbox{32}{Options} \\
+\bitbox{32}{Data}
+\end{bytefield}
+\end{figure}
+\end{frame}
+
+\subsection*{Perceiving Congestion}
+\label{sec:pc}
+
+\begin{frame}{Perceiving Congestion}
+\begin{columns}
+\begin{column}{0.5\textwidth}
+\begin{block}{Round-trip Time (\textit{RTT})}
+Describes the amount of time between sending a data packet and receiving its acknowledgment packet.
+\end{block}
+\end{column}
+\begin{column}{0.5\textwidth}
+\begin{block}{Retransmission Timeout (\textit{RTO})}
+The maximum amount of time the sender will wait for the acknowledgment of a packet.
+\end{block}
+\end{column}
+\end{columns}
+~\\
+\only<2->{
 \begin{itemize}
-		\item A variant based on \textit{{\textbackslash}includegraphics} which takes the same arguments as \textit{{\textbackslash}includegraphics}
-		\item An environment for any content
-		\begin{illustration}
+\item Classic TCP perceives congestion by detecting packet loss.
+\item Other implementations also perceive congestion by determining variations in RTT over time.
+\end{itemize}
+}
+\end{frame}
+
+\begin{frame}[fragile]{Perceiving Packet Loss - Selective Acknowledgments}
+\centering
+\begin{figure}
+\centering
 \begin{tikzpicture}
-				\draw[draw=red] (0,0) circle (1em);
-				\draw[draw=blue] (0,0) rectangle(2em,2em);
+	\node (Sender) {Sender};
+	\node (Receiver) [right = 5cm of Sender] {Receiver};
+	\node (s1) [below = 0.5 of Sender] {};
+	\node (s2) [below = 0.35cm of s1] {};
+	\node (s3) [below = 0.35cm of s2] {};
+	\node (r1) [below = 1cm of Receiver] {};
+	\node (r2) [below = 0.35cm of r1] {};
+	\node (r3) [below = 0.35cm of r2] {};
+	\node (r4) [below = 1cm of r3] {};
+	\node (r5) [below = 0.35cm of r4] {};
+	\node (r6) [below = 0.35cm of r5] {};
+	\node (s4) [below = 2cm of s3]  {};
+	\node (s5) [below = 0.35cm of s4] {};
+	\node (s6) [below = 0.35cm of s5] {};
+	\node (Sender2) [below = 1cm of s6] {};
+	\node (Receiver2) [below = 1.55cm of r6] {};
+	\only<2->{\node (loss) [below right = -0.2cm and 3cm of s2] {\lightning};}
+
+	\only<1->{\path (s1) edge [above] node {\tiny A} (r1);}
+	\only<2->{\path (s2) edge [dashed, above] node {\tiny B} (loss);}
+	\only<3->{\path (s3) edge [below] node {\tiny C} (r3);}
+	\only<4->{\path (r4) edge [above] node {\tiny A} (s4);}
+	\only<5->{\path (r6) edge [above] node {\tiny A} (s6);}
+
+	\draw[-] (Sender) -- (Sender2);
+	\draw[-] (Receiver) -- (Receiver2);
 \end{tikzpicture}
-		\end{illustration}
+\end{figure}
+\end{frame}
+
+\subsection*{Adjusting for Congestion}
+\label{sec:ac}
+\begin{frame}{Adjusting for Congestion}
+~\\
+\begin{block}{Goal}
+\begin{itemize}
+\item Do not fill up buffers in the network.
+\item Achieve and hold maximum throughput.
+\item Adapt to the current state of congestion as quickly as possible.
+\end{itemize}
+\end{block}
+\only<2->{
+\begin{itemize}
+\item CA algorithms are based on Additive-Increase Multiplicative-Decrease principle.
+\item Variable $cwnd$ (congestion window) describes the maximum amount of data sent before ACKs are needed.
+\end{itemize}
+}
+\end{frame}
+
+\begin{frame}{Adaption - Additive Increase}
+~\\
+\begin{block}{Formula}
+$cwnd \coloneqq cwnd + u$
+\end{block}
+\begin{itemize}
+\item Increases the data rate slowly to reach the networks maximum throughput.
+\item Prevents the rapid filling of buffers.
+\item Proposed value for $u$: $\frac{1}{cwnd}$
+\end{itemize}
+\end{frame}
+
+\begin{frame}{Adaption - Multiplicative Decrease}
+~\\
+\begin{block}{Formula}
+$cwnd \coloneqq d \cdot cwnd, 0<d<1$
+\end{block}
+\begin{itemize}
+\item Decreases the data rate drastically.
+\item Allows the nodes to recover and buffer lengths to decrease.
+\item Proposed value for $d$: $\frac{1}{2}$
+\end{itemize}
+\end{frame}
+
+\begin{frame}{Adaption - Slow-Start}
+~\\
+\begin{block}{Formula}
+$cwnd \coloneqq cwnd + 1$ for each received ACK packet.
+\\$\Rightarrow$ Exponential increase over time.
+\end{block}
+\begin{itemize}
+\item Drastically increases the protocols performance.
+\item Used after connection is established or after multiple packet losses have been detected.
+\item Switches off after a threshold has been exceeded or congestion has been perceived.
+\end{itemize}
+\end{frame}
+
+\begin{frame}{Visualization of TCPs Congestion Avoidance}
+\centering
+\begin{figure}
+  \centering
+    \begin{tikzpicture}[y=.2cm, x=.5cm]
+    	% Ranges and dotted lines
+    	\fill[yellow!10]	(0,0) rectangle (6,35);
+    	\fill[red!10] 		(6,0) rectangle (7,35);
+    	\fill[green!10] 	(7,0) rectangle (11,35);
+    	\fill[red!10]		(11,0) rectangle (13,35);
+    	\fill[green!10]		(13,0) rectangle (15.3,35);
+    	\draw[dotted] 		(6,0) -- (6,35);
+    	\draw[dotted] 		(7,0) -- (7,35);
+    	\draw[dotted] 		(11,0) -- (11,35);
+    	\draw[dotted] 		(13,0) -- (13,35);
+    	% Axis
+    	\draw[->] 			(0,0) -- coordinate (x axis mid) (17,0);
+    	\draw[->] 			(0,0) -- coordinate (y axis mid) (0,35);
+    	% Labels
+    	\node[below=0.1cm] at (x axis mid) [transition] {Transmission round};
+    	\node[rotate=90, above=0.1cm] at (y axis mid) [transition] {Congestion window size};
+    	% Plot
+    	\draw plot[mark=*]
+    		file {tcp.data};
+    \end{tikzpicture}
+\end{figure}
+\end{frame}
+
+\section{Network-assisted Congestion Control}
+\label{sec:cc}
+\AtBeginSection[]
+{
+  \begin{frame}
+	\frametitle{Outline}
+    \tableofcontents[currentsection]
+  \end{frame}
+}
+
+\begin{frame}{Network-assisted Congestion Control}
+\begin{itemize}
+\item Congestion Detection algorithms are very implicit:
+	\begin{itemize}
+	\item Packet loss may also have other causes
+	\item Methods are inaccurate, and do not inform about the degree of congestion.
+	\end{itemize}
+\item[$\Rightarrow$] Creates unnecessary network delay.
+~\\~\\
+\item<2-> Explicit congestion notification is crucial to keep the network stable.
+\item[$\Rightarrow$]<2-> Modify the IP header to include a \textit{Congestion Experienced} Flag (CE).
+\item<3-> \textit{Explicit congestion notification} (ECN), an extention to TCP/IP, implements this.
+\end{itemize}
+\end{frame}
+
+\begin{frame}{Principle of Network-assisted Congestion Control}
+\begin{center}
+\begin{figure}
+	\centering
+		\begin{tikzpicture}[->, node distance = 2.1cm]
+			%network
+			\node[client, scale=0.7] (H1) {};
+			\node[router, scale=0.7] (R1) [right=of H1] {};
+			\node[router, scale=0.7] (R2) [right=of R1] {};
+			\node[router, scale=0.7] (R3) [right=of R2] {};
+			\node[client, scale=0.7] (H2) [right=of R3] {};
+			\path [line] (H1) -- (R1);
+			\path [line] (R1) -- (R2);
+			\path [line] (R2) -- (R3);
+			\path [line] (R3) -- (H2);
+			\path [line, dotted] (H2) -- ++(0,-35pt) -| (H1);
+			\node[below=0.9cm of R2] (text) [transition] {Congested Signal};
+
+			%packets
+			\node[messageclosed,scale=0.5, fill=black!10] (P1) [above right=0.2cm and 1cm of H1] {};
+			\node[messageclosed,scale=0.5, fill=black!50] (P2) [above left=0.7cm and 0.3cm of R2] {};
+			\node[messageclosed,scale=0.5, fill=black!10] (P3) [above left=0.4cm and 0.2cm of R2] {};
+			\node[messageclosed,scale=0.5, fill=black!10] (P4) [above left=0.1cm and 0.1cm of R2] {};
+			\node[messageclosed,scale=0.5, fill=black!50] (P5) [above right=0.2cm and 1cm of R3] {};
+			\node[above left = 1.5cm and 0cm of R1] (P) [transition] {Packet};
+			\node[above right = 1.5cm and -0.3cm of R2] (CN) [transition] {Congested Node};
+			\node[above left = 1.5cm and -0.4cm of H2] (MP) [transition] {Marked Packet};
+			\path [line] (P) -- (P1);
+			\path [line] (CN) -- (R2);
+			\path [line] (MP) -- (P5);
+		\end{tikzpicture}
+\end{figure}
+\end{center}
+\end{frame}
+
+\subsection{XCP}
+\label{sec:xcp}
+\begin{frame}{XCP}
+~\\
+\begin{block}{XCP}
+\textit{eXplicit Congestion Protocol} generalizes the ECN proposal.
+\end{block}
+\begin{itemize}
+\item Improves upon ECN by introducing new packet control concepts.
+\item Additionally gives feedback about the degree of congestion.
+\end{itemize}
+\only<2->{
+\begin{block}{XCP Header}
+\begin{itemize}
+\item $H\_cwnd$
+\item $H\_rtt$
+\item $H\_feedback$
+\end{itemize}
+\end{block}
+}
+\end{frame}
+
+\begin{frame}{XCP Sender}
+\begin{itemize}
+\item As a TCP Sender, the XCP sender also keeps track of $cwnd$ and $RTT$ and deals with packet loss (retransmission).
+~\\~\\
+\item On data packet depature, the XCP sender will:
+	\begin{itemize}
+	\item $H\_cwnd \coloneqq cwnd$
+	\item $H\_rtt \coloneqq RTT$
+	\item $H\_feedback \coloneqq \frac{r \cdot RTT - cwnd}{cwnd}$, where $r$ is the desired data rate.
+	\end{itemize}
+\item[$\Rightarrow$] The desired data rate can be reached in just one RTT.
+~\\~\\
+\item On ACK packet arrival, the XCP sender will adjust $cwnd$ according to the received feedback:
+	\begin{itemize}
+	\item $cwnd \coloneqq max(cwnd + H\_feedback, s)$, where $s$ is the packet size.
 	\end{itemize}
@@
 
-\singlemessageframe{This Frame contains\\only a single Message!}
+\begin{frame}{XCP Router}
+\begin{columns}
+\begin{column}{.5\textwidth}
+\begin{block}{Efficiency Controller (EC)}
+\begin{itemize}
+\item Maximizes link utilization.
+\item Minimizes queue lengths and packet drops.
+\item If the node is congested, feedback should be negative, otherwise positive.
+\end{itemize}
+\end{block}
+\end{column}
 
-\begin{frame}{Sizes}%
-	\showsizes[cm]%
-	\showfonts
+\begin{column}{.5\textwidth}
+\only<2->{
+\begin{block}{Fairness Controller (FC)}
+\begin{itemize}
+\item Divides feedback to packets fairly.
+\item If feedback > 0, equal division.
+\item If feedback < 0, division in proportion to the packets data rates.
+\end{itemize}
+\end{block}
+}
+\end{column}
+\end{columns}
+
+\only<3->{
+\begin{itemize}
+\item Computation of feedback reflecting the current degree of congestion at this node.
+\item Allocates bandwidth equally to all XCP senders.
+\end{itemize}
+}
+\end{frame}
+
+\subsection{Evaluation of Congestion Control}
+\label{sec:eval}
+\begin{frame}{Evaluation of Congestion Control}
+Pros:
+\begin{itemize}
+\item XCP and ECN can easily outperform TCP/IP in certain conditions.
+\item Many networks could benefit from using congestion control.
+\end{itemize}
+\only<2->{
+~\\Cons:
+\begin{itemize}
+\item Many routers run on hardware only $\Rightarrow$ unrecognized packets will be dropped.
+\item Network-assisted congestion control often relies on active queue management.
+\end{itemize}
+}
+~\\
+\begin{itemize}
+\item<3->[$\Rightarrow$] Network-assisted congestion control protocols will not appear widespread in WANs.
+\item<4->[$\Rightarrow$] SANs could implement congestion control more easily.
+\end{itemize}
+\end{frame}
+
+\subsection{DCTP}
+\label{sec:dctp}
+\begin{frame}{DCTCP}
+~\\
+\begin{block}{Data center TCP}
+\begin{itemize}
+\item Extension to TCP for use in data centers.
+\item Based on ECN.
+\end{itemize}
+\end{block}
+\begin{itemize}
+\item<2-> Tackles the TCP incast problem.
+\item<3-> Prevents bursty retransmissions by lowering RTO.
+\item<3-> Prevents link under-utilization:
+\begin{itemize}
+\item <4-> Adjusts to congestion according to the fraction of packets that experienced congestion.
+\end{itemize}
+~\\~\\
+\item<5->[$\Rightarrow$] Implemented in many data centers.
+\end{itemize}
+\end{frame}
+
+\section{Conclusion}
+\label{sec:conclusion}
+\AtBeginSection[]
+{
+  \begin{frame}
+	\frametitle{Outline}
+    \tableofcontents[currentsection]
+  \end{frame}
+}
+
+\begin{frame}{Conclusion}
+\only<1->{
+\begin{block}{Origins of Congestion}
+\begin{itemize}
+\item Buffer size and Buffer management cause delay and packet loss.
+\item Queuing delay results in increased network delay.
+\end{itemize}
+\end{block}
+}
+\only<2->{
+\begin{block}{Congestion Avoidance}
+\begin{itemize}
+\item Simple way to tackle congestion.
+\item Only gets implicit congestion notification.
+\item Often causes instead of resolving network stress.
+\end{itemize}
+\end{block}
+}
+\only<3->{
+\begin{block}{Network-assisted Congestion Control}
+\begin{itemize}
+\item Can explicitly control congestion.
+\item Results in overall better network performance.
+\item Implementations in WANs very challenging, successful in data centers.
+\end{itemize}
+\end{block}
+}
+\end{frame}
+
+% Bibliography
+
+\begin{frame}[noframenumbering,allowframebreaks]{References}
+    \printbibliography
+\end{frame}
+
+% Additional slides for questions
+
+\begin{frame}[noframenumbering]{Typical Data Center Topology}
+\centering
+\begin{figure}
+	\centering
+		\begin{tikzpicture}[node distance = 2.5cm]
+			\node[router, scale=0.8] (Router) {};
+			\node (Internet) [right = 2cm of Router, text width=2cm] {\small Outgoing Connection};
+			\node[switch, scale=0.8] (Sw1) [below left = 1cm and 1.3cm of Router] {};
+			\node[switch, scale=0.8] (Sw2) [below right = 1cm and 1.3cm of Router] {};
+			\node[server, scale=1] (S1) [below left = 1cm and 0.4cm of Sw1] [transition] {Server};
+			\node[server, scale=1] (S2) [below = 3.1cm of Router] [transition] {Server};
+			\node[server, scale=1] (S3) [below right = 1cm and 0.4cm of Sw2] [transition] {Server};
+
+			\draw[-] (Router) -- (Sw1);
+			\draw[-] (Router) -- (Sw2);
+			\draw[-, dotted] (Router) -- (Internet);
+			\draw[-] (Sw1) -- (S1);
+			\draw[-] (Sw1) -- (S2);
+			\draw[-] (Sw1) -- (S3);
+			\draw[-] (Sw2) -- (S1);
+			\draw[-] (Sw2) -- (S2);
+			\draw[-] (Sw2) -- (S3);
+
+			\draw[decorate,decoration={brace,amplitude=10pt},xshift=-4pt,yshift=0pt] (S1.north -| S1.west) -- (Router.center -| S1.west) node [black,midway,xshift=-1cm,align=right] [transition] {High\\ 'Fan-In'};
+		\end{tikzpicture}
+\end{figure}
 \end{frame}

### 2020-07-09 slides/slides.tex
Final Version

@@
 		\begin{itemize}
-		\item Links have limitations.
-		\item<4-> Data is split up into smaller packets.
-		\item<5-> Network nodes contain buffers storing packets.
+		\item Links have limitations
+		\item<4-> Data is split up into smaller packets
+		\item<5-> Network nodes contain buffers storing packets
 		\end{itemize}
-		\only<5->{
+		\only<6->{
 		\begin{alertblock}{}
-		Origins of Congestion are versatile.
+		Origins of Congestion are versatile
 		\end{alertblock}
@@
 \begin{block}{Purpose}
-Buffers store packets in network nodes waiting for transmission.
+Buffers store packets in network nodes waiting for transmission
 \end{block}
@@
 \begin{itemize}
-\item Buffer size should be large enough to handle packet bursts.
-\item Too large buffers result in an increase of queuing delay and \textit{over-buffering}.
+\item Buffer size should be large enough to handle packet bursts
+\item Too large buffers result in an increase of queuing delay and \textit{over-buffering}
 \end{itemize}
@@
 \begin{enumerate}
-\item[$\Rightarrow$] Buffers should be designed according to their application environment.
+\item[$\Rightarrow$] Buffers should be designed according to their application environment
 \end{enumerate}
@@
 \begin{itemize}
-\item Refers to frequently full, excessively large buffers.
-\item<2-> Queued packets will not be dropped, resulting in:
-\only<2->{
+\item Refers to frequently full, excessively large buffers
+~\\~\\
+\item<2-> Queued packets will not be dropped
 	\begin{itemize}
-	\item Unnecessary network delay (\textit{latency}) and packet delay fluctuation (\textit{jitter}).
-	\item Increased congestion, as TCP is not being notified of congestion.
-	\item Decreases network throughput.
+	\item<2->[$\Rightarrow$] Increased congestion, as TCP is not being notified of congestion
+	\item<2->[$\Rightarrow$] Decreased network throughput
+	\item<2->[$\Rightarrow$] Unnecessary network delay (\textit{latency}), packet delay fluctuation (\textit{jitter})
 	\end{itemize}
-}
-\item<3-> With memory getting cheaper over the years, buffers saw an increase in size.
+~\\
+\item<3-> With memory getting cheaper over the years, buffers saw an increase in size
 \end{itemize}
@@
 		\begin{block}{Transmission Delay}
-			Depends on bandwidth and length of the connection.
+			Depends on bandwidth and length of the connection
 		\end{block}
@@
 		\begin{block}{Processing Delay}
-			Time it takes to process a single packet.
+			Time it takes to process a single packet
 		\end{block}
@@
 		\begin{block}{Inter-Packet Delay}
-			Allows the receiver to prepare for transmission.
+			Allows the receiver to prepare for transmission
 		\end{block}
@@
 		\begin{block}{Queuing Delay}
-			Time a packet waits in the buffer waiting for transmission.
+			Time a packet waits in the buffer waiting for transmission
 		\end{block}
@@
 ~\\~\\
-\only<2->{
+\only<3->{
 \begin{itemize}
-\item Congestion control algorithms can improve queuing delay, number of queued packets and buffer management.
+\item Congestion control algorithms improve
+	\begin{itemize}
+	\item Queuing delay
+	\item Buffer lengths
+	\item Buffer management
+	\end{itemize}
 \end{itemize}
@@
 \begin{itemize}
-\item Lives on transport layer, communicates with application layer via sockets ($\rightarrow$ ports) and uses IP as internet protocol ($\rightarrow$ TCP/IP).
-\item Ensures successful transmission of all data.
+\item Transport protocol
+\item Uses IP as internet protocol ($\rightarrow$ TCP/IP)
+\item<2-> Ensures successful transmission of all data
 \end{itemize}
@@
 \begin{itemize}
-\item<2-> Transmitted data needs to be acknowledged.
-\item<2-> Adjusts to the current state of congestion.
-\item<3-> Many different implementations exist.
+\item<3-> Transmitted data needs to be acknowledged
+\item<3-> Adjusts to the current state of congestion
+~\\~\\
+\item<4-> Many different implementations exist
 \end{itemize}
-\end{frame}
-
-\begin{frame}[fragile]{TCP Header}
-\centering
-\begin{figure}
-\centering
-\begin{bytefield}[bitheight=2.2\baselineskip, bitwidth=0.635\baselineskip]{32}
-\bitbox{16}{Source Port} & \bitbox{16}{Destination Port} \\
-\bitbox{32}{Sequence Number} \\
-\bitbox{32}{Acknowledgment Number} \\
-\bitbox{4}{Data Offset} & \bitbox{4}{Rsrvd} & \bitbox{1}{\tiny C\\W\\R} & \bitbox{1}{\tiny E\\C\\E} & \bitbox{1}{\tiny U\\R\\G} & \bitbox{1}{\tiny A\\C\\K} & \bitbox{1}{\tiny P\\S\\H} & \bitbox{1}{\tiny R\\S\\T} & \bitbox{1}{\tiny S\\Y\\N} & \bitbox{1}{\tiny F\\I\\N} & \bitbox{16}{Window Size} \\
-\bitbox{16}{Checksum} & \bitbox{16}{Urgent Pointer} \\
-\bitbox{32}{Options} \\
-\bitbox{32}{Data}
-\end{bytefield}
-\end{figure}
 \end{frame}
@@
 \begin{block}{Round-trip Time (\textit{RTT})}
-Describes the amount of time between sending a data packet and receiving its acknowledgment packet.
+Time between sending a data packet and receiving its acknowledgment packet
 \end{block}
@@
 \begin{block}{Retransmission Timeout (\textit{RTO})}
-The maximum amount of time the sender will wait for the acknowledgment of a packet.
+Fixed upper limit to RTT
 \end{block}
@@
 \end{columns}
-~\\
+~\\~\\
 \only<2->{
 \begin{itemize}
-\item Classic TCP perceives congestion by detecting packet loss.
-\item Other implementations also perceive congestion by determining variations in RTT over time.
+\item Classic TCP perceives congestion by detecting packet loss
+\item<3-> Other implementations also perceive congestion by determining variances in RTT over time
 \end{itemize}
@@
 	\only<4->{\path (r4) edge [above] node {\tiny A} (s4);}
-	\only<5->{\path (r6) edge [above] node {\tiny A} (s6);}
+	\only<5->{\path (r6) edge [above] node {\tiny A,C} (s6);}
 
@@
 \begin{itemize}
-\item Do not fill up buffers in the network.
-\item Achieve and hold maximum throughput.
-\item Adapt to the current state of congestion as quickly as possible.
+\item Do not fill up buffers in the network
+\item Achieve and hold maximum throughput
+\item Quickly adapt to congestion
 \end{itemize}
@@
 \begin{itemize}
-\item CA algorithms are based on Additive-Increase Multiplicative-Decrease principle.
-\item Variable $cwnd$ (congestion window) describes the maximum amount of data sent before ACKs are needed.
+\item Based on Additive-Increase Multiplicative-Decrease principle
+\item<3-> $cwnd$ (congestion window) variable
+	\begin{itemize}
+	\item Current data rate
+	\item Maintained by TCP Sender
+	\end{itemize}
 \end{itemize}
@@
 \begin{block}{Formula}
-$cwnd \coloneqq cwnd + u$
+$cwnd \coloneqq cwnd + u$ for each received ACK packet
 \end{block}
 \begin{itemize}
-\item Increases the data rate slowly to reach the networks maximum throughput.
-\item Prevents the rapid filling of buffers.
-\item Proposed value for $u$: $\frac{1}{cwnd}$
+\item Linear increase in data rate
+\item Slow approach to the networks maximum throughput
+\item Prevents the rapid filling of buffers
+~\\~\\
+\item<2-> Proposed value for $u$: $\frac{1}{cwnd}$
+\end{itemize}
+\end{frame}
+
+\begin{frame}{Adaption - Slow-Start}
+~\\
+\begin{block}{Formula}
+$cwnd \coloneqq cwnd + 1$ for each received ACK packet
+\\$\Rightarrow$ Exponential increase over time
+\end{block}
+\begin{itemize}
+\item Drastic Increase in data rate
+\item Quick increase in protocol performance
+~\\~\\
+\item<2-> Used after
+	\begin{itemize}
+	\item Connection establishment
+	\item Multiple packet losses
+	\end{itemize}
+\item<3-> Used until
+	\begin{itemize}
+	\item Slow-Start threshold has been exceeded
+	\item Perceived congestion
+	\end{itemize}
 \end{itemize}
@@
 \begin{itemize}
-\item Decreases the data rate drastically.
-\item Allows the nodes to recover and buffer lengths to decrease.
-\item Proposed value for $d$: $\frac{1}{2}$
-\end{itemize}
-\end{frame}
-
-\begin{frame}{Adaption - Slow-Start}
-~\\
-\begin{block}{Formula}
-$cwnd \coloneqq cwnd + 1$ for each received ACK packet.
-\\$\Rightarrow$ Exponential increase over time.
-\end{block}
-\begin{itemize}
-\item Drastically increases the protocols performance.
-\item Used after connection is established or after multiple packet losses have been detected.
-\item Switches off after a threshold has been exceeded or congestion has been perceived.
+\item Decreases the data rate drastically
+\item Allows congested nodes and buffers to recover
+~\\~\\
+\item<2-> Proposed value for $d$: $\frac{1}{2}$
 \end{itemize}
@@
 \begin{itemize}
-\item Congestion Detection algorithms are very implicit:
+\item Congestion Detection algorithms are very implicit
 	\begin{itemize}
 	\item Packet loss may also have other causes
-	\item Methods are inaccurate, and do not inform about the degree of congestion.
+	\item No information about the degree of congestion
 	\end{itemize}
-\item[$\Rightarrow$] Creates unnecessary network delay.
+\item[$\Rightarrow$] Creates unnecessary network delay
 ~\\~\\
-\item<2-> Explicit congestion notification is crucial to keep the network stable.
-\item[$\Rightarrow$]<2-> Modify the IP header to include a \textit{Congestion Experienced} Flag (CE).
-\item<3-> \textit{Explicit congestion notification} (ECN), an extention to TCP/IP, implements this.
+\item<2-> \textit{Explicit congestion notification} (ECN)
+	\begin{itemize}
+	\item Extension to TCP/IP
+	\item Adds \textit{Congestion Experienced} Flag (CE) to the IP header
+	\end{itemize}
 \end{itemize}
@@
 			\path [line] (R3) -- (H2);
-			\path [line, dotted] (H2) -- ++(0,-35pt) -| (H1);
-			\node[below=0.9cm of R2] (text) [transition] {Congested Signal};
+			\only<5->{\path [line, dotted] (H2) -- ++(0,-35pt) -| (H1);}
+			\only<5->{\node[below=0.9cm of R2] (text) [transition] {Congested Signal};}
 
@@
 			\node[messageclosed,scale=0.5, fill=black!10] (P1) [above right=0.2cm and 1cm of H1] {};
-			\node[messageclosed,scale=0.5, fill=black!50] (P2) [above left=0.7cm and 0.3cm of R2] {};
-			\node[messageclosed,scale=0.5, fill=black!10] (P3) [above left=0.4cm and 0.2cm of R2] {};
-			\node[messageclosed,scale=0.5, fill=black!10] (P4) [above left=0.1cm and 0.1cm of R2] {};
-			\node[messageclosed,scale=0.5, fill=black!50] (P5) [above right=0.2cm and 1cm of R3] {};
+			\only<3->{\node[messageclosed,scale=0.5, fill=black!50] (P2) [above left=0.7cm and 0.3cm of R2] {};}
+			\only<2->{\node[messageclosed,scale=0.5, fill=black!10] (P3) [above left=0.4cm and 0.2cm of R2] {};}
+			\only<2->{\node[messageclosed,scale=0.5, fill=black!10] (P4) [above left=0.1cm and 0.1cm of R2] {};}
+			\only<4->{\node[messageclosed,scale=0.5, fill=black!50] (P5) [above right=0.2cm and 1cm of R3] {};}
 			\node[above left = 1.5cm and 0cm of R1] (P) [transition] {Packet};
-			\node[above right = 1.5cm and -0.3cm of R2] (CN) [transition] {Congested Node};
-			\node[above left = 1.5cm and -0.4cm of H2] (MP) [transition] {Marked Packet};
+			\only<2->{\node[above right = 1.5cm and -0.3cm of R2] (CN) [transition] {Congested Node};}
+			\only<4->{\node[above left = 1.5cm and -0.4cm of H2] (MP) [transition] {Marked Packet};}
 			\path [line] (P) -- (P1);
-			\path [line] (CN) -- (R2);
-			\path [line] (MP) -- (P5);
+			\only<2->{\path [line] (CN) -- (R2);}
+			\only<4->{\path [line] (MP) -- (P5);}
 		\end{tikzpicture}
@@
 \begin{block}{XCP}
-\textit{eXplicit Congestion Protocol} generalizes the ECN proposal.
+\textit{eXplicit Congestion Protocol} generalizes the ECN proposal
 \end{block}
 \begin{itemize}
-\item Improves upon ECN by introducing new packet control concepts.
-\item Additionally gives feedback about the degree of congestion.
+\item Introduces new packet control concepts
+\item Gives feedback about the degree of congestion
 \end{itemize}
+~\\
 \only<2->{
@@
 \begin{itemize}
-\item As a TCP Sender, the XCP sender also keeps track of $cwnd$ and $RTT$ and deals with packet loss (retransmission).
+\item Maintains $cwnd$ and $RTT$
+\item Handles packet loss
 ~\\~\\
-\item On data packet depature, the XCP sender will:
+\item<2-> On data packet departure
 	\begin{itemize}
@@
 	\item $H\_rtt \coloneqq RTT$
-	\item $H\_feedback \coloneqq \frac{r \cdot RTT - cwnd}{cwnd}$, where $r$ is the desired data rate.
+	\item $H\_feedback \coloneqq \frac{r \cdot RTT - cwnd}{cwnd}$, where $r$ is the desired data rate
 	\end{itemize}
-\item[$\Rightarrow$] The desired data rate can be reached in just one RTT.
+\item<2-> [$\Rightarrow$] The desired data rate can be reached in just one RTT
 ~\\~\\
-\item On ACK packet arrival, the XCP sender will adjust $cwnd$ according to the received feedback:
+\item<3-> On ACK packet arrival
 	\begin{itemize}
-	\item $cwnd \coloneqq max(cwnd + H\_feedback, s)$, where $s$ is the packet size.
+	\item $cwnd \coloneqq max(cwnd + H\_feedback, s)$, where $s$ is the packet size
 	\end{itemize}
@@
 \begin{itemize}
-\item Maximizes link utilization.
-\item Minimizes queue lengths and packet drops.
-\item If the node is congested, feedback should be negative, otherwise positive.
+\item Maximizes link utilization
+\item Minimizes queue lengths and packet drops
 \end{itemize}
@@
 \begin{itemize}
-\item Divides feedback to packets fairly.
-\item If feedback > 0, equal division.
-\item If feedback < 0, division in proportion to the packets data rates.
+\item Divides feedback to all packets fairly
 \end{itemize}
@@
 \end{columns}
-
+~\\
 \only<3->{
 \begin{itemize}
-\item Computation of feedback reflecting the current degree of congestion at this node.
-\item Allocates bandwidth equally to all XCP senders.
+\item Computed feedback reflects the degree of congestion until this router
 \end{itemize}
+}
+~\\
+\only<4->{
+\begin{alertblock}{}
+Allocates bandwidth equally to all XCP senders
+\end{alertblock}
 }
@@
 \begin{itemize}
-\item XCP and ECN can easily outperform TCP/IP in certain conditions.
-\item Many networks could benefit from using congestion control.
+\item XCP and ECN outperforms TCP/IP
 \end{itemize}
@@
 \begin{itemize}
-\item Many routers run on hardware only $\Rightarrow$ unrecognized packets will be dropped.
-\item Network-assisted congestion control often relies on active queue management.
+\item Many routers run on hardware only 
+	\begin{itemize}
+	\item[$\Rightarrow$] Unrecognized packets will be dropped
+	\end{itemize}
+\item Relies on active queue management
 \end{itemize}
@@
 \begin{itemize}
-\item<3->[$\Rightarrow$] Network-assisted congestion control protocols will not appear widespread in WANs.
-\item<4->[$\Rightarrow$] SANs could implement congestion control more easily.
+\item<3->[$\Rightarrow$] Not used in WANs
+\item<4->[$\Rightarrow$] Implementation in SANs more easily
 \end{itemize}
@@
 \begin{itemize}
-\item Extension to TCP for use in data centers.
-\item Based on ECN.
+\item Extension to TCP for use in data centers
+\item Based on ECN
 \end{itemize}
@@
 \begin{itemize}
-\item<2-> Tackles the TCP incast problem.
-\item<3-> Prevents bursty retransmissions by lowering RTO.
-\item<3-> Prevents link under-utilization:
+\item<2-> Tackles the TCP incast problem
+\item<3-> Prevents bursty retransmissions by lowering RTO
+\item<3-> Prevents link under-utilization
 \begin{itemize}
-\item <4-> Adjusts to congestion according to the fraction of packets that experienced congestion.
+\item <4-> Adjusts to congestion according to the fraction of packets that experienced congestion
 \end{itemize}
 ~\\~\\
-\item<5->[$\Rightarrow$] Implemented in many data centers.
+\item<5->[$\Rightarrow$] Implemented in many data centers
 \end{itemize}
@@
 \begin{itemize}
-\item Buffer size and Buffer management cause delay and packet loss.
-\item Queuing delay results in increased network delay.
+\item Buffer size and Buffer management cause delay and packet loss
+\item Queuing delay results in increased network delay
 \end{itemize}
@@
 \begin{itemize}
-\item Simple way to tackle congestion.
-\item Only gets implicit congestion notification.
-\item Often causes instead of resolving network stress.
+\item Simple way to tackle congestion
+\item Only gets implicit congestion notification
+\item Often causes instead of resolving network stress
 \end{itemize}
@@
 \begin{itemize}
-\item Can explicitly control congestion.
-\item Results in overall better network performance.
-\item Implementations in WANs very challenging, successful in data centers.
+\item Can explicitly control congestion
+\item Results in overall better network performance
+\item Implementations in WANs very challenging, successful in data centers
 \end{itemize}
